---
title: "Descritption of the Covid-19 epidemic in Europe and Western and Central Asia"
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: Francesco Grandesso Epicentre (MSF)
output: 
  rmdformats::readthedown:
    use_bookdown: TRUE
editor_options: 
  chunk_output_type: console
---


```{r setup, include = FALSE}

knitr::opts_chunk$set(cache = FALSE, 
                      echo = FALSE, 
                      tidy = TRUE, 
                      collapse = TRUE, 
                      dpi = 150, 
                      fig.align = "center", 
                      warning = FALSE, 
                      message = FALSE, 
                      encoding = "UTF-8")

source(here::here('R', 'setup.R'), encoding = 'UTF-8')

library(treemap)
library(patchwork)

source(file.path(path.R, "utils_management.R"), encoding = "UTF-8")
source(file.path(path.R, "utils_vis.R")       , encoding = "UTF-8")


date_min <- NULL
date_max <- as.Date('2020-11-08')
week_report <- ISOweek::ISOweek(date_max) %>% gsub("W","w", .)

set_paths(path.local, week_report)

```

```{r upload-data, include = FALSE}

## --- ECDC Covid data
dta_ecdc <- get_ecdc_data() %>% 
  prepare_ecdc_data() %>% 
  tidyr::drop_na(iso_a3) %>% 
  filter(between(date, left = NULL, right = date_max) & (continent == 'Europe' | region %in% c('Western Asia', 'Central Asia')))


## ---  Population data
df_countries <- readRDS(file.path(path.local.data, paste0('df_countries','.RDS'))) %>% 
  filter(continent == 'Europe' | region %in% c('Western Asia', 'Central Asia')) %>% 
  left_join(readRDS(file.path(path.local.data, paste0('df_pop_country','.RDS'))))

# sf Robinson projection
sf_robinson <- readRDS(file.path(path.local.data, paste0('sf_world','.RDS'))) %>% 
  filter(continent == 'Europe' | region %in% c('Western Asia', 'Central Asia'))

## --- GIS data - Mercator projection
sf_mercator <- rnaturalearth::ne_countries(type = "countries", returnclass = "sf") %>% 
  tibble::as_tibble() %>% 
  sf::st_as_sf() %>% 
  dplyr::select(country = name_long, iso_a3, iso_a2, pop_est) %>% 
  dplyr::mutate(
    continent = suppressWarnings(countrycode::countrycode(iso_a3, origin = "iso3c", destination = "continent")),
    region = suppressWarnings(countrycode::countrycode(iso_a3, origin = "iso3c", destination = "region23"))
  ) %>% 
  dplyr::filter(stringr::str_detect(country, "Antarctic", negate = TRUE)) %>% 
  cbind(sf::st_coordinates(suppressWarnings(sf::st_centroid(., of_largest_polygon = TRUE)))) %>% 
  dplyr::rename(lon = X, lat = Y) 


# Specific path for outputs of this analysis
path.local.Europe <- file.path(path.local, 'Europe')
dir.create(path.local.Europe, showWarnings = FALSE, recursive = TRUE) 

```

```{r prepare-data, include = FALSE}

## Grouping regions

grp_region_EuropeNSW      <- c('Northern Europe', 'Southern Europe', 'Western Europe')
grp_region_EuropeE_AsiaWC <- c('Eastern Europe', 'Western Asia', 'Central Asia')


## CASES Incidence proportion

tbl_cases_prop <- dta_ecdc %>% 
  group_by(country) %>% 
  summarise(
    n = sum(cases)) %>% 
  left_join(df_countries) %>% 
  mutate(
   p = n/pop * 100)

tbl_cases_prop_2w <- dta_ecdc %>% 
  filter(between(date, left = date_max - 13, right = date_max)) %>%
  group_by(country) %>% 
  summarise(
    n_2w = sum(cases)) %>% 
    left_join(df_countries) %>% 
  mutate(
   p_2w = n_2w/pop * 100) %>% 
  left_join(tbl_cases_prop) %>% 
  select(region, country, n, p, n_2w, p_2w)


## DEATHS Incidence proportion

tbl_deaths_prop <- dta_ecdc %>% 
  group_by(country) %>% 
  summarise(
    n = sum(deaths)) %>% 
  left_join(df_countries) %>% 
  mutate(
   p = n/pop * 100)

tbl_deaths_prop_2w <- dta_ecdc %>% 
  filter(between(date, left = date_max - 13, right = date_max)) %>%
  group_by(country) %>% 
  summarise(
    n_2w = sum(deaths)) %>% 
    left_join(df_countries) %>% 
  mutate(
   p_2w = n_2w/pop * 100) %>% 
  left_join(tbl_deaths_prop) %>% 
  select(region, country, n, p, n_2w, p_2w)

```


Data source: [ECDC data](https://opendata.ecdc.europa.eu/covid19/casedistribution/csv) updated until `r format(date_max, '%d %B %Y')`

**How to read the treemaps**

**Treemapping** is a method to display a set of tiled rectangles whose area is proportional to a specified dimension of the data. The rectangles are filled with a sequential colour scheme representing a second dimension of the data. When the size and colour dimensions are correlated, one can often easily see patterns that would be difficult to spot in other ways.

The **treemaps presented in this report** display the size of the rectangles proportional to the total number of cases (or deaths) in each country, and the colour of the rectangles representing a second variable whose definition and colour scheme is displayed in the legend. 

The relation of the the two variables are also presented in scatter plots.


# Overview

## Total number and incidence proportion of confirmed cases

### `r combine_words(grp_region_EuropeNSW)`

```{r, fig.cap = paste("TREEMAP: Total number and incidence proportion of confirmed cases per country in", combine_words(grp_region_EuropeNSW))}

tbl_cases_prop %>% 
  filter(region %in% grp_region_EuropeNSW) %>% 
  treemap(index = c('region', 'country'), 
          vSize = 'n', 
          vColor = 'p',
          type = 'dens', 
          palette = 'YlGn', 
          range = c(0, max(tbl_cases_prop$p)), 
          title = 'Total cases', 
          title.legend = 'Incidence proportion of cases (%)')

```

```{r, fig.cap = paste("SCATTER PLOT: Total number and incidence proportion of confirmed cases per country in", combine_words(grp_region_EuropeNSW))}

countries_zero_count <- tbl_cases_prop %>% 
  filter(n == 0, region %in% grp_region_EuropeNSW) %>% 
  pull(country)

txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))

ggplot(tbl_cases_prop %>% filter(n != 0, region %in% grp_region_EuropeNSW)) + 
  geom_point(aes(n, p, col = region)) + 
  ggrepel::geom_text_repel(aes(n, p, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 5), 
                     labels = scales::label_number_si()) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    colour = 'Region', 
    caption = paste0("Countries with zero count: \n", txt_countries_zero_count), 
    x = "Total count of cases (log scale)",
    y = "Incidence proportion % (log scale)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```

### `r combine_words(grp_region_EuropeE_AsiaWC)`

```{r, fig.cap = paste("TREEMAP: Total number and incidence proportion of confirmed cases per country in", combine_words(grp_region_EuropeE_AsiaWC))}

tbl_cases_prop %>% 
  filter(region %in% grp_region_EuropeE_AsiaWC) %>% 
  treemap(index = c('region', 'country'), 
          vSize = 'n', 
          vColor = 'p',
          type = 'dens', 
          palette = 'YlGn', 
          range = c(0, max(tbl_cases_prop$p)), 
          title = 'Total cases', 
          title.legend = 'Incidence proportion of cases (%)')

```

```{r, fig.cap = paste("SCATTER PLOT: Total number and incidence proportion of confirmed cases per country in", combine_words(grp_region_EuropeE_AsiaWC))}

countries_zero_count <- tbl_cases_prop %>% 
  filter(n == 0, region %in% grp_region_EuropeE_AsiaWC) %>% 
  pull(country)

txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))

ggplot(tbl_cases_prop %>% filter(n != 0, region %in% grp_region_EuropeE_AsiaWC)) + 
  geom_point(aes(n, p, col = region)) + 
  ggrepel::geom_text_repel(aes(n, p, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 5), 
                     labels = scales::label_number_si()) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    colour = 'Region', 
    caption = paste0("Countries with zero count: \n", txt_countries_zero_count), 
    x = "Total count of cases (log scale)",
    y = "Incidence proportion % (log scale)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```


## Total number and incidence proportion of associated deaths

### `r combine_words(grp_region_EuropeNSW)`

```{r, fig.cap = paste("TREEMAP: Total number and incidence proportion of associated deaths per country in", combine_words(grp_region_EuropeNSW))}

tbl_deaths_prop %>% 
  filter(region %in% grp_region_EuropeNSW) %>% 
  treemap(index = c('region', 'country'), 
          vSize = 'n', 
          vColor = 'p',
          type = 'dens', 
          palette = 'OrRd', 
          range = c(0, max(tbl_deaths_prop$p)), 
          title = 'Total deaths', 
          title.legend = 'Incidence proportion of deaths (%)')

```

### `r combine_words(grp_region_EuropeE_AsiaWC)`

```{r, fig.cap = paste("TREEMAP: Total number and incidence proportion of associated deaths per country in", combine_words(grp_region_EuropeE_AsiaWC))}

tbl_deaths_prop %>% 
  filter(region %in% grp_region_EuropeE_AsiaWC) %>% 
  treemap(index = c('region', 'country'), 
          vSize = 'n', 
          vColor = 'p',
          type = 'dens', 
          palette = 'OrRd', 
          range = c(0, max(tbl_deaths_prop$p)), 
          title = 'Total deaths', 
          title.legend = 'Incidence proportion of deaths (%)')

```


# During the last two weeks

## Number of confirmed cases in the last 14 days

### `r combine_words(grp_region_EuropeNSW)`

```{r, fig.cap = paste("Number of confirmed cases since the beginning of the epidemic and in the last 14 days per country in", combine_words(grp_region_EuropeNSW))}

tbl_cases_prop_2w %>% 
  filter(region %in% grp_region_EuropeNSW) %>% 
  treemap(index = c('region', 'country'), 
          vSize = 'n', 
          vColor = 'n_2w',
          type = 'dens', 
          palette = 'YlGn', 
          range = c(0, max(tbl_cases_prop_2w$n_2w)), 
          title = 'Total cases', 
          title.legend = 'Cases in the last 14 days')

```

### `r combine_words(grp_region_EuropeE_AsiaWC)`

```{r, fig.cap = paste("Number of confirmed cases since the beginning of the epidemic and in the last 14 days per country in", combine_words(grp_region_EuropeE_AsiaWC))}

tbl_cases_prop_2w %>% 
  filter(region %in% grp_region_EuropeE_AsiaWC) %>% 
  treemap(index = c('region', 'country'), 
          vSize = 'n', 
          vColor = 'n_2w',
          type = 'dens', 
          palette = 'YlGn', 
          range = c(0, max(tbl_cases_prop_2w$n_2w)), 
          title = 'Total cases', 
          title.legend = 'Cases in the last 14 days')

```


## Number of associated deaths in the last 14 days

### `r combine_words(grp_region_EuropeNSW)`

```{r, fig.cap = paste("Number of confirmed associated deaths since the beginning of the epidemic and in the last 14 days per country in", combine_words(grp_region_EuropeNSW))}

tbl_deaths_prop_2w %>% 
  filter(region %in% grp_region_EuropeNSW) %>% 
  treemap(index = c('region', 'country'), 
          vSize = 'n', 
          vColor = 'n_2w',
          type = 'dens', 
          palette = 'OrRd', 
          range = c(0, max(tbl_deaths_prop_2w$n_2w)), 
          title = 'Total deaths', 
          title.legend = 'Deaths in the last 14 days')

```

### `r combine_words(grp_region_EuropeE_AsiaWC)`

```{r, fig.cap = paste("Number of confirmed associated deaths since the beginning of the epidemic and in the last 14 days per country in", combine_words(grp_region_EuropeE_AsiaWC))}

tbl_deaths_prop_2w %>% 
  filter(region %in% grp_region_EuropeE_AsiaWC) %>% 
  treemap(index = c('region', 'country'), 
          vSize = 'n', 
          vColor = 'n_2w',
          type = 'dens', 
          palette = 'OrRd', 
          range = c(0, max(tbl_deaths_prop_2w$n_2w)), 
          title = 'Total deaths', 
          title.legend = 'Deaths in the last 14 days')

```


## Overall incidence proportions _versus_ incidence proportions in the last 14 days

### Maps

<!-- Map of overall incidence proportion -->
```{r maps-incidence-proportions, fig.cap = "MAP: Overall incidence proportion of cases and recent incidence proportions of cases (last 14 days)"}

## Breaks for incidence
breaks_inc <- c(0, 0.01, 0.1, 1, Inf)
labels_inc <- c("0 - 0.01%", "0.01 - 0.1%", "0.1 - 1%", "1%+")


## --- OVERALL ---

## Put incidence table in to sf
sf_case_inc <- df_countries %>% 
  left_join(tbl_cases_prop) %>% 
  mutate(
    inc_brks = cut(p, breaks_inc, labels = labels_inc, include.lowest = TRUE, right = FALSE)) %>% 
  full_join(
    select(sf_mercator, iso_a3),
    by = "iso_a3"
  ) %>% 
  st_as_sf()

## Map 
map_case_inc <- ggplot(sf_case_inc) + 
  geom_sf(aes(fill = inc_brks), size = .1) + 
  scale_fill_brewer(
    name = 'Incidence proportion', 
    palette = "Blues", 
    na.value = "lightgrey", 
    drop = FALSE, 
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(70 / length(labels_inc), units = "mm"),
      title.hjust = 0.5,
      nrow = 1,
      label.position = "bottom",
      title.position = 'top')) + 
  coord_sf(
    xlim = c(-30, 80), 
    ylim = c(28, 73), 
    expand = FALSE) + 
  labs(title = 'Overall') + 
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        legend.position = "bottom")


## --- LAST TWO WEEKS ---

## Put incdence table (last 2 weks) in to sf
sf_case_inc_2weeks <- df_countries %>% 
  left_join(tbl_cases_prop_2w) %>% 
  mutate(
    inc_brks = cut(p_2w, breaks_inc, labels = labels_inc, include.lowest = TRUE, right = FALSE)) %>% 
  full_join(
    select(sf_mercator, iso_a3),
    by = "iso_a3"
  ) %>% 
  st_as_sf()

## Map
map_case_inc_2weeks <- ggplot(sf_case_inc_2weeks) + 
  geom_sf(aes(fill = inc_brks), size = .1) + 
  scale_fill_brewer(
    name = 'Incidence proportion', 
    palette = "Blues", 
    na.value = "lightgrey", 
    drop = FALSE, 
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(70 / length(labels_inc), units = "mm"),
      title.hjust = 0.5,
      nrow = 1,
      label.position = "bottom",
      title.position = 'top')) + 
  coord_sf(
    xlim = c(-30, 80), 
    ylim = c(28, 73), 
    expand = FALSE) + 
  labs(title = 'Last 14 days') + 
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        legend.position = "bottom")


## Combine maps
map_case_inc + 
  map_case_inc_2weeks + 
  plot_layout(ncol = 2, guides = 'collect') & theme(legend.position = 'bottom')

```

### Scatter plots

```{r, fig.cap = paste("SCATTER PLOT: Overall incidence proportion of cases and recent incidence proportions of cases (last 14 days), in", combine_words(grp_region_EuropeNSW))}

tbl_case_inc_vs_inc_2weeks <- tbl_cases_prop_2w %>% 
  filter(region %in% grp_region_EuropeNSW) %>% 
  select(region, country, overall = p, last2weeks = p_2w)

countries_zero_count <- tbl_case_inc_vs_inc_2weeks %>% 
  filter(last2weeks == 0) %>% 
  pull(country)

txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))


# The dots plot
ggplot(tbl_case_inc_vs_inc_2weeks %>% filter(last2weeks != 0)) + 
  geom_point(aes(overall, last2weeks, col = region)) + 
  ggrepel::geom_text_repel(aes(overall, last2weeks, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    colour = 'Region', 
    caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    x = "Incidence proportion since the beginning of the epidemic % (log scale)",
    y = "Incidence proportion in the last 14 days % (log scale)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```

```{r, fig.cap = paste("SCATTER PLOT: Overall incidence proportion of cases and recent incidence proportions of cases (last 14 days), in", combine_words(grp_region_EuropeE_AsiaWC))}

tbl_case_inc_vs_inc_2weeks <- tbl_cases_prop_2w %>% 
  filter(region %in% grp_region_EuropeE_AsiaWC) %>% 
  select(region, country, overall = p, last2weeks = p_2w)

countries_zero_count <- tbl_case_inc_vs_inc_2weeks %>% 
  filter(last2weeks == 0) %>% 
  pull(country)

txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))


# The dots plot
dotsgplot <- ggplot(tbl_case_inc_vs_inc_2weeks %>% filter(last2weeks != 0)) + 
  geom_point(aes(overall, last2weeks, col = region)) + 
  ggrepel::geom_text_repel(aes(overall, last2weeks, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    colour = 'Region', 
    caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    x = "Incidence proportion since the beginning of the epidemic % (log scale)",
    y = "Incidence proportion in the last 14 days % (log scale)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

dotsgplot

```



# Trends

```{r}
# Get trends

linear_trend <- function(dta, series, last_date, n_days = 12, ma_window = 3, min_sum = 30){
  
  dta <- dta %>% select(date, cnt = all_of(series))
  
  dates_extent <- c(last_date - (n_days - 1), last_date)
    
  dta <- dta %>% 
    filter(between(date, dates_extent[1], dates_extent[2])) %>% 
    tidyr::complete(date = seq.Date(min(date, na.rm = TRUE), 
                                    max(date, na.rm = TRUE), by = 1), 
                    fill = list(cnt = NA_real_))
  
  
  # Moving average
  dta$ma <- forecast::ma(dta$cnt, order = ma_window)
  dta$ma <- na_if(dta$ma, 0) # Replace 0 values as NA
  
  
  # Empty matrix of predictions
  m_preds <- matrix(data = NA, 
                    nrow = dim(dta)[1], 
                    ncol = 3, 
                    dimnames = list(c(1:dim(dta)[1]), c('fit', 'lwr', 'upr')))
  
  
  # Run model with conditions
  if (dim(dta)[1] > ma_window & sum(dta$cnt, na.rm = TRUE) > min_sum) {
    
    mdl <- lm(log(ma) ~ date, data = dta)
    
    preds <- exp(predict(mdl, interval = 'confidence'))
    
    matched_rows <- match(rownames(preds), rownames(m_preds))
    matched_cols <- match(colnames(preds), colnames(m_preds))
    m_preds[matched_rows, matched_cols] <- preds
    
    tbl_preds <- tibble(date = seq.Date(from = dates_extent[1], to = dates_extent[2], by = 1), 
                        cnt  = dta$cnt, 
                        ma   = dta$ma, 
                        fit  = as.double(m_preds[, 'fit']), 
                        lwr  = as.double(m_preds[, 'lwr']), 
                        upr  = as.double(m_preds[, 'upr']))
    
    mdl_coeffs <- tibble(coeff = coefficients(mdl)[[2]], 
                         lwr   = confint(mdl)[2,1], 
                         upr   = confint(mdl)[2,2])
    
  } else {
    mdl <- NA_character_
    
    tbl_preds <- tibble(date = seq.Date(from = dates_extent[1], to = dates_extent[2], by = 1), 
                        cnt  = dta$cnt, 
                        ma   = dta$ma, 
                        fit  = as.double(m_preds[, 'fit']), 
                        lwr  = as.double(m_preds[, 'lwr']), 
                        upr  = as.double(m_preds[, 'upr']))
    
    mdl_coeffs <- tibble(coeff = NA_real_, 
                         lwr   = NA_real_, 
                         upr   = NA_real_)
  }
  
  return(list(mdl = mdl, preds = tbl_preds, coeffs = mdl_coeffs))
}
```

```{r}
dta_ecdc$date

i <- incidence::incidence(dates = dta_ecdc$date, groups  = dta_ecdc$country)
i

plot(i) + theme_bw()

```



```{r}
lst_dta_ecdc <- dta_ecdc %>% 
  multisplit("iso_a3")


# -- Linear model trends - last 14 days 

lst_mdl_cases_14d <- list()

for(i in names(lst_dta_ecdc)){
  mdl <- linear_trend(dta = lst_dta_ecdc[[i]], series = 'cases', n_days = 14, last_date = date_max)
  lst_mdl_cases_14d[[i]] <- mdl
}


lst_mdl_deaths_14d <- list()

for(i in names(lst_dta_ecdc)){
  mdl <- linear_trend(dta = lst_dta_ecdc[[i]], series = 'deaths', n_days = 14, last_date = date_max)
  lst_mdl_deaths_14d[[i]] <- mdl
}


# -- Linear model trends - last 30 days 

lst_mdl_cases_30d <- list()

for(i in names(lst_dta_ecdc)){
  mdl <- linear_trend(dta = lst_dta_ecdc[[i]], series = 'cases', n_days = 30, last_date = date_max)
  lst_mdl_cases_30d[[i]] <- mdl
}


lst_mdl_deaths_30d <- list()

for(i in names(lst_dta_ecdc)){
  mdl <- linear_trend(dta = lst_dta_ecdc[[i]], series = 'deaths', n_days = 30, last_date = date_max)
  lst_mdl_deaths_30d[[i]] <- mdl
}

```

```{r, eval = FALSE}
# To plot both cases and deaths into a single graphic plot
country_six_plots2 <- function(dta_all, dta_long, dta_short, date_max) {
    
  # Parameters
  main_colour  <- c(cases = '#1A62A3', deaths = '#e10000')
  date_min     <- dta_all %>% filter(cnt != 0) %>% pull(date) %>% min()

  
  # Plots
  plot_all <- ggplot(dta_all, aes(x = date, y = cnt)) + 
    facet_wrap(~obs, scales = "free_y", ncol = 1) + 
    geom_col(aes(colour = obs, fill = obs)) + 
    scale_colour_manual(values = main_colour) + 
    scale_fill_manual(values = main_colour) + 
    scale_x_date(limits = c(date_min, NA), breaks = '2 months', date_labels = "%b-%Y") +
    xlab('') + 
    ylab('frequency') + 
    labs(subtitle = 'Since the first cases reported') + 
    theme_light() + 
    theme(legend.position = "none", 
          strip.text = element_text(size = 11))
  
  plot_long <- ggplot(dta_long, aes(x = date, y = cnt)) + 
    facet_wrap(~ obs, scales = "free_y", ncol = 1) + 
    geom_point(aes(colour = obs), size = 2) + 
    scale_colour_manual(values = main_colour) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr, fill = obs), alpha = 0.4) + 
    geom_line(aes(y = fit, colour = obs), size = 1) + 
    scale_fill_manual(values = main_colour) + 
    scale_x_date(limits = c(min(dta_long$date), max(dta_long$date)), date_labels = "%d-%b") +
    xlab('') + 
    ylab(paste0('frequency and fitted values')) + 
    labs(subtitle = paste('Last', (max(dta_long$date) - min(dta_long$date) + 1), 'days')) + 
    theme_light() + 
    theme_light() + 
    theme(legend.position = "none", 
          strip.text = element_text(size = 11))
  
  plot_short <- ggplot(dta_short, aes(x = date, y = cnt)) + 
    facet_wrap(~ obs, scales = "free_y", ncol = 1) + 
    geom_point(aes(colour = obs), size = 2) + 
    scale_colour_manual(values = main_colour) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr, fill = obs), alpha = 0.4) + 
    geom_line(aes(y = fit, colour = obs), size = 1) + 
    scale_fill_manual(values = main_colour) + 
    scale_x_date(limits = c(min(dta_short$date), max(dta_short$date)), breaks = '4 days', date_labels = "%d-%b") +
    xlab('') + 
    ylab(paste0('frequency and fitted values')) + 
    labs(subtitle = paste('Last', (max(dta_short$date) - min(dta_short$date) + 1), 'days')) + 
    theme_light() + 
    theme_light() + 
    theme(legend.position = "none", 
          strip.text = element_text(size = 11))
  
  
  # Arrange plots
  multiplot <- ggarrange(plot_all, 
                         plot_long, 
                         plot_short, 
                         ncol = 3, 
                         widths = c(2, 1.4, 1.1))
  
  return(multiplot)
}


# -- Lists of predictions
lst_preds_cases_30d <- lapply(lst_mdl_cases_30d, function (x) x[['preds']])
lst_preds_cases_14d <- lapply(lst_mdl_cases_14d, function (x) x[['preds']])

lst_preds_deaths_30d <- lapply(lst_mdl_deaths_30d, function (x) x[['preds']])
lst_preds_deaths_14d <- lapply(lst_mdl_deaths_14d, function (x) x[['preds']])


country_list <- df_countries

for(i in names(lst_dta_ecdc)){
  
  #i = 'USA'
  name_country <- country_list %>% filter(iso_a3 == i) %>% pull(country) %>% gsub(" ", "_", .)

  dta_all <- lst_dta_ecdc[[i]] %>% 
    select(date, cases, deaths) %>% 
    pivot_longer(-date, names_to = 'obs', values_to = 'cnt')
  
  dta_preds_long <- lst_preds_cases_30d[[i]] %>% 
    mutate(obs = 'cases') %>% 
    add_row(lst_preds_deaths_30d[[i]] %>% 
              mutate(obs = 'deaths')) %>% 
    arrange(date, obs)
  
  dta_preds_short <- lst_preds_cases_14d[[i]] %>% 
    mutate(obs = 'cases') %>% 
    add_row(lst_preds_deaths_14d[[i]] %>% 
              mutate(obs = 'deaths')) %>% 
    arrange(date, obs)
  
  country_multiplot <- country_six_plots2(dta_all, dta_preds_long, dta_preds_short)
  
  ggsave(filename = file.path(path.local.Europe, paste0('trends_', name_country, '.png')),
         plot = country_multiplot, 
         scale = 1, 
         width = 9,
         dpi = 320)
}
  
```


## Growth rate vs cumulative incidence rate in the last 30 days

```{r, fig.width = 6, fig.height = 6}

# Cumulative case count
tbl_case_count_30days <- dta_ecdc %>% 
  filter(between(date, left = date_max - 29, right = date_max)) %>% 
  filter(region %in% c('Northern Europe', 'Southern Europe', 'Western Europe')) %>% 
  group_by(iso_a3, region, country) %>% 
  summarise(cases = sum(cases, na.rm = TRUE)) %>% 
  ungroup()

# -- Incidence rate
tbl_case_inc_30days <- df_countries %>% 
  filter(region %in% c('Northern Europe', 'Southern Europe', 'Western Europe')) %>% 
  left_join(tbl_case_count_30days) %>% 
  mutate(
    incdc = cases / pop * 100)

# -- Growth rate
tbl_growth_rate_cases_30d <- lapply(lst_mdl_cases_30d, function (x) x[['coeffs']]) %>% 
  bind_rows(.id = "iso_a3") %>% 
  transform(gr = exp(coeff) - 1) %>% 
  select(iso_a3, gr)

# -- Plot
tbl_30days <- tbl_case_inc_30days %>% 
  left_join(tbl_growth_rate_cases_30d)

ggplot(data = tbl_30days) + 
  geom_point(aes(x = gr, y = incdc, col = region)) + 
  ggrepel::geom_text_repel(aes(gr, incdc, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_x_continuous(labels = scales::label_percent(scale = 100)) + 
  labs(
    #caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    y = "Cumulative incidence rate in the last 30 days (log scale %)",
    x = "Estimated growth rate in cases during the last 30 days (%)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```


```{r, fig.width = 6, fig.height = 6}

# Cumulative case count
tbl_case_count_30days <- dta_ecdc %>% 
  filter(between(date, left = date_max - 29, right = date_max)) %>% 
  filter(region %in% c('Eastern Europe', 'Western Asia', 'Central Asia')) %>% 
  group_by(iso_a3, region, country) %>% 
  summarise(cases = sum(cases, na.rm = TRUE)) %>% 
  ungroup()

# -- Incidence rate
tbl_case_inc_30days <- df_countries %>% 
  filter(region %in% c('Eastern Europe', 'Western Asia', 'Central Asia')) %>% 
  left_join(tbl_case_count_30days) %>% 
  mutate(
    incdc = cases / pop * 100)

# -- Growth rate
tbl_growth_rate_cases_30d <- lapply(lst_mdl_cases_30d, function (x) x[['coeffs']]) %>% 
  bind_rows(.id = "iso_a3") %>% 
  transform(gr = exp(coeff) - 1) %>% 
  select(iso_a3, gr)

# -- Plot
tbl_30days <- tbl_case_inc_30days %>% 
  left_join(tbl_growth_rate_cases_30d)

ggplot(data = tbl_30days) + 
  geom_point(aes(x = gr, y = incdc, col = region)) + 
  ggrepel::geom_text_repel(aes(gr, incdc, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_x_continuous(labels = scales::label_percent(scale = 100)) + 
  labs(
    #caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    y = "Cumulative incidence rate in the last 30 days (log scale %)",
    x = "Estimated growth rate in cases during the last 30 days (%)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```



## Growth rate vs cumulative incidence rate in the last 14 days

```{r, fig.width = 6, fig.height = 6}

# Cumulative count cases
tbl_case_count_14days <- dta_ecdc %>% 
  filter(between(date, left = date_max - 13, right = date_max)) %>% 
  filter(region %in% c('Northern Europe', 'Southern Europe', 'Western Europe')) %>% 
  group_by(iso_a3, region, country) %>% 
  summarise(cases = sum(cases, na.rm = TRUE)) %>% 
  ungroup()

# -- Incidence rate
tbl_case_inc_14days <- df_countries %>% 
  filter(region %in% c('Northern Europe', 'Southern Europe', 'Western Europe')) %>% 
  left_join(tbl_case_count_14days) %>% 
  mutate(
    incdc = cases / pop * 100)

# -- Growth rates
tbl_growth_rate_cases_14d <- lapply(lst_mdl_cases_14d, function (x) x[['coeffs']]) %>% 
  bind_rows(.id = "iso_a3") %>% 
  transform(gr = exp(coeff) - 1) %>% 
  select(iso_a3, gr)


tbl_14days <- tbl_case_inc_14days %>% 
  left_join(tbl_growth_rate_cases_14d)

ggplot(data = tbl_14days) + 
  geom_point(aes(x = gr, y = incdc, col = region)) + 
  ggrepel::geom_text_repel(aes(gr, incdc, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_x_continuous(labels = scales::label_percent(scale = 100)) + 
  labs(
    #caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    y = "Cumulative incidence rate in the last 14 days (log scale %)",
    x = "Estimated growth rate in the last 14 days (%)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```


```{r, fig.width = 6, fig.height = 6}

# Cumulative count cases
tbl_case_count_14days <- dta_ecdc %>% 
  filter(between(date, left = date_max - 13, right = date_max)) %>% 
  filter(region %in% c('Eastern Europe', 'Western Asia', 'Central Asia')) %>% 
  group_by(iso_a3, region, country) %>% 
  summarise(cases = sum(cases, na.rm = TRUE)) %>% 
  ungroup()

# -- Incidence rate
tbl_case_inc_14days <- df_countries %>% 
  filter(region %in% c('Eastern Europe', 'Western Asia', 'Central Asia')) %>% 
  left_join(tbl_case_count_14days) %>% 
  mutate(
    incdc = cases / pop * 100)

# -- Growth rates
tbl_growth_rate_cases_14d <- lapply(lst_mdl_cases_14d, function (x) x[['coeffs']]) %>% 
  bind_rows(.id = "iso_a3") %>% 
  transform(gr = exp(coeff) - 1) %>% 
  select(iso_a3, gr)


tbl_14days <- tbl_case_inc_14days %>% 
  left_join(tbl_growth_rate_cases_14d)

ggplot(data = tbl_14days) + 
  geom_point(aes(x = gr, y = incdc, col = region)) + 
  ggrepel::geom_text_repel(aes(gr, incdc, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_x_continuous(labels = scales::label_percent(scale = 100)) + 
  labs(
    #caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    y = "Cumulative incidence rate in the last 14 days (log scale %)",
    x = "Estimated growth rate in the last 14 days (%)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```

<!-- 
4.	Excess of mortality per country?
a.	We could replicate with a nicer display this analysis: https://www.ft.com/content/a2901ce8-5eb7-4633-b89c-cbdf5b386938
b.	Data Source: https://github.com/Financial-Times/coronavirus-excess-mortality-data
-->

