---
title: "Description of the Covid-19 epidemic in `r params$continent`"
params:
  continent: Africa
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: Francesco Grandesso Epicentre (MSF)
output: 
  rmdformats::readthedown:
    use_bookdown: TRUE
editor_options: 
  chunk_output_type: console
---


```{r setup, include = FALSE}

if(Sys.getlocale(category = "LC_TIME") == "French_France.1252") {
  Sys.setlocale(category = "LC_TIME", locale = "English_United Kingdom.1252")
  Sys.setenv(LANG = "en_GB.UTF-8") 
}

knitr::opts_chunk$set(cache     = FALSE, 
                      echo      = FALSE, 
                      tidy      = TRUE, 
                      collapse  = TRUE, 
                      dpi       = 150, 
                      fig.align = "center", 
                      warning   = FALSE, 
                      message   = FALSE, 
                      encoding  = "UTF-8")


if (!exists('path.root')) {
  source(here::here('R', 'setup.r'), encoding = 'UTF-8')
}

source(file.path(path.R, "utils_management.R"), encoding = "UTF-8")
source(file.path(path.R, "utils_vis.R")       , encoding = "UTF-8")
source(file.path(path.R, "utils_modelling.R")       , encoding = "UTF-8")

# Packages
library(tidyverse)
library(treemap)
library(patchwork)
library(countrycode)
library(geofacet)
library(glue)
library(slider)


# Set dates
dates_and_week <- set_date_frame(create_folders = TRUE)

date_min_report <- dates_and_week[[1]]
date_min_report <- as.Date("2020-01-22")
date_max_report <- dates_and_week[[2]]
week_report     <- dates_and_week[[3]]

params$continent


# Set paths
set_paths(path.local, week_report)
path.local.deepdive <- file.path(path.local.week.deepdive,
                                 params$continent)
if(!exists(path.local.deepdive)) {
  dir.create(path.local.deepdive, showWarnings = FALSE, recursive = TRUE)
}
```



```{r upload-data, include = FALSE}

## --- ECDC data
dta_ecdc <- covidutils::get_ecdc_data()


## --- JHU Covid data
dta_jhu <- get_owid_jhcsse() %>% 
  tidyr::drop_na(iso_a3) %>% 
  filter(between(date, left = date_min_report, right = date_max_report),
         (continent == params$continent))


## --- Population data
df_countries <- readRDS(file.path(path.local.data, paste0('df_countries','.RDS'))) %>% 
  filter(continent == params$continent) %>% 
  left_join(readRDS(file.path(path.local.data, paste0('df_pop_country','.RDS'))))


# sf Robinson projection
sf_robinson <- readRDS(file.path(path.local.data, paste0('sf_world','.RDS'))) %>% 
  filter(continent == params$continent)


## --- GIS data - Mercator projection
sf_mercator <- rnaturalearth::ne_countries(type = "countries", returnclass = "sf") %>% 
  tibble::as_tibble() %>% 
  sf::st_as_sf() %>% 
  dplyr::select(country = name_long, iso_a3, iso_a2, pop_est) %>% 
  dplyr::mutate(
    continent = suppressWarnings(countrycode::countrycode(iso_a3, origin = "iso3c",
                                                          destination = "continent")),
    region = suppressWarnings(countrycode::countrycode(iso_a3, origin = "iso3c", 
                                                       destination = "region23"))) %>% 
  dplyr::filter(stringr::str_detect(country, "Antarctic", negate = TRUE)) %>% 
  cbind(sf::st_coordinates(suppressWarnings(sf::st_centroid(., of_largest_polygon = TRUE)))) %>% 
  dplyr::rename(lon = X, lat = Y) 

```



```{r prepare-data, include = FALSE}

# dta_ecdc <- dta_ecdc %>% 
#    prepare_ecdc_geodata_geofacet() %>% 
#   drop_na(iso_a3) %>% 
#   filter(between(date, left = NULL, 
#                  right = date_max_report))


# Define lims
xlim_continent <- switch(params$continent,
       "Africa"   = c(-25, 60),
       "Europe"   = c(-30, 80),
       "Americas" = c(-160, -30),
       "Asia"     = c(20, 170))

ylim_continent <- switch(params$continent,
       "Africa"   = c(-35, 40),
       "Europe"   = c(28, 73),
       "Americas" = c(-70, 90),
       "Asia"     = c(-20, 90))



## CASES Incidence proportion

tbl_cases_prop <- dta_jhu %>% 
  group_by(country) %>% 
  summarise(
    n = sum(cases)) %>% 
  left_join(df_countries) %>% 
  mutate(
   p = n/pop * 100)

tbl_cases_prop_30d <- dta_jhu %>% 
  filter(between(date, left = date_max_report - 29, right = date_max_report)) %>%
  group_by(country) %>% 
  summarise(
    n_30d = sum(cases)) %>% 
    left_join(df_countries) %>% 
  mutate(
   p_30d = n_30d/pop * 100) %>% 
  left_join(tbl_cases_prop) %>% 
  select(region, country, n, p, n_30d, p_30d)


tbl_cases_prop_2w <- dta_jhu %>% 
  filter(between(date, left = date_max_report - 13, right = date_max_report)) %>%
  group_by(country) %>% 
  summarise(
    n_2w = sum(cases)) %>% 
    left_join(df_countries) %>% 
  mutate(
   p_2w = n_2w/pop * 100) %>% 
  left_join(tbl_cases_prop) %>% 
  select(region, country, n, p, n_2w, p_2w)


## DEATHS Incidence proportion

tbl_deaths_prop <- dta_jhu %>% 
  group_by(country) %>% 
  summarise(
    n = sum(deaths)) %>% 
  left_join(df_countries) %>% 
  mutate(
   p = n/pop * 100)

tbl_deaths_prop_2w <- dta_jhu %>% 
  filter(between(date, left = date_max_report - 13, right = date_max_report)) %>%
  group_by(country) %>% 
  summarise(
    n_2w = sum(deaths)) %>% 
    left_join(df_countries) %>% 
  mutate(
   p_2w = n_2w/pop * 100) %>% 
  left_join(tbl_deaths_prop) %>% 
  select(region, country, n, p, n_2w, p_2w)

```



Data source: [JHU CSSE data](https://github.com/owid/covid-19-data/tree/master/public/data) updated until `r format(date_max_report, '%d %B %Y')`

**How to read the treemaps**

**Treemapping** is a method to display a set of tiled rectangles whose area is proportional to a specified dimension of the data. The rectangles are filled with a sequential colour scheme representing a second dimension of the data. When the size and colour dimensions are correlated, one can often easily see patterns that would be difficult to spot in other ways.

The **treemaps presented in this report** display the size of the rectangles proportional to the total number of cases (or deaths) in each country, and the colour of the rectangles representing a second variable whose definition and colour scheme is displayed in the legend. 

The relation of the the two variables are also presented in scatter plots.



# Overall

## Scatter plot all countries

```{r, fig.cap = paste("SCATTER PLOT: Total number and incidence proportion of confirmed cases per country")}

countries_zero_count <- tbl_cases_prop %>% 
  filter(n == 0) %>% 
  pull(country)

txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 
                                   'none', 
                                   combine_words(countries_zero_count))

ggplot(tbl_cases_prop %>% filter(n != 0)) + 
  geom_point(aes(n, p, col = region)) + 
  ggrepel::geom_text_repel(aes(n, p, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 5), 
                     labels = scales::label_number_si()) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    colour = 'Region', 
    caption = paste0("Countries with zero count: \n", txt_countries_zero_count), 
    x = "Total count of cases\n(log scale)",
    y = "Incidence proportion %\n(log scale)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```




# During the last 30 days

## Scatter plot total cases _versus_ cases in the last 30 days

```{r, fig.cap = paste("SCATTER PLOT: Number of cases and cases in the last 30 days")}

countries_zero_count <- tbl_cases_prop_30d %>% 
  filter(n == 0) %>% 
  pull(country)

txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))

scatter_plot_cases_cases_30d <- ggplot(tbl_cases_prop_30d %>% filter(n != 0)) + 
  geom_point(aes(n, n_30d, col = region)) + 
  ggrepel::geom_text_repel(aes(n, n_30d, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 5), 
                     labels = scales::label_number_si()) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 10), 
                     labels = scales::label_number_si()) + 
  labs(
    colour = 'Region', 
    caption = paste0("Countries with zero count: \n", txt_countries_zero_count), 
    x = "Total number of cases\n(log scale)",
    y = "Number of cases in the last 30 days\n(log scale)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")


scatter_plot_cases_cases_30d

ggsave(file.path(path.local.deepdive, paste0('scatter_plot_cases_cases_30d', '_',
                                                    week_report, '.png')), 
       plot = scatter_plot_cases_cases_30d, 
       scale = 1, 
       dpi = 320)
```


## Scatter plot cases _versus_ incidence proportion of cases in the last 30 days

```{r, fig.cap = paste("SCATTER PLOT: Number of cases and cases in the last 30 days")}

countries_zero_count <- tbl_cases_prop_30d %>% 
  filter(n == 0) %>% 
  pull(country)

txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))

scatter_plot_cases_prop_30d <- ggplot(tbl_cases_prop_30d %>% filter(n != 0)) + 
  geom_point(aes(n_30d, p_30d, col = region)) + 
  ggrepel::geom_text_repel(aes(n_30d, p_30d, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0), n = 5), 
                     labels = scales::label_number_si()) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 1), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    colour = 'Region', 
    caption = paste0("Countries with zero count: \n", txt_countries_zero_count), 
    x = "Number of cases in the last 30 days\n(log scale)",
    y = "Incidence proportion of cases in the last 30 days\n(log scale)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

scatter_plot_cases_prop_30d

ggsave(file.path(path.local.deepdive, 
                 paste0('scatter_plot_cases_prop_30d', '_', week_report, '.png')), 
       plot = scatter_plot_cases_prop_30d, 
       scale = 1, 
       dpi = 320)
```


## Scatter plot overall incidence proportion of cases _versus_ incidence proportion of cases in the last 30 days

```{r, fig.cap = paste("SCATTER PLOT: Number of cases and cases in the last 30 days")}

countries_zero_count <- tbl_cases_prop_30d %>% 
  filter(n == 0) %>% 
  pull(country)

txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))

scatter_plot_prop_prop_30d <- ggplot(tbl_cases_prop_30d %>% filter(n != 0)) + 
  geom_point(aes(p, p_30d, col = region)) + 
  ggrepel::geom_text_repel(aes(p, p_30d, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 1), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 1), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    colour = 'Region', 
    caption = paste0("Countries with zero count: \n", txt_countries_zero_count), 
    x = "Overall incidence proportion of cases\n(log scale)",
    y = "Incidence proportion of cases in the last 30 days\n(log scale)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

scatter_plot_prop_prop_30d

ggsave(file.path(path.local.deepdive, paste0('scatter_plot_prop_prop_30d', '_', week_report, '.png')), 
       plot = scatter_plot_prop_prop_30d, 
       scale = 1, 
       dpi = 320)
```



# During the last 14 days

## Overall incidence proportions _versus_ incidence proportions in the last 14 days

### Map

<!-- Map of overall incidence proportion -->
```{r maps-incidence-proportions, fig.cap = "MAP: Overall incidence proportion of cases and recent incidence proportions of cases (last 14 days)"}

## Breaks for incidence
breaks_inc <- c(0, 0.01, 0.1, 1, Inf)
labels_inc <- c("0 - 0.01%", "0.01 - 0.1%", "0.1 - 1%", "1%+")

## --- OVERALL ---

## Put incidence table in to sf
sf_case_inc <- df_countries %>% 
  left_join(tbl_cases_prop) %>% 
  mutate(
    inc_brks = cut(p, breaks_inc, labels = labels_inc, 
                   include.lowest = TRUE, right = FALSE)) %>% 
  full_join(
    select(sf_mercator, iso_a3),
    by = "iso_a3"
  ) %>% 
  st_as_sf()



# ggplot(sf_case_inc) +
#   geom_sf(aes(fill = inc_brks), size = .1) +
#    coord_sf(
#     xlim = c(20, 170),
#     ylim = c(-20, 90),
#     expand = FALSE,
#     datum = NA) +
#   labs(title = 'Overall')




## Map 
map_case_inc <- ggplot(sf_case_inc) + 
  geom_sf(aes(fill = inc_brks), size = .1) + 
  scale_fill_brewer(
    name = 'Incidence proportion', 
    palette = "Blues", 
    na.value = "lightgrey", 
    drop = FALSE, 
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(70 / length(labels_inc), units = "mm"),
      title.hjust = 0.5,
      nrow = 1,
      label.position = "bottom",
      title.position = 'top')) + 
  coord_sf(
    xlim = xlim_continent, 
    ylim = ylim_continent, 
    expand = FALSE,
    datum = NA) + 
  labs(title = 'Overall') + 
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        legend.position = "bottom")


## --- LAST TWO WEEKS ---

## Put incdence table (last 2 weks) in to sf
sf_case_inc_2weeks <- df_countries %>% 
  left_join(tbl_cases_prop_2w) %>% 
  mutate(
    inc_brks = cut(p_2w, breaks_inc, labels = labels_inc, include.lowest = TRUE, right = FALSE)) %>% 
  full_join(
    select(sf_mercator, iso_a3),
    by = "iso_a3"
  ) %>% 
  st_as_sf()

## Map
map_case_inc_2weeks <- ggplot(sf_case_inc_2weeks) + 
  geom_sf(aes(fill = inc_brks), size = .1) + 
  scale_fill_brewer(
    name = 'Incidence proportion', 
    palette = "Blues", 
    na.value = "lightgrey", 
    drop = FALSE, 
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(70 / length(labels_inc), units = "mm"),
      title.hjust = 0.5,
      nrow = 1,
      label.position = "bottom",
      title.position = 'top')) + 
  coord_sf(
    xlim = xlim_continent, 
    ylim = ylim_continent,
    expand = FALSE,
    datum = NA) + 
  labs(title = 'Last 14 days') + 
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        legend.position = "bottom")


## Combine maps
map_case_inc <- map_case_inc + 
  map_case_inc_2weeks + 
  plot_layout(ncol = 2, guides = 'collect') & theme(legend.position = 'bottom')

map_case_inc

## Save map
ggsave(file.path(path.local.deepdive, paste0('map_case_inc', '_', week_report, '.png')), 
       plot = map_case_inc, 
       scale = 1, 
       dpi = 320)

```


### Scatter plots

```{r, fig.cap = paste("SCATTER PLOT: Overall incidence proportion of cases and recent incidence proportions of cases in the last 14 days")}

tbl_case_inc_vs_inc_2weeks <- tbl_cases_prop_2w %>% 
  select(region, country, overall = p, last2weeks = p_2w)

countries_zero_count <- tbl_case_inc_vs_inc_2weeks %>% 
  filter(last2weeks == 0) %>% 
  pull(country)

txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))


# The dots plot
ggplot(tbl_case_inc_vs_inc_2weeks %>% filter(last2weeks != 0)) + 
  geom_point(aes(overall, last2weeks, col = region)) + 
  ggrepel::geom_text_repel(aes(overall, last2weeks, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 1), n = 5), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 2), n = 10), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    colour = 'Region', 
    caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    x = "Incidence proportion since the beginning of the epidemic %\n(log scale)",
    y = "Incidence proportion in the last 14 days %\n(log scale)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```


# Trends


```{r linear-models}
lst_dta_jhu <- dta_jhu %>% 
  multisplit("iso_a3")


# -- Linear model trends - last 30 days 

lst_mdl_cases_30d <- list()

for(i in names(lst_dta_jhu)){
  mdl <- linear_trend_deepdive(dta = lst_dta_jhu[[i]], 
                               series = 'cases', 
                               n_days = 30, 
                               last_date = date_max_report)
  lst_mdl_cases_30d[[i]] <- mdl
}


lst_mdl_deaths_30d <- list()

for(i in names(lst_dta_jhu)){
  mdl <- linear_trend_deepdive(dta = lst_dta_jhu[[i]], 
                               series = 'deaths', 
                               n_days = 30, 
                               last_date = date_max_report)
  lst_mdl_deaths_30d[[i]] <- mdl
}



# -- Linear model trends - last 14 days 

lst_mdl_cases_14d <- list()

for(i in names(lst_dta_jhu)){
  mdl <- linear_trend_deepdive(dta = lst_dta_jhu[[i]], 
                               series = 'cases', 
                               n_days = 12, 
                               last_date = date_max_report)
  lst_mdl_cases_14d[[i]] <- mdl
}


lst_mdl_deaths_14d <- list()

for(i in names(lst_dta_jhu)){
  mdl <- linear_trend_deepdive(dta = lst_dta_jhu[[i]], 
                               series = 'deaths', 
                               n_days = 12, 
                               last_date = date_max_report)
  lst_mdl_deaths_14d[[i]] <- mdl
}

```


```{r tbl-coeffs}
# -- Lists of predictions
lst_preds_cases_30d <- lapply(lst_mdl_cases_30d, function (x) x[['preds']])
lst_preds_cases_14d <- lapply(lst_mdl_cases_14d, function (x) x[['preds']])

lst_preds_deaths_30d <- lapply(lst_mdl_deaths_30d, function (x) x[['preds']])
lst_preds_deaths_14d <- lapply(lst_mdl_deaths_14d, function (x) x[['preds']])


tbl_coeff_cases_30d <- lapply(lst_mdl_cases_30d, function (x) x[['coeffs']]) %>% 
  tibble::enframe(name = "iso_a3") %>% 
  unnest(cols = c(value)) %>% 
  mutate(
    trend = case_when(
      lwr > 0 ~ "Increasing", 
      upr < 0 ~ "Declining", 
      upr > 0 & lwr < 0 ~ "Stable", 
      TRUE ~ NA_character_) %>% 
      factor(levels = c('Increasing', 'Stable', 'Declining')))


tbl_coeff_cases_14d <- lapply(lst_mdl_cases_14d, function (x) x[['coeffs']]) %>% 
  tibble::enframe(name = "iso_a3") %>% 
  unnest(cols = c(value)) %>% 
  mutate(
    trend = case_when(
      lwr > 0 ~ "Increasing", 
      upr < 0 ~ "Declining", 
      upr > 0 & lwr < 0 ~ "Stable", 
      TRUE ~ NA_character_) %>% 
      factor(levels = c('Increasing', 'Stable', 'Declining')))

tbl_coeff_deaths_30d <- lapply(lst_mdl_deaths_30d, function (x) x[['coeffs']]) %>% 
  tibble::enframe(name = "iso_a3") %>% 
  unnest(cols = c(value))

tbl_coeff_deaths_14d <- lapply(lst_mdl_deaths_14d, function (x) x[['coeffs']]) %>% 
  tibble::enframe(name = "iso_a3") %>% 
  unnest(cols = c(value))

```


## Map

```{r maps-trends}

sf_case_trend_30d <- df_countries %>% 
  select(iso_a3, country) %>% 
  left_join(tbl_coeff_cases_30d) %>% 
  full_join(
    select(sf_mercator, iso_a3),
    by = "iso_a3") %>% 
  st_as_sf()


sf_case_trend_14d <- df_countries %>% 
  select(iso_a3, country) %>% 
  left_join(tbl_coeff_cases_14d) %>% 
  full_join(
    select(sf_mercator, iso_a3),
    by = "iso_a3") %>% 
  st_as_sf()


RdAmGn <- c('#D95F02', '#E6AB02', '#1B9E77') 

map_case_trend_30d <- ggplot(sf_case_trend_30d) + 
  geom_sf(aes(fill = trend), size = .1, alpha = 0.8) + 
  scale_fill_manual(
    name = NULL, 
    values = RdAmGn, 
    drop = FALSE, 
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(70 / length(labels_inc), units = "mm"),
      title.hjust = 0.5,
      nrow = 1,
      label.position = "bottom",
      title.position = 'top')) + 
  coord_sf(
    xlim = xlim_continent, 
    ylim = ylim_continent, 
    expand = FALSE,
    datum = NA) + 
  labs(title = 'Last 30 days') + 
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        legend.position = "bottom")


map_case_trend_14d <- ggplot(sf_case_trend_14d) + 
  geom_sf(aes(fill = trend), size = .1, alpha = 0.8) + 
  scale_fill_manual(
    name = NULL, 
    values = RdAmGn, 
    drop = FALSE, 
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(70 / length(labels_inc), units = "mm"),
      title.hjust = 0.5,
      nrow = 1,
      label.position = "bottom",
      title.position = 'top')) + 
  coord_sf(
    xlim = xlim_continent, 
    ylim = ylim_continent, 
    expand = FALSE, 
    datum = NA) + 
  labs(title = 'Last 12 days') + 
  theme_light() +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(), 
        legend.position = "bottom",
        plot.margin = margin(0, 0, 0, 0, "pt"))



## Combine maps
map_cases_trends <- map_case_trend_30d + 
  map_case_trend_14d + 
  plot_layout(ncol = 2, guides = 'collect') & theme(legend.position = 'bottom')

map_cases_trends

ggsave(file.path(path.local.deepdive, paste0('map_cases_trends', '_', 
                                                    week_report, '.png')), 
       plot = map_cases_trends, 
       scale = 1, 
       dpi = 320)
```



## Epidemic curves

```{r fig-epicurves}
iso_a3_increasing <- tbl_coeff_cases_30d %>% 
  filter(trend == 'Increasing') %>% 
  pull(iso_a3)

dta_jhu_increasing <- dta_jhu %>% 
  filter(iso_a3 %in% iso_a3_increasing)


epicurves <- ggplot(dta_jhu_increasing, aes(x = date, y = cases)) + 
    facet_wrap(~country, scales = "free_y", ncol = 4) + 
    geom_col(fill = '#1A62A3') + 
    scale_x_date(breaks = '2 months', date_labels = "%b-%Y") +
    xlab('') + 
    ylab('frequency') + 
    theme_light() + 
    theme(legend.position = "none", 
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          strip.text = element_text(size = 11))

epicurves

ggsave(file.path(path.local.deepdive, paste0('epicurves', '_', week_report, '.png')), 
       plot = epicurves, 
       height = 9,
       width = 12,
       scale = 0.8)
```


```{r fig-preds}

tbl_preds_cases_30d <- lapply(lst_mdl_cases_30d, function (x) x[['preds']]) %>% 
  tibble::enframe(name = "iso_a3") %>% 
  unnest(cols = c(value)) %>% 
  left_join(df_countries) %>% 
  filter(iso_a3 %in% iso_a3_increasing)

preds <- ggplot(tbl_preds_cases_30d, aes(x = date, y = cnt)) + 
    facet_wrap(~ country, scales = "free_y", ncol = 4) + 
    geom_point(colour = '#1A62A3', size = 2) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr), fill = '#1A62A3', alpha = 0.4) + 
    geom_line(aes(y = fit), colour = '#1A62A3', size = 1) + 
    scale_x_date(date_labels = "%d-%b") +
    xlab('') + 
    ylab(paste0('frequency and fitted values')) + 
    theme_light() + 
    theme(legend.position = "none", 
          axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
          strip.text = element_text(size = 11))

preds

ggsave(file.path(path.local.deepdive, paste0('preds', '_', 
                                                    week_report, '.png')), 
       plot = preds, 
       height = 9,
       width = 12,
       scale = 0.8)

```
