---
title: "Descritption of the Covid-19 epidemic in the Americas"
date: '`r Sys.Date()`'
author:
  - name: Francesco Grandesso
    affiliation: Epicentre (MSF)
output: 
  distill::distill_article:
    toc: TRUE
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

<!--
---
title: "EpiSitrep South-America"
author: "Francesco Grandesso"
date: "23/09/2020"
output: html_document
editor_options: 
chunk_output_type: console
---
-->


```{r setup, include = FALSE}
knitr::opts_chunk$set(cache = FALSE, 
                      echo = FALSE, 
                      tidy = TRUE, 
                      collapse = TRUE, 
                      dpi = 150, 
                      fig.align = "center", 
                      warning = FALSE, 
                      message = FALSE, 
                      encoding = "UTF-8")

source(here::here('R', 'setup.R'), encoding = 'UTF-8')
source(file.path(path.R, "utils_management.R"), encoding = "UTF-8")
source(file.path(path.R, "utils_vis.R")       , encoding = "UTF-8")

library(treemap)


# Set dates
dates_and_week <- set_date_frame(create_folders = TRUE)

date_min_report <- dates_and_week[[1]]
date_max_report <- dates_and_week[[2]]
week_report     <- dates_and_week[[3]]


# Set paths
set_paths(path.local, week_report)
path.local.deepdive.Americas <- file.path(path.local.worldwide,
                                          'deepdive_Americas')
if(!exists(path.local.deepdive.Americas)) {
  dir.create(path.local.deepdive.Americas, showWarnings = FALSE, recursive = TRUE)
}
```


```{r upload-data, include = FALSE}

## --- ECDC data
dta_ecdc <- covidutils::get_ecdc_data()


## --- JHU Covid data
dta_jhu <- get_owid_jhcsse() %>% 
  tidyr::drop_na(iso_a3) %>% 
  filter(between(date, left = NULL, right = date_max_report) & (continent == 'Americas'))



# List of Americas countries with population
df_countries <- readRDS(file.path(path.local.data, paste0('df_countries','.RDS'))) %>% 
  filter(continent == 'Americas') %>% 
  left_join(readRDS(file.path(path.local.data, paste0('df_pop_country','.RDS'))))


# sf Robinson projection
sf_robinson_Americas <- readRDS(file.path(path.local.data, paste0('sf_world','.RDS'))) %>%
  filter(continent == 'Americas')


## Mercator projection
sf_mercator_Americas <- rnaturalearth::ne_countries(type = "countries", returnclass = "sf") %>% 
  tibble::as_tibble() %>% 
  sf::st_as_sf() %>% 
  dplyr::select(country = name_long, iso_a3, iso_a2, pop_est) %>% 
  dplyr::mutate(
    continent = suppressWarnings(countrycode::countrycode(iso_a3, origin = "iso3c", destination = "continent")),
    region = suppressWarnings(countrycode::countrycode(iso_a3, origin = "iso3c", destination = "region23"))
  ) %>% 
  dplyr::filter(stringr::str_detect(country, "Antarctic", negate = TRUE)) %>% 
  cbind(sf::st_coordinates(suppressWarnings(sf::st_centroid(., of_largest_polygon = TRUE)))) %>% 
  dplyr::rename(lon = X, lat = Y) %>% 
  filter(continent == 'Americas')
```



```{r prepare-data, include = FALSE}

dta_ecdc <- dta_ecdc %>% 
  prepare_ecdc_data_geofacet() %>% 
  drop_na(iso_a3) %>% 
  filter(between(date, left = NULL, right = date_max_report))


## CASES Incidence proportion
tbl_cases_prop <- dta_jhu %>% 
  group_by(country) %>% 
  summarise(
    n = sum(cases)) %>% 
  left_join(df_countries) %>% 
  mutate(
    p = n / pop * 100)

tbl_cases_prop_30d <- dta_jhu %>% 
  filter(between(date, left = date_max_report - 29, right = date_max_report)) %>%
  group_by(country) %>% 
  summarise(
    n_30d = sum(cases)) %>% 
  left_join(df_countries) %>% 
  mutate(
    p_30d = n_30d/pop * 100) %>% 
  left_join(tbl_cases_prop) %>% 
  select(region, country, n, p, n_30d, p_30d)


tbl_cases_prop_2w <- dta_jhu %>% 
  filter(between(date, left = date_max_report - 13, right = date_max_report)) %>%
  group_by(country) %>% 
  summarise(
    n_2w = sum(cases)) %>% 
  left_join(df_countries) %>% 
  mutate(
    p_2w = n_2w/pop * 100) %>% 
  left_join(tbl_cases_prop) %>% 
  select(region, country, n, p, n_2w, p_2w)


## DEATHS Incidence proportion

tbl_deaths_prop <- dta_jhu %>% 
  group_by(country) %>% 
  summarise(
    n = sum(deaths)) %>% 
  left_join(df_countries) %>% 
  mutate(
    p = n/pop * 100)

tbl_deaths_prop_2w <- dta_jhu %>% 
  filter(between(date, left = date_max_report - 13, right = date_max_report)) %>%
  group_by(country) %>% 
  summarise(
    n_2w = sum(deaths)) %>% 
  left_join(df_countries) %>% 
  mutate(
    p_2w = n_2w/pop * 100) %>% 
  left_join(tbl_deaths_prop) %>% 
  select(region, country, n, p, n_2w, p_2w)

```


Data source: [ECDC data](https://opendata.ecdc.europa.eu/covid19/casedistribution/csv) updated until `r format(date_max_report, '%d %B %Y')`


# 1 - Overview of the burden of cases and deaths

## 1.1 - Cumulative number of confirmed cases per country 

```{r, fig.width = 5, fig.height = 3}

selected <- c('Northern America', 'Central America')

dta_jhu %>% 
  filter(region %in% selected) %>% 
  treemap(index = "country", vSize = "cases", type = "index" , title = paste0('Cumulative cases - ', combine_words(selected)))

```

```{r, fig.width = 5, fig.height = 3}

selected <- c('South America')

dta_jhu %>% 
  filter(region %in% selected) %>% 
  treemap(index = "country", vSize = "cases", type = "index" , title = paste0('Cumulative cases - ', combine_words(selected)))

```


## 1.2 - Cumulative number of associeted deaths per country  

```{r, fig.width = 5, fig.height = 3}

selected <- c('Northern America', 'Central America')

dta_jhu %>% 
  filter(region %in% selected) %>% 
  treemap(index = "country", vSize = "deaths", type = "index" , title = paste0('Cumulative deaths - ', combine_words(selected)))

```

```{r, fig.width = 5, fig.height = 3}

selected <- c('South America')

dta_jhu %>% 
  filter(region %in% selected) %>% 
  treemap(index = "country", vSize = "deaths", type = "index" , title = paste0('Cumulative deaths - ', combine_words(selected)))

```


## 1.3 - Number of confirmed cases in the last 14 days

```{r, fig.width = 5, fig.height = 3}

selected <- c('Northern America', 'Central America')

dta_jhu %>% 
  filter(between(date, left = date_max_report-13, right = date_max_report)) %>% 
  filter(region %in% selected) %>% 
  treemap(index = "country", vSize = "cases", type = "index" , title = paste0('Cumulative cases in the last 14 days - ', combine_words(selected)))

```

```{r, fig.width = 5, fig.height = 3}

selected <- c('South America')

dta_jhu %>% 
  filter(between(date, left = date_max_report-13, right = date_max_report)) %>% 
  filter(region %in% selected) %>% 
  treemap(index = "country", vSize = "cases", type = "index" , title = paste0('Cumulative cases in the last 14 days - ', combine_words(selected)))

```


## 1.4 - Number of associated deaths in the last 14 days

```{r, fig.width = 5, fig.height = 3}

selected <- c('Northern America', 'Central America')

dta_jhu %>% 
  filter(between(date, left = date_max_report-13, right = date_max_report)) %>% 
  filter(region %in% selected) %>% 
  treemap(index = "country", vSize = "deaths", type = "index" , title = paste0('Cumulative deaths in the last 14 days - ', combine_words(selected)))

```

```{r, fig.width = 5, fig.height = 3}

selected <- c('South America')

dta_jhu %>% 
  filter(between(date, left = date_max_report-13, right = date_max_report)) %>% 
  filter(region %in% selected) %>% 
  treemap(index = "country", vSize = "deaths", type = "index" , title = paste0('Cumulative deaths in the last 14 days - ', combine_words(selected)))

```



# 2 - Overall incidence of confirmed cases versus recent (last 14 days)

## 2.1 - Map of overall cumulative incidence + map of recent incidence

<!-- Map of cumulative incidences -->
```{r tbl-cumulative-incidences}

# Cumulative count cases
tbl_case_count <- dta_jhu %>% 
  group_by(iso_a3, region, country) %>% 
  summarise(cases = sum(cases, na.rm = TRUE)) %>% 
  ungroup()

# Breaks for incidence
breaks_inc <- c(0, 0.01, 0.1, 1, Inf)
labels_inc <- c("0 - 0.01%", "0.01 - 0.1%", "0.1 - 1%", "1%+")


# Cumulative incidence cases
tbl_case_inc <- df_countries %>% 
  left_join(tbl_case_count) %>% 
  mutate(
    inc = cases / pop * 100, 
    inc_brks = cut(inc, breaks_inc, labels = labels_inc, include.lowest = TRUE, right = FALSE))


sf_case_inc <- tbl_case_inc %>% 
  inner_join(
    select(sf_mercator_Americas, iso_a3),
    by = "iso_a3"
  ) %>% 
  st_as_sf()

map_case_inc <- ggplot(sf_case_inc) + 
  geom_sf(aes(fill = inc_brks), size = .1) + 
  scale_fill_brewer(
    name = 'Cumulative incidence rate', 
    palette = "Blues", 
    drop = FALSE, 
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(70 / length(labels_inc), units = "mm"),
      title.hjust = 0.5,
      nrow = 1,
      label.position = "bottom",
      title.position = 'top')) + 
  labs(title = 'Overall') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        legend.position = "bottom")

```

```{r tbl-cumulative-incidences-2weeks}

# Cumulative count cases
tbl_case_count_2weeks <- dta_jhu %>% 
  filter(between(date, left = date_max_report-13, right = date_max_report)) %>% 
  group_by(iso_a3, region, country) %>% 
  summarise(cases = sum(cases, na.rm = TRUE)) %>% 
  ungroup()

# Breaks for incidence
breaks_inc <- c(0, 0.01, 0.1, 1, Inf)
labels_inc <- c("0 - 0.01%", "0.01 - 0.1%", "0.1 - 1%", "1%+")


# Cumulative incidence cases
tbl_case_inc_2weeks <- df_countries %>% 
  left_join(tbl_case_count_2weeks) %>% 
  mutate(
    inc = cases / pop * 100, 
    inc_brks = cut(inc, breaks_inc, labels = labels_inc, include.lowest = TRUE, right = FALSE))


sf_case_inc_2weeks <- tbl_case_inc_2weeks %>% 
  inner_join(
    select(sf_mercator_Americas, iso_a3),
    by = "iso_a3"
  ) %>% 
  st_as_sf()

map_case_inc_2weeks <- ggplot(sf_case_inc_2weeks) + 
  geom_sf(aes(fill = inc_brks), size = .1) + 
  scale_fill_brewer(
    name = 'Cumulative incidence rate', 
    palette = "Blues", 
    drop = FALSE, 
    guide = guide_legend(
      keyheight = unit(3, units = "mm"),
      keywidth = unit(70 / length(labels_inc), units = "mm"),
      title.hjust = 0.5,
      nrow = 1,
      label.position = "bottom",
      title.position = 'top')) + 
  labs(title = 'Last 14 days') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5), 
        legend.position = "bottom")

```

```{r}

library(patchwork)
map_case_inc + 
  map_case_inc_2weeks + 
  plot_layout(guides='collect') & theme(legend.position = 'bottom')

```


##	2.2 - Cumulative incidence vs recent incidence

```{r, fig.widht = 6, fig.height = 6}

tbl_case_inc_vs_inc_2weeks <- tbl_case_inc %>% 
  select(region, country, overall = inc) %>% 
  left_join(tbl_case_inc_2weeks %>% 
              select(region, country, last2weeks = inc))

countries_zero_count <- tbl_case_inc_vs_inc_2weeks %>% filter(last2weeks == 0) %>% pull(country)
txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))


# The dots plot
dotsgplot <- ggplot(tbl_case_inc_vs_inc_2weeks %>% filter(last2weeks != 0)) + 
  geom_point(aes(overall, last2weeks, col = region)) + 
  ggrepel::geom_text_repel(aes(overall, last2weeks, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    x = "Overall cumulative incidence rate (log scale %)",
    y = "Cumulative incidence rate in the last 14 days (log scale %)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

dotsgplot

```


## 2.3 - Cumulative incidence vs total number of cases reported

```{r, fig.widht = 6, fig.height = 6}

tbl_case_count_vs_case_inc <- tbl_case_count %>% 
  select(region, country, cases) %>% 
  left_join(tbl_case_inc %>% 
              select(region, country, inc))

countries_zero_count <- tbl_case_count_vs_case_inc %>% filter(cases == 0) %>% pull(country)
txt_countries_zero_count <- ifelse(length(countries_zero_count) == 0, 'none', combine_words(countries_zero_count))


# The dots plot
dotsgplot <- ggplot(tbl_case_count_vs_case_inc %>% filter(cases != 0)) + 
  geom_point(aes(cases, inc, col = region)) + 
  ggrepel::geom_text_repel(aes(cases, inc, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_x_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_number_si()) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_percent(scale = 1)) + 
  labs(
    caption = paste0("Countries with zero count: \n", txt_countries_zero_count), 
    x = "Overall cumulative count of cases (log scale)",
    y = "Overall cumulative incidence rate (log scale %)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

dotsgplot

```





# 3 Trends

```{r}
# Get trends

linear_trend <- function(dta, series, last_date, n_days = 12, ma_window = 3, min_sum = 30){
  
  dta <- dta %>% select(date, cnt = all_of(series))
  
  dates_extent <- c(last_date - (n_days - 1), last_date)
  
  dta <- dta %>% 
    filter(between(date, dates_extent[1], dates_extent[2])) %>% 
    tidyr::complete(date = seq.Date(min(date, na.rm = TRUE), 
                                    max(date, na.rm = TRUE), by = 1), 
                    fill = list(cnt = NA_real_))
  
  
  # Moving average
  dta$ma <- forecast::ma(dta$cnt, order = ma_window)
  dta$ma <- na_if(dta$ma, 0) # Replace 0 values as NA
  
  
  # Empty matrix of predictions
  m_preds <- matrix(data = NA, 
                    nrow = dim(dta)[1], 
                    ncol = 3, 
                    dimnames = list(c(1:dim(dta)[1]), c('fit', 'lwr', 'upr')))
  
  
  # Run model with conditions
  if (dim(dta)[1] > ma_window & sum(dta$cnt, na.rm = TRUE) > min_sum) {
    
    mdl <- lm(log(ma) ~ date, data = dta)
    
    preds <- exp(predict(mdl, interval = 'confidence'))
    
    matched_rows <- match(rownames(preds), rownames(m_preds))
    matched_cols <- match(colnames(preds), colnames(m_preds))
    m_preds[matched_rows, matched_cols] <- preds
    
    tbl_preds <- tibble(date = seq.Date(from = dates_extent[1], to = dates_extent[2], by = 1), 
                        cnt  = dta$cnt, 
                        ma   = dta$ma, 
                        fit  = as.double(m_preds[, 'fit']), 
                        lwr  = as.double(m_preds[, 'lwr']), 
                        upr  = as.double(m_preds[, 'upr']))
    
    mdl_coeffs <- tibble(coeff = coefficients(mdl)[[2]], 
                         lwr   = confint(mdl)[2,1], 
                         upr   = confint(mdl)[2,2])
    
  } else {
    mdl <- NA_character_
    
    tbl_preds <- tibble(date = seq.Date(from = dates_extent[1], to = dates_extent[2], by = 1), 
                        cnt  = dta$cnt, 
                        ma   = dta$ma, 
                        fit  = as.double(m_preds[, 'fit']), 
                        lwr  = as.double(m_preds[, 'lwr']), 
                        upr  = as.double(m_preds[, 'upr']))
    
    mdl_coeffs <- tibble(coeff = NA_real_, 
                         lwr   = NA_real_, 
                         upr   = NA_real_)
  }
  
  return(list(mdl = mdl, preds = tbl_preds, coeffs = mdl_coeffs))
}


lst_dta_jhu <- dta_jhu %>% 
  multisplit("iso_a3")


# -- Linear model trends - last 14 days 

lst_mdl_cases_14d <- list()

for(i in names(lst_dta_jhu)){
  mdl <- linear_trend(dta = lst_dta_jhu[[i]], series = 'cases', n_days = 14, last_date = date_max_report)
  lst_mdl_cases_14d[[i]] <- mdl
}


lst_mdl_deaths_14d <- list()

for(i in names(lst_dta_jhu)){
  mdl <- linear_trend(dta = lst_dta_jhu[[i]], series = 'deaths', n_days = 14, last_date = date_max_report)
  lst_mdl_deaths_14d[[i]] <- mdl
}


# -- Linear model trends - last 30 days 

lst_mdl_cases_30d <- list()

for(i in names(lst_dta_jhu)){
  mdl <- linear_trend(dta = lst_dta_jhu[[i]], series = 'cases', n_days = 30, last_date = date_max_report)
  lst_mdl_cases_30d[[i]] <- mdl
}


lst_mdl_deaths_30d <- list()

for(i in names(lst_dta_jhu)){
  mdl <- linear_trend(dta = lst_dta_jhu[[i]], series = 'deaths', n_days = 30, last_date = date_max_report)
  lst_mdl_deaths_30d[[i]] <- mdl
}

```

```{r, eval = FALSE}
# To plot both cases and deaths into a single graphic plot
country_six_plots2 <- function(dta_all, dta_long, dta_short, date_max) {
  
  # Parameters
  main_colour  <- c(cases = '#1A62A3', deaths = '#e10000')
  date_min     <- dta_all %>% filter(cnt != 0) %>% pull(date) %>% min()
  
  
  # Plots
  plot_all <- ggplot(dta_all, aes(x = date, y = cnt)) + 
    facet_wrap(~obs, scales = "free_y", ncol = 1) + 
    geom_col(aes(colour = obs, fill = obs)) + 
    scale_colour_manual(values = main_colour) + 
    scale_fill_manual(values = main_colour) + 
    scale_x_date(limits = c(date_min, NA), breaks = '2 months', date_labels = "%b-%Y") +
    xlab('') + 
    ylab('frequency') + 
    labs(subtitle = 'Since the first cases reported') + 
    theme_light() + 
    theme(legend.position = "none", 
          strip.text = element_text(size = 11))
  
  plot_long <- ggplot(dta_long, aes(x = date, y = cnt)) + 
    facet_wrap(~ obs, scales = "free_y", ncol = 1) + 
    geom_point(aes(colour = obs), size = 2) + 
    scale_colour_manual(values = main_colour) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr, fill = obs), alpha = 0.4) + 
    geom_line(aes(y = fit, colour = obs), size = 1) + 
    scale_fill_manual(values = main_colour) + 
    scale_x_date(limits = c(min(dta_long$date), max(dta_long$date)), date_labels = "%d-%b") +
    xlab('') + 
    ylab(paste0('frequency and fitted values')) + 
    labs(subtitle = paste('Last', (max(dta_long$date) - min(dta_long$date) + 1), 'days')) + 
    theme_light() + 
    theme_light() + 
    theme(legend.position = "none", 
          strip.text = element_text(size = 11))
  
  plot_short <- ggplot(dta_short, aes(x = date, y = cnt)) + 
    facet_wrap(~ obs, scales = "free_y", ncol = 1) + 
    geom_point(aes(colour = obs), size = 2) + 
    scale_colour_manual(values = main_colour) + 
    geom_ribbon(aes(ymin = lwr, ymax = upr, fill = obs), alpha = 0.4) + 
    geom_line(aes(y = fit, colour = obs), size = 1) + 
    scale_fill_manual(values = main_colour) + 
    scale_x_date(limits = c(min(dta_short$date), max(dta_short$date)), breaks = '4 days', date_labels = "%d-%b") +
    xlab('') + 
    ylab(paste0('frequency and fitted values')) + 
    labs(subtitle = paste('Last', (max(dta_short$date) - min(dta_short$date) + 1), 'days')) + 
    theme_light() + 
    theme_light() + 
    theme(legend.position = "none", 
          strip.text = element_text(size = 11))
  
  
  # Arrange plots
  multiplot <- ggarrange(plot_all, 
                         plot_long, 
                         plot_short, 
                         ncol = 3, 
                         widths = c(2, 1.4, 1.1))
  
  return(multiplot)
}


# -- Lists of predictions
lst_preds_cases_30d <- lapply(lst_mdl_cases_30d, function (x) x[['preds']])
lst_preds_cases_14d <- lapply(lst_mdl_cases_14d, function (x) x[['preds']])

lst_preds_deaths_30d <- lapply(lst_mdl_deaths_30d, function (x) x[['preds']])
lst_preds_deaths_14d <- lapply(lst_mdl_deaths_14d, function (x) x[['preds']])


country_list <- df_countries

for(i in names(lst_dta_jhu)){
  
  #i = 'USA'
  name_country <- country_list %>% filter(iso_a3 == i) %>% pull(country) %>% gsub(" ", "_", .)
  
  dta_all <- lst_dta_jhu[[i]] %>% 
    select(date, cases, deaths) %>% 
    pivot_longer(-date, names_to = 'obs', values_to = 'cnt')
  
  dta_preds_long <- lst_preds_cases_30d[[i]] %>% 
    mutate(obs = 'cases') %>% 
    add_row(lst_preds_deaths_30d[[i]] %>% 
              mutate(obs = 'deaths')) %>% 
    arrange(date, obs)
  
  dta_preds_short <- lst_preds_cases_14d[[i]] %>% 
    mutate(obs = 'cases') %>% 
    add_row(lst_preds_deaths_14d[[i]] %>% 
              mutate(obs = 'deaths')) %>% 
    arrange(date, obs)
  
  country_multiplot <- country_six_plots2(dta_all, dta_preds_long, dta_preds_short)
  
  ggsave(filename = file.path(path.local.South_America, paste0('trends_', name_country, '.png')),
         plot = country_multiplot, 
         scale = 1, 
         width = 9,
         dpi = 320)
}

```


## 3.1 - Growth rate vs cumulative incidence rate in the last 30 days

```{r, fig.widht = 6, fig.height = 6}

# Cumulative case count
tbl_case_count_30days <- dta_jhu %>% 
  filter(between(date, left = date_max_report-29, right = date_max_report)) %>% 
  group_by(iso_a3, region, country) %>% 
  summarise(cases = sum(cases, na.rm = TRUE)) %>% 
  ungroup()

# -- Incidence rate
tbl_case_inc_30days <- df_countries %>% 
  left_join(tbl_case_count_30days) %>% 
  mutate(
    incdc = cases / pop * 100)

# -- Growth rate
tbl_growth_rate_cases_30d <- lapply(lst_mdl_cases_30d, function (x) x[['coeffs']]) %>% 
  bind_rows(.id = "iso_a3") %>% 
  transform(gr = exp(coeff) - 1) %>% 
  select(iso_a3, gr)

# -- Plot
tbl_30days <- tbl_case_inc_30days %>% 
  left_join(tbl_growth_rate_cases_30d)

ggplot(data = tbl_30days) + 
  geom_point(aes(x = gr, y = incdc, col = region)) + 
  ggrepel::geom_text_repel(aes(gr, incdc, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_x_continuous(labels = scales::label_percent(scale = 100)) + 
  labs(
    #caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    y = "Cumulative incidence rate in the last 30 days (log scale %)",
    x = "Estimated growth rate in cases during the last 30 days (%)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```


## 3.2 - Growth rate vs cumulative incidence rate in the last 14 days

```{r, fig.widht = 6, fig.height = 6}

# Cumulative count cases
tbl_case_count_14days <- dta_jhu %>% 
  filter(between(date, left = date_max_report-13, right = date_max_report)) %>% 
  group_by(iso_a3, region, country) %>% 
  summarise(cases = sum(cases, na.rm = TRUE)) %>% 
  ungroup()

# -- Incidence rate
tbl_case_inc_14days <- df_countries %>% 
  left_join(tbl_case_count_14days) %>% 
  mutate(
    incdc = cases / pop * 100)

# -- Growth rates
tbl_growth_rate_cases_14d <- lapply(lst_mdl_cases_14d, function (x) x[['coeffs']]) %>% 
  bind_rows(.id = "iso_a3") %>% 
  transform(gr = exp(coeff) - 1) %>% 
  select(iso_a3, gr)


tbl_14days <- tbl_case_inc_14days %>% 
  left_join(tbl_growth_rate_cases_14d)

ggplot(data = tbl_14days) + 
  geom_point(aes(x = gr, y = incdc, col = region)) + 
  ggrepel::geom_text_repel(aes(gr, incdc, label = country, colour = region), size = 3, show.legend = FALSE) + 
  scale_y_continuous(trans = 'log10', 
                     breaks = scales::trans_breaks("log10", function(x) round(10^x, 0)), 
                     labels = scales::label_percent(scale = 1)) + 
  scale_x_continuous(labels = scales::label_percent(scale = 100)) + 
  labs(
    #caption = paste0("Countries with zero count in the last 14 days: \n", txt_countries_zero_count), 
    y = "Cumulative incidence rate in the last 14 days (log scale %)",
    x = "Estimated growth rate in the last 14 days (%)") + 
  theme_light() + 
  theme(
    plot.caption = element_text(hjust = 0, face = "italic"), 
    plot.caption.position =  "panel")

```


4.	Excess of mortality per country?
a.	We could replicate with a nicer display this analysis: https://www.ft.com/content/a2901ce8-5eb7-4633-b89c-cbdf5b386938
b.	Data Source: https://github.com/Financial-Times/coronavirus-excess-mortality-data
