---
title: "MSF Covid-19 EpiSitrep MSF data analysis"
description: |
  This document presents the analyses on the MSF Covid-19 linelists. It is subject to change according to needs and requests. The R scripts for both worldwide and linelist MSF analysis are provided on the following Github repository https://github.com/epicentre-msf/covid19-episitrep-overall for epidemiologists in MSF to use (R skills required). The scripts will be updated from week to week as analysis progresses as requested.
author:
  - name: Francesco Grandesso
    url: mailto:francesco.GRANDESSO@epicentre.msf.org
    affiliation: Epicentre
    affiliation_url: https://epicentre.msf.org/
  - name: Paul Campbell
    url: mailto:paul.CAMPBELLepicentre.msf.org
    affiliation: Epicentre
    affiliation_url: https://epicentre.msf.org/
date: '`r Sys.Date()`'
output: 
  distill::distill_article:
    toc: TRUE
    toc_depth: 1
---


<!-- Setup and uploading functions -->
```{r setup, warning = FALSE, message = FALSE}
if (!exists('path.root')) {
  source(here::here('R', 'setup.r'), encoding = 'UTF-8')
}

source(file.path(path.R, "utils_management.R"), encoding = "UTF-8")
source(file.path(path.R, "utils_vis.R")       , encoding = "UTF-8")
source(file.path(path.R, "set_time_frame.R")  , encoding = "UTF-8")

knitr::opts_chunk$set(cache = FALSE, 
                      echo = FALSE, 
                      tidy = TRUE, 
                      collapse = TRUE, 
                      dpi = 150, 
                      fig.width = 8, 
                      fig.height = 5, 
                      fig.align = "center", 
                      warning = FALSE, 
                      message = FALSE, 
                      encoding = "UTF-8")

```


<!-- Import and manage data -->
```{r get_data}
# The list of countries
df_countries <- readRDS(file.path(path.local.worldwide.data, 'df_countries.RDS'))

# The MSF linelist
update_msf_data_anyways <- FALSE

get_updated_msf_data <- ifelse(isTRUE(update_msf_data_anyways) | 
                                 !file.exists(file.path(path.local.msf.data, 'dta_MSF.RData')), 
                               TRUE, 
                               FALSE)

source(file.path(path.R, 'run_get_msf_data.R'), encoding = 'UTF-8') 

```



# Overview

**Information** of completeness of data (number of MSF projects that are active on Covid19 care, reporting individual data (linelist), reporting aggregated data and not reporting data). Number of projects will be listed by continent and region. 
 
```{r, eval = FALSE}

# This will be a summary table
# We still not have the necessary information to produce this table

```


**Table** of total number of patients screened, by sex and age group

```{r, eval = FALSE}

# Also fo this table we do not have the necessary data yet

```


**Table** with list of projects by country with:

  - counts by Covid status: confirmed, probable, suspected, not a case and unknown 
  - dates of first and last consultation/hospitalisation

```{r, eval = TRUE}

# ALREADY DONE

```


  
## - TO BE DONE ASAP (PAUL)
**Histogram** of number of consultation and admissions aggregated by week: overall and by region (or continent)

```{r}

# TO BE DONE ASAP (PAUL)

```


## - TO BE DONE ASAP (PAUL)
**Histogram** of number of confirmed, probable, suspected, not a case and unknown aggregated by week: overall and by region (or continent)

```{r}

# TO BE DONE ASAP (PAUL)

```


## - To be done (FRANCESCO)

**Table** by continent (or region):
  
- Proportion of positive tests
- Proportion of suspect cases sent home after screening
- Proportion of asymptomatic among the tested positive

```{r}

```


**Table** by Covid status

  - Age (median, IQR, min-max)
  - Sex

```{r}

# Done

```


**Table** by continent (or region):

- Frequency and proportion of symptoms by Covid status 
- Frequency and proportion of patients with at least 1 comorbidity

```{r}

# Already a table to b adapted

```


## - TO BE DONE ASAP (FRANCESCO)

**Table Case-fatality risk** by Covid status
- Case-fatality risk among suspect, probable and confirmed, and by consultation/admission

```{r}

```


## - TO BE DONE ASAP (FRANCESCO)

# Description of Covid19 confirmed patients

**Age/sex pyramid** by continent

```{r}


```


**Table** by continent (by region, if possible):

  -	Characteristics: Number and proportion of cases exposed through healthcare facility (healthcare worker + visited) 
  -	Characteristics: Number and proportion of cases with chronic diseases (comorbidities)

```{r}

```
  
  
**Table Patient’s care 1**

  - Admitted to a hospital/inpatient department, 
  - received oxygen, 
  - admitted to ICU, 
  - supported with ventilator, 
  - supported with ECMO
  
```{r}

# ALREADY DONE

```


**Table Patient’s care 2**

- medication received (under discussion)

```{r}

# List of medications to be finalised

```



# - TO BE DONE ASAP (Francesco)

**Table Outcome 1 - Count outcome** by country

count of cured, deaths, transferred… 
  
```{r}

```


**Table Outcome 2 - Hospital case fatality risk** by continent (region ??)
  
  - overall
  - by age group
  - by sex
  - by comorbidity

```{r}

# DONE

```


**Table Patients' characteristics**

Frequency and proportion of symptoms, by case severity


```{r}

# TO BE DONE as soon as the proper definition of severity is available

```


# Evolution overtime of the following indicators among the Covid19 confirmed cases


**With all confirmed cases as denominator: time-series and trends plot** with week as time unit of the following indicator: 

Time-series plot: 

  -	Proportion of admitted over the confirmed
  -	Proportion of patients with severe symptoms
  -	Delay from symptoms onset to consultation/admission
  -	Length of staying for patients hospitalised
  -	Case fatality risk for patients hospitalised

## - TO BE DONE ASAP (PAUL)

```{r}

# TO BE DONE ASAP (PAUL)

```


**With all confirmed cases with severe or critical symptoms as denominator: time-series and trends plot** with week as time unit of the following indicator: 

  -	Proportion of patients who received oxygen
  -	Proportion of patients admitted to ICU
  -	Proportion of patients supported by ventilator
  -	Proportion of patients supported by ECMO

```{r}

# TO BE DONE as soon as patients are properly classified by severity

```
