---
title: "Covid-19 MSF data analysis - OC level"
params:
  OC: OCB
author:
  - name: Francesco Grandesso
    url: mailto:francesco.GRANDESSO@epicentre.msf.org
    affiliation: Epicentre
    affiliation_url: https://epicentre.msf.org/
  - name: Paul Campbell
    url: mailto:paul.CAMPBELL@epicentre.msf.org
    affiliation: Epicentre
    affiliation_url: https://epicentre.msf.org/
  - name: Mathilde Mousset
    url: mailto:mathilde.MOUSSET@epicentre.msf.org
    affiliation: Epicentre
    affiliation_url: https://epicentre.msf.org/
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  rmdformats::readthedown:
    highlight: "tango"
    css: styles_blues.css
    lightbox: TRUE
    gallery: TRUE
    use_bookdown: TRUE
    toc_depth: 3
editor_options:
  chunk_output_type: console
---

<!-- Setup and uploading functions -->

```{r setup, warning = FALSE, message = FALSE, include = FALSE}

if(Sys.getlocale(category = "LC_TIME") == "French_France.1252") {
  Sys.setlocale(category = "LC_TIME", locale = "English_United Kingdom.1252")
  Sys.setenv(LANG = "en_GB.UTF-8") 
}

if (!exists('path.root')) {
  source(here::here('R', 'setup.r'), encoding = 'UTF-8')
}


knitr::opts_chunk$set(cache = FALSE,
                      echo  = FALSE,
                      tidy  = TRUE,
                      collapse = TRUE,
                      dpi = 150,
                      fig.width  = 8,
                      fig.height = 5,
                      fig.align  = "center",
                      warning  = FALSE,
                      message  = FALSE,
                      encoding = "UTF-8")


# Upload functions
source(file.path(path.R, 'utils_management.R'), encoding = 'UTF-8')
source(file.path(path.R, 'utils_get_data.R')  , encoding = 'UTF-8')
source(file.path(path.R, 'utils_vis.R')       , encoding = 'UTF-8')


# Set the left and right censoring for date of consultation.
# The right-censoring create also the week value and the folders where to save outputs
dates_and_week <- set_date_frame(create_folders = TRUE)

date_min_report <- dates_and_week[[1]]
date_min_report <- as.Date("2020-01-22")
date_max_report <- dates_and_week[[2]]
week_report     <- dates_and_week[[3]]



### SETUP PATHS ###
# If week folder not created already by overall sitrep, create it.
path.local.week.oc       <- file.path(path.local, week_report, "oc")
path.local.msf.data.oc   <- file.path(path.local.week.oc, params$OC, 'data')
path.local.msf.tables.oc <- file.path(path.local.week.oc, params$OC, 'tables')
path.local.msf.graphs.oc <- file.path(path.local.week.oc, params$OC, 'graphs')
path.local.msf.graphs.oc.per.project <- file.path(path.local.week.oc, params$OC, 'graphs', 'per_project')

if (!dir.exists(path.local.week.oc)) {
  dir.create(path.local.week.oc,   showWarnings = FALSE, recursive = TRUE)
}
if (!dir.exists(path.local.msf.data.oc)) {
  dir.create(path.local.msf.data.oc,   showWarnings = FALSE, recursive = TRUE)
}
if (!dir.exists(path.local.msf.graphs.oc)) {
  dir.create(path.local.msf.graphs.oc, showWarnings = FALSE, recursive = TRUE)
}
if (!dir.exists(path.local.msf.graphs.oc.per.project)) {
  dir.create(path.local.msf.graphs.oc.per.project, showWarnings = FALSE, recursive = TRUE)
}
if (!dir.exists(path.local.msf.tables.oc)) {
  dir.create(path.local.msf.tables.oc, showWarnings = FALSE, recursive = TRUE)
}


### GRAPHICAL PARAMETERS ###
caption_graph <- glue('Analysis and graphic by Epicentre | Data source: MSF linelists as of {format(date_max_report, "%d %b %Y")}')
```


<!-- Import and manage data -->

```{r get-data, include = FALSE}

# === === === === === === ===
# Geo and population data
# === === === === === === ===

df_countries <- readRDS(file.path(path.local.data, 'df_countries.RDS'))



# === === === === === === ===
# Linelist data
# === === === === === === ===

dta_linelist <- covidutils::get_msf_linelist(force = TRUE) %>%
  covidutils::prepare_msf_dta(shorten_var_names = FALSE) %>%
  rename(iso_a3 = country) %>%
  left_join(df_countries %>% select(continent, region, iso_a3, country), by = 'iso_a3') %>%
  mutate(
    country   = as.factor(country),
    continent = as.factor(continent),
    delay_before_consultation = as.numeric(MSF_date_consultation - patcourse_dateonset)) %>%
  filter(between(MSF_date_consultation,
                 left = date_min_report, right = date_max_report) | is.na(MSF_date_consultation)) %>%  # Filter date limits using date of consultation
  filter(across(
    .cols = ends_with(params$OC),
    .fns  = ~ .x == TRUE))


# Data for per project analysis: only the confirmed, probable and suspects
dta_linelist_project <- dta_linelist %>%
  filter(ind_MSF_covid_status %in% c('Confirmed', 'Probable', 'Suspected')) %>%
  mutate(MSF_severity = ifelse(is.na(MSF_severity), "Non available", MSF_severity),
         MSF_severity = factor(MSF_severity,
                           levels = c("Mild", "Moderate", "Severe", "Critical",
                                      "Non available")),
         country_project = paste(iso_a3, project, sep = "_") )



# === === === === === === ===
# Aggregated data
# === === === === === === ===

dta_weekly_aggregated        <- get_msf_aggregated(force = TRUE)
dta_project_aggregated_dates <- get_msf_aggregated_dates(force = TRUE)

### Remove temporarily Quito that grouped the whole activity in a single week
dta_weekly_aggregated <- dta_weekly_aggregated %>%
  filter(project != 'Quito-Fixed & mobile brigades')


### Expand the aggregated
dta_aggregated_expanded <- dta_weekly_aggregated %>%
  filter(!project %in% c('Lesvos-Moria', 'Lesvos-Moria Ped', "Quito")) %>%
  select(-week) %>%
  mutate(
    date = as.Date(date),
    epi_week_consultation = make_epiweek_date(date) %>% as.Date(),
    country = case_when(
      country == 'DRC' ~ 'Democratic Republic of the Congo',
      country == "Côte d'Ivoire" ~ 'Côte d’Ivoire',
      country == "DRC" ~ "Democratic Republic of Congo",
      TRUE ~ country)) %>%
  left_join(df_countries, by = 'country') %>%
  pivot_longer(cols = c('confirmed', 'probable', 'suspected', 'non_cases', 'unknown'),
               names_to = 'ind_MSF_covid_status') %>%
  filter(!is.na(value)) %>%
  mutate(obs = purrr::map(value, ~rep_len(1, .x))) %>%
  unnest(cols = c(obs)) %>%
  select(-c(value, obs)) %>%
  mutate(
    ind_MSF_covid_status = factor(ind_MSF_covid_status,
                          levels = c('confirmed', 'probable', 'suspected',
                                     'non_cases', 'unknown'),
                          labels = c('Confirmed', 'Probable', 'Suspected',
                                     'Not a case', 'Unknown'))) %>%
  rename('OC' = oc,
         'site_name' = project) %>%
  filter(between(date, left = date_min_report, right = date_max_report) | is.na(date)) %>%
  filter(OC == params$OC)


# === === === === === === ===
# Merge datasets
# === === === === === === ===

# The dataset with linelist and the expanded aggregated data together
dta_linelist_with_aggregated <- dta_linelist %>%
  select(continent, region, country, site_name, iso_a3, OC,
         epi_week_consultation, ind_MSF_covid_status, project) %>%
  add_row(dta_aggregated_expanded %>%
            select(continent, region, country, site_name, iso_a3,
                   OC, epi_week_consultation, ind_MSF_covid_status) %>%
            mutate(
              country = paste(country, '(*)'),
              site_name = paste(site_name, '(*)'))) %>%
  mutate(
    project = case_when(
      is.na(project) ~ site_name,
      TRUE           ~ project),
    ind_MSF_covid_status = recode(ind_MSF_covid_status, "(Unknown)" = "Unknown")
  ) %>% 
  mutate(continent = case_when(
      country == "Cote d'Ivoire (*)" ~ 'Africa', 
      country == "Democratic Republic of Congo (*)" ~ 'Africa',
      country == "Palestine (*)" ~ "Asia",
      TRUE ~ continent))


tbl_aggregated_expanded_with_dates <- dta_aggregated_expanded %>%
  distinct(sheet, OC, country, site_name) %>%
  full_join(dta_project_aggregated_dates, by = 'sheet') %>%
  mutate(
    country = case_when(
      country == 'DRC' ~ 'Democratic Republic of the Congo',
      country == "Côte d'Ivoire" ~ 'Côte d’Ivoire',
      TRUE ~ country)) %>%
  pivot_longer(
    cols = date_first:date_last,
    names_to = "type_date",
    values_to = "MSF_date_consultation") %>%
  left_join(df_countries, by = 'country') %>%
  mutate(
    country = paste(country, '(*)'),
    site_name = paste(site_name, '(*)'))
```

---
subtitle: "Data as of week `r glue::glue("{format(date_max_report, '%V')}, ending {weekdays(date_max_report)} {format(date_max_report, '%d %B %Y')}")`"
---

# Overview `r params$OC`

This section will present **information of completeness of data** (number of MSF projects that are active on Covid-19 care, reporting individual data (linelist), reporting aggregated data and not reporting data). Number of projects will be listed by continent and region.

At the date the required data are not yet available. And this section will be updated as soon as they will be available.

Additional graphics and tables are also available at the [Epicentre COVID-19 Epi Dashboard](https://reports.msf.net/public/covid19/).

Reports of previous weeks are available [here](https://reports.msf.net/secure/app_direct/covid19-additional-analysis/additional_episitrep_outputs_oc/archive/).

<!-- Some figure outputs to be used in the text -->
```{r numbers, include = FALSE}
nb_msf_sites_linelist   <- length(unique(dta_linelist$site_name))              # Nb of sites using the linelist
nb_msf_sites_aggregated <- length(unique(dta_aggregated_expanded$site_name))   # Nb of sites using aggregated data
nb_msf_sites            <- nb_msf_sites_linelist + nb_msf_sites_aggregated     # Sum of the two above


call_msf_countries <- unique(c(as.character(dta_linelist$country),
                               as.character(dta_aggregated_expanded$country))) %>%
  sort() # List of countries reporting data
nb_msf_countries   <- length(call_msf_countries)

nb_msf_obs_linelist   <- dta_linelist %>% count() %>% pull(n)         # Nb of observations in the linelist
nb_msf_obs_aggregated <- dta_aggregated_expanded %>% count() %>% pull(n)  # Nb of observations in aggreagated data
nb_msf_obs            <- nb_msf_obs_linelist + nb_msf_obs_aggregated

nb_msf_confirmed <- sum(dta_linelist_with_aggregated$ind_MSF_covid_status == 'Confirmed', na.rm = TRUE)
nb_msf_probable  <- sum(dta_linelist_with_aggregated$ind_MSF_covid_status == 'Probable' , na.rm = TRUE)
nb_msf_suspected <- sum(dta_linelist_with_aggregated$ind_MSF_covid_status == 'Suspected', na.rm = TRUE)


nb_msf_confirmed_linelist <- sum(dta_linelist$ind_MSF_covid_status == 'Confirmed', na.rm = TRUE)

nb_msf_confirmed_with_comorbidity <- with(dta_linelist,
                                          sum(ind_MSF_covid_status == 'Confirmed' & ind_Comcond_count > 0,
                                              na.rm = TRUE))

nb_msf_conf_prob_susp <- with(dta_linelist,
                              sum(ind_MSF_covid_status %in% c('Confirmed', 'Probable', 'Suspected') & ind_outcome_patcourse_status %in% c('Cured', 'Died'), na.rm = TRUE))

nb_msf_conf_prob_susp_who_died <- with(dta_linelist,
                                       sum(ind_MSF_covid_status %in% c('Confirmed', 'Probable', 'Suspected') & ind_outcome_patcourse_status == 'Died', na.rm = TRUE))

nb_msf_conf_above65 <- with(dta_linelist,
                            sum(ind_MSF_covid_status == 'Confirmed' & age_in_years >= 65 & ind_outcome_patcourse_status %in% c('Cured', 'Died'), na.rm = TRUE))

nb_msf_conf_above65_who_died <- with(dta_linelist,
                                     sum(ind_MSF_covid_status == 'Confirmed' & age_in_years >= 65 & ind_outcome_patcourse_status == 'Died', na.rm = TRUE))


median_age_all <- dta_linelist %>%
  pull(age_in_years) %>%
  median(na.rm = TRUE)


median_age_all_by_sex <- dta_linelist %>%
  group_by(patinfo_sex) %>%
  summarise(
    age_median = median(age_in_years, na.rm = TRUE))


median_age_confirmed <- dta_linelist %>%
  filter(ind_MSF_covid_status == 'Confirmed') %>%
  pull(age_in_years) %>%
  median(na.rm = TRUE)


median_age_confirmed_by_sex <- dta_linelist %>%
  filter(ind_MSF_covid_status == 'Confirmed') %>%
  group_by(patinfo_sex) %>%
  summarise(
    age_median = median(age_in_years, na.rm = TRUE))


median_age_confirmed_died <- dta_linelist %>%
  filter(ind_MSF_covid_status == 'Confirmed' & ind_outcome_patcourse_status == 'Died') %>%
  pull(age_in_years) %>%
  median(na.rm = TRUE)
```


**Number of MSF sites**: `r nb_msf_sites` (`r nb_msf_sites_linelist` sites with linelist data and `r nb_msf_sites_aggregated` with aggregated data).

**Total patients consulted/admitted**: `r format(nb_msf_obs, big.mark = ',')` in `r length(call_msf_countries)` countries (`r combine_words(call_msf_countries)`).

**Number of consultation reported through the linelist**: `r format(nb_msf_obs_linelist, big.mark = ',')`.

**Number of consultation reported through the aggregated data reporting**: `r format(nb_msf_obs_aggregated, big.mark = ',')`.

Of all consultations/admissions patients were in relation with the Covid-19 infection:

-   `r format(nb_msf_confirmed, big.mark = ',')` confirmed
-   `r format(nb_msf_probable , big.mark = ',')` probable
-   `r format(nb_msf_suspected, big.mark = ',')` suspected

------------------------------------------------------------------------

# All patients consulted/admitted

This section provides an overview of **all patients that attended an MSF supported Covid-19 health facility**: number of patients consulted, but not admitted and number of consulted and admitted to the health facility.

Other results are presented by Covid status (confirmed, probable, suspected, not a case and unknown) and listed by country or site and grouped by continent.

## Time & place

### Weekly number of consultations (including admission) by Covid status

Aggregated data included

<!-- Histograms of patients Covid status -->
```{r hist-covid-status, fig.cap = 'Weekly number of consultations (including admission) by Covid status'}
# DONE by Paul

tbl_epicurve_status <- dta_linelist_with_aggregated %>%
  count(ind_MSF_covid_status, epi_week_consultation)

hist_epicurve_status <- tbl_epicurve_status %>%
  ggplot(aes(epi_week_consultation, n, fill = ind_MSF_covid_status)) +
  geom_col() +
  ggthemes::scale_fill_tableau(name = "Status", palette = "Tableau 20") +
  scale_x_date(name = "Week of Consultation",
               date_breaks = "2 week", date_labels = "%V",
               expand = expansion(mult = c(0.01, 0.01)),
               sec.axis = ggplot2::sec_axis(trans = ~ as.Date(.))) + #labels = function(x) format(x, "%b-%Y")
  scale_y_continuous(name = "Patients", expand = expansion(mult = c(0, 0.02))) +
  theme_light() +
  theme(legend.position = 'top', panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())

ggsave(filename = paste0('hist_epicurve_status', '_', week_report, '.png'),
       path = path.local.msf.graphs.oc,
       plot = hist_epicurve_status,
       width = 6,
       height = 4,
       scale = 1.1,
       dpi = 320)


hist_epicurve_status
```


<!-- Histograms BY CONTINENT of patients Covid status -->
```{r hist-covid-status-facet-continent, fig.cap = 'Weekly number of consultations (including admission) by Covid status, in each continent where MSF is operating'}
# DONE by Francesco (copied from the Paul graph just above)

tbl_epicurve_status_continent <- dta_linelist_with_aggregated %>%
  count(continent, ind_MSF_covid_status, epi_week_consultation)


hist_epicurve_status_continent <- tbl_epicurve_status_continent %>%
  ggplot(aes(epi_week_consultation, n, fill = ind_MSF_covid_status)) +
  facet_wrap(vars(continent), scales = 'free_y') +
  geom_col() +
  ggthemes::scale_fill_tableau(name = NULL, palette = "Tableau 20") +
  scale_x_date(name = "Week of Consultation", date_breaks = "2 week", date_labels = "%V", expand = expansion(mult = c(0.01, 0.01)),
               sec.axis = ggplot2::sec_axis(trans = ~ as.Date(.))) + #labels = function(x) format(x, "%b-%Y")
  scale_y_continuous(name = "Patients", expand = expansion(mult = c(0, 0.02))) +
  labs(y = "Patients", caption = 'NOTE: free y axis scale among graphics') +
  theme_light() +
  theme(legend.position = 'top', panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())

ggsave(file.path(path.local.msf.graphs.oc, paste0('hist_epicurve_status_continent', '_', week_report, '.png')),
       plot = hist_epicurve_status_continent,
       width = 6,
       height = 4,
       scale = 1.1,
       dpi = 320)

hist_epicurve_status_continent
```

### PCR tests

```{r hist-testing}
dta_testing <- dta_linelist %>%
  select(country, project, epi_week_consultation, merge_admit, ind_MSF_covid_status,
         ind_outcome_patcourse_status, matches("test|lab|sample")) %>%
  rowwise %>%
  mutate(
    test_1_10 = ifelse(MSF_test_type   == "PCR", 1, 0),
    test_2_10 = ifelse(MSF_test_type_2 == "PCR", 1, 0),
    test_3_10 = ifelse(MSF_test_type_3 == "PCR", 1, 0),
    test_4_10 = ifelse(MSF_test_type_4 == "PCR", 1, 0),
    n_test_PCR = sum(test_1_10, test_2_10, test_3_10, test_4_10, na.rm = TRUE),
    merge_test_PCR = ifelse(n_test_PCR >= 1, "Yes", "No"),
    
    test_result_1 = ifelse(MSF_test_type   == "PCR" & MSF_test_results       == "Positive", 1, 0),
    test_result_2 = ifelse(MSF_test_type_2 == "PCR" & MSF_test_results_2     == "Positive", 1, 0),
    test_result_3 = ifelse(MSF_test_type_3 == "PCR" & outcome_lab_result == "Positive", 1, 0),
    test_result_4 = ifelse(MSF_test_type_4 == "PCR" & MSF_test_results_4     == "Positive", 1, 0),
    n_test_PCR_positive = sum(test_result_1, test_result_2,
                              test_result_3, test_result_4, na.rm = TRUE),
    merge_test_PCR_positive = ifelse(n_test_PCR_positive >= 1, "Yes", "No")
  )


# Graph

scaling_factor <- 0.01

dta_testing %>%
  group_by(epi_week_consultation) %>%
  summarise(n_patients_PCR = sum(merge_test_PCR == "Yes", na.rm = TRUE),
            n_patients_PCR_positive = sum(merge_test_PCR_positive == "Yes", na.rm = TRUE),
            perct_patient_pos = ifelse(n_patients_PCR == 0,
                                       NA_integer_,
                                       n_patients_PCR_positive / n_patients_PCR),
            
            nb_test_PCR = sum(n_test_PCR, na.rm = TRUE),
            nb_tests_PCR_positive =sum(n_test_PCR_positive, na.rm = TRUE),
            perct_test_pos = ifelse(nb_test_PCR == 0,
                                    NA_integer_,
                                    nb_tests_PCR_positive / nb_test_PCR)) %>%
  
  
  ggplot(aes(x = epi_week_consultation)) +
  geom_col(aes(y = n_patients_PCR),
           fill = "#A0CBE8") +
  geom_line(aes(y = perct_patient_pos / scaling_factor),
            colour = "red") +
  labs(x = "Week of consultation",
       y = "Number of patients tested",
       title = "Number of patient tested and positivity rates") +
  
  scale_x_date(name = "Week of Consultation",
               date_breaks = "2 weeks",
               date_labels = "%V",
               sec.axis = ggplot2::sec_axis(trans = ~ .),
               expand = expansion(mult = c(0.01, 0.01))) +
  
  scale_y_continuous(sec.axis = ggplot2::sec_axis(~ .* scaling_factor,
                                                  name = "Positivity rate",
                                                  labels = scales::percent_format(accuracy = 1)),
                     expand = expansion(mult = c(0, 0.02)))

```

### Weekly number of admissions

Aggregated data excluded

<!-- Histogram OVERALL of number of consultation/admissions -->
```{r hist-consultations-sdmissions, fig.cap = paste('Weekly number of consultations and admissions in an MSF supported COvid-19 health service as of', format(date_max_report, "%d %B %Y"))}
# DONE by Paul

tbl_epicurve_consultation <- dta_linelist %>%
  filter(merge_admit != "Unknown") %>%
  count(epi_week_consultation, merge_admit) %>%
  drop_na() %>%
  mutate(merge_admit = recode(merge_admit, "Yes" = "Admitted", "No" = "Not Admitted"))

tbl_admit_prop <- dta_linelist %>%
  filter(merge_admit != "Unknown") %>%
  drop_na(epi_week_consultation) %>%
  group_by(epi_week_consultation) %>%
  summarise(
    n = n(),
    total = sum(merge_admit %in% c("Yes", "No")),
    admitted = sum(merge_admit == "Yes"),
    admit_prop = admitted / total
  ) %>%
  ungroup()

n_max   <- max(tbl_admit_prop$n, na.rm = TRUE)
cfr_max <- rounder(max(tbl_admit_prop$admit_prop, na.rm = TRUE), .1)
cfr_max <- ifelse(cfr_max > 1, 1, cfr_max)
scaling_factor <- cfr_max / n_max # sets the y limit of CFR rounded up to nearest 10% above cfr max

missing_consultation <- sum(is.na(dta_linelist$epi_week_consultation))
missing_admission    <- sum(dta_linelist$merge_admit == "Unknown")

hist_epicurve_consult_admit <- tbl_epicurve_consultation %>%
  ggplot(aes(epi_week_consultation, n)) +
  geom_col(aes(fill = merge_admit)) +
  geom_line(data = tbl_admit_prop,
            aes(y = admit_prop / scaling_factor, colour = "Admitted %"),
            key_glyph = "timeseries", size = 1) +
  ggthemes::scale_fill_tableau(name = NULL, palette = "Tableau 20") +
  scale_x_date(name = "Week of Consultation", date_breaks = "2 weeks", date_labels = "%V",
               sec.axis = ggplot2::sec_axis(trans = ~ .), expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(sec.axis = ggplot2::sec_axis(~ . * scaling_factor, name = "Admitted", labels = scales::percent_format(accuracy = 1)),
                     expand = expansion(mult = c(0, 0.02))) +
  labs(y = "Patients", colour = NULL, caption = glue::glue("Missing data: Consultation Date {missing_consultation}, Admission: {missing_admission}")) +
  theme_light() +
  theme(legend.position = "top", panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())


ggsave(file.path(path.local.msf.graphs.oc, paste0('hist_epicurve_consult_admit', '_', week_report, '.png')),
       plot = hist_epicurve_consult_admit,
       width = 6,
       height = 4,
       scale = 1.1,
       dpi = 320)


hist_epicurve_consult_admit
```


<!-- Histogram BY CONTINENT of number of consultation/admissions -->
```{r hist-consultations-admissions-facet-continent, fig.cap = paste('Weekly number of consultations and admissions in an MSF supported COvid-19 health service by continent as of', format(date_max_report, "%d %B %Y"))}
# DONE by Paul

tbl_epicurve_consultation_continent <- dta_linelist %>%
  filter(merge_admit != "Unknown") %>%
  count(continent, epi_week_consultation, merge_admit) %>%
  drop_na() %>%
  mutate(merge_admit = recode(merge_admit, "Yes" = "Admitted", "No" = "Not Admitted"))

tbl_admit_prop_continent <- dta_linelist %>%
  filter(merge_admit != "Unknown") %>%
  drop_na(epi_week_consultation) %>%
  group_by(continent, epi_week_consultation) %>%
  summarise(
    n = n(),
    total = sum(merge_admit %in% c("Yes", "No")),
    admitted = sum(merge_admit == "Yes"),
    admit_prop = admitted / total
  ) %>%
  ungroup()

n_max <- max(tbl_admit_prop_continent$n, na.rm = TRUE)
cfr_max <- rounder(max(tbl_admit_prop_continent$admit_prop, na.rm = TRUE), .1)
cfr_max <- ifelse(cfr_max > 1, 1, cfr_max)
scaling_factor <- cfr_max / n_max # sets the y limit of CFR rounded up to nearest 10% above cfr max

missing_consultation <- sum(is.na(dta_linelist$epi_week_consultation))
missing_admission    <- sum(dta_linelist$merge_admit == "Unknown")

hist_epicurve_consult_admit_continent <- tbl_epicurve_consultation_continent %>%
  ggplot(aes(epi_week_consultation, n)) +
  facet_wrap(~continent) +
  geom_col(aes(fill = merge_admit)) +
  geom_line(data = tbl_admit_prop_continent,
            aes(y = admit_prop / scaling_factor, colour = "Admitted %"),
            key_glyph = "timeseries", size = 0.5) +
  ggthemes::scale_fill_tableau(name = NULL, palette = "Tableau 20") +
  scale_x_date(name = "Week of Consultation", date_breaks = "3 week", date_labels = "%V",
               sec.axis = ggplot2::sec_axis(trans = ~ .), expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(sec.axis = ggplot2::sec_axis(~ . * scaling_factor, name = "Admitted", labels = scales::percent_format(accuracy = 1)),
                     expand = expansion(mult = c(0, 0.02))) +
  labs(y = "Patients", colour = NULL, caption = glue::glue("Missing data: Consultation Date {missing_consultation}, Admission: {missing_admission}")) +
  theme_light() +
  theme(legend.position = "top", panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())


ggsave(file.path(path.local.msf.graphs.oc, paste0('hist_epicurve_consult_admit_continent', '_', week_report, '.png')),
       plot = hist_epicurve_consult_admit_continent,
       width = 6,
       height = 4,
       scale = 1.1,
       dpi = 320)

hist_epicurve_consult_admit_continent
```


```{r gtbl-consultation-continent, layout = "l-body-outset"}
tbl_consultation_continent <- dta_linelist %>%
  filter(merge_admit != "Unknown") %>%
  mutate(merge_admit = recode(merge_admit, "Yes" = "Admitted", "No" = "Not Admitted")) %>%
  count(continent, merge_admit) %>%
  pivot_wider(names_from = merge_admit, values_from = n) %>%
  adorn_totals(c('col', 'row')) %>%
  mutate(
    admitted_percent = Admitted / Total* 100)

gtbl_consultation_continent <- tbl_consultation_continent %>%
  gt(rowname_col = "continent") %>%
  cols_label(
    admitted_percent = "Admitted (%)") %>%
  fmt_number(
    columns = vars(Admitted, `Not Admitted`, Total),
    decimals = 0) %>%
  fmt_number(
    columns = vars(admitted_percent),
    decimals = 1) %>%
  cols_align(
    align = 'right',
    columns = vars(Admitted, `Not Admitted`, Total, admitted_percent)) %>%
  cols_width(
    vars(Admitted, `Not Admitted`, Total) ~ px(120),
    everything() ~ px(100)) %>%
  tab_options(
    column_labels.font.weight = "bold") %>%
  tab_header(title = "Number of admissions by continent")

gtbl_consultation_continent
```


### Summary tables of patients consulted/admitted to an MSF supported COvid-19 health facility

<!--
=== SUMMARY TABLE BY COUNTRY ===
Counts of patients consulted/admitted to an MSF supported health facility by Covid
status (confirmed, probable, suspected, not a case and unknown) listed by country
-->
```{r, tbl-covid-status-country, layout = "l-page"}
# DONE by Francesco
tbl_countries_ind_MSF_covid_status <- dta_linelist_with_aggregated %>%
  select(continent, country, site_name, ind_MSF_covid_status) %>%
  group_by(continent, country) %>%
  summarise(
    total = n(),
    n_confirmed = sum(ind_MSF_covid_status == 'Confirmed'),
    n_probable  = sum(ind_MSF_covid_status == 'Probable'),
    n_suspected = sum(ind_MSF_covid_status == 'Suspected'),
    n_not_case  = sum(ind_MSF_covid_status == 'Not a case'),
    n_unknown   = sum(ind_MSF_covid_status == '(Unknown)'),
    p_confirmed = n_confirmed / total,
    p_probable  = n_probable  / total,
    p_suspected = n_suspected / total,
    p_not_case  = n_not_case  / total,
    p_unknown   = n_unknown   / total) %>%
  ungroup() %>%
  arrange(continent, country)

tbl_countries_dates <- dta_linelist %>%
  select(continent, country, site_name, MSF_date_consultation) %>%
  add_row(tbl_aggregated_expanded_with_dates %>% select(continent, country, site_name, MSF_date_consultation)) %>%
  group_by(country) %>%
  summarise(
    date_adm_first = format(min(MSF_date_consultation, na.rm = TRUE), '%d-%m-%Y'),
    date_adm_last  = format(max(MSF_date_consultation, na.rm = TRUE), '%d-%m-%Y'))

gtbl_countries_ind_MSF_covid_status <- tbl_countries_ind_MSF_covid_status %>%
  left_join(tbl_countries_dates) %>%
  gt(rowname_col = "country",
     groupname_col = "continent") %>%
  cols_label(
    country     = 'Country',
    total       = 'Total',
    n_confirmed = 'n',
    n_probable  = 'n',
    n_suspected = 'n',
    n_not_case  = 'n',
    n_unknown   = 'n',
    p_confirmed = '(%)',
    p_probable  = '(%)',
    p_suspected = '(%)',
    p_not_case  = '(%)',
    p_unknown   = '(%)',
    date_adm_first = 'First',
    date_adm_last  = 'Last') %>%
  tab_spanner(
    label = html('Dates of<br>consultation/hospitalisation'),
    columns = vars(date_adm_first, date_adm_last)) %>%
  tab_spanner(
    label = html('Confirmed'),
    columns = ends_with('_confirmed')) %>%
  tab_spanner(
    label = html('Probable'),
    columns = ends_with('_probable')) %>%
  tab_spanner(
    label = html('Suspected'),
    columns = ends_with('_suspected')) %>%
  tab_spanner(
    label = html('Not a case'),
    columns = ends_with('_not_case')) %>%
  tab_spanner(
    label = html('(Unknown)'),
    columns = ends_with('_unknown')) %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = "({x})") %>%
  fmt_missing(
    columns = starts_with('date_adm_'),
    missing_text = "Unknown") %>%
  cols_align(
    align = 'right',
    columns = vars(total)) %>%
  cols_align(
    align = 'left',
    columns = starts_with('p_')) %>%
  cols_align(
    align = 'center',
    columns = starts_with('date_adm_')) %>%
  tab_style(
    style = list(
      cell_text(align = "left")),
    locations = cells_column_labels(columns = starts_with('p_'))) %>%
  summary_rows(
    groups = TRUE,
    columns = vars(total, n_confirmed, n_probable, n_suspected, n_not_case, n_unknown),
    fns = list('Country Total' = ~ sum(.)),
    missing_text = "",
    formatter = fmt_number,
    decimals = 0) %>%
  grand_summary_rows(
    columns = vars(total, n_confirmed, n_probable, n_suspected, n_not_case, n_unknown),
    fns = list('Grand Total' = ~ sum(.)),
    missing_text = "",
    formatter = fmt_number,
    decimals = 0) %>%
  tab_source_note(
    source_note = "(*) These data were reported as aggregated only and are not included in further analysis below") %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = 'bold',
    grand_summary_row.text_transform = 'uppercase',
    data_row.padding = px(1),
    row_group.padding = px(1),
    summary_row.padding = px(1),
    grand_summary_row.padding = px(1))

gtsave(gtbl_countries_ind_MSF_covid_status,
       file.path(path.local.msf.tables.oc, paste0('gtbl_countries_ind_MSF_covid_status', '_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_countries_ind_MSF_covid_status,
       file.path(path.local.msf.tables.oc, paste0(week_report, '_SUMMARY-TABLE_MSF-countries_by_patients_Covid-status.html'))) %>%
  invisible()


gtbl_countries_ind_MSF_covid_status %>%
  tab_header(
    title = html(paste('SUMMARY TABLE<br>
                 Counts of patients consulted/admitted to an MSF supported COvid-19 health service by Covid status listed by country as of', format(date_max_report, "%d %b %Y"))))
```


<!--
=== SUMMARY TABLE BY SITE ===
Not in the EpiSitrep -- Too long
Output in html file that will be available through a link
-->
```{r tbl-covid-status-site, layout = "l-page"}
# DONE by Francesco

tbl_sites_ind_MSF_covid_status <- dta_linelist_with_aggregated %>%
  mutate(country = gsub(" \\(\\*\\)", "", country)) %>%
  group_by(country, site_name) %>%
  summarise(
    total = n(),
    n_confirmed = sum(ind_MSF_covid_status == 'Confirmed'),
    n_probable  = sum(ind_MSF_covid_status == 'Probable'),
    n_suspected = sum(ind_MSF_covid_status == 'Suspected'),
    n_not_case  = sum(ind_MSF_covid_status == 'Not a case'),
    n_unknown   = sum(ind_MSF_covid_status == '(Unknown)'),
    p_confirmed = n_confirmed / total,
    p_probable  = n_probable  / total,
    p_suspected = n_suspected / total,
    p_not_case  = n_not_case  / total,
    p_unknown   = n_unknown   / total) %>%
  ungroup()

tbl_site_dates <- dta_linelist %>%
  select(continent, country, site_name, MSF_date_consultation) %>%
  add_row(tbl_aggregated_expanded_with_dates %>% select(continent, country, site_name, MSF_date_consultation)) %>%
  group_by(site_name) %>%
  summarise(
    date_adm_first = format(min(MSF_date_consultation, na.rm = TRUE), '%d-%m-%Y'),
    date_adm_last  = format(max(MSF_date_consultation, na.rm = TRUE), '%d-%m-%Y'))

gtbl_sites_ind_MSF_covid_status <- tbl_sites_ind_MSF_covid_status %>%
  left_join(tbl_site_dates) %>%
  gt(rowname_col = "site_name",
     groupname_col = "country") %>%
  cols_label(
    site_name   = 'Site',
    total       = 'Total',
    n_confirmed = 'n',
    n_probable  = 'n',
    n_suspected = 'n',
    n_not_case  = 'n',
    n_unknown   = 'n',
    p_confirmed = '(%)',
    p_probable  = '(%)',
    p_suspected = '(%)',
    p_not_case  = '(%)',
    p_unknown   = '(%)',
    date_adm_first = 'First',
    date_adm_last  = 'Last') %>%
  tab_spanner(
    label = html('Dates of<br>consultation/hospitalisation'),
    columns = vars(date_adm_first, date_adm_last)) %>%
  tab_spanner(
    label = html('Confirmed'),
    columns = ends_with('_confirmed')) %>%
  tab_spanner(
    label = html('Probable'),
    columns = ends_with('_probable')) %>%
  tab_spanner(
    label = html('Suspected'),
    columns = ends_with('_suspected')) %>%
  tab_spanner(
    label = html('Not a case'),
    columns = ends_with('_not_case')) %>%
  tab_spanner(
    label = html('(Unknown)'),
    columns = ends_with('_unknown')) %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = "({x})") %>%
  fmt_missing(
    columns = starts_with('date_adm_'),
    missing_text = "Unknown") %>%
  cols_align(
    align = 'right',
    columns = vars(total)) %>%
  cols_align(
    align = 'left',
    columns = starts_with('p_')) %>%
  cols_align(
    align = 'center',
    columns = starts_with('date_adm_')) %>%
  tab_style(
    style = list(
      cell_text(align = "left")),
    locations = cells_column_labels(columns = starts_with('p_'))) %>%
  summary_rows(
    groups = TRUE,
    columns = vars(total, n_confirmed, n_probable, n_suspected, n_not_case, n_unknown),
    fns = list('Country Total' = ~ sum(.)),
    missing_text = "",
    formatter = fmt_number,
    decimals = 0) %>%
  grand_summary_rows(
    columns = vars(total, n_confirmed, n_probable, n_suspected, n_not_case, n_unknown),
    fns = list('Grand Total' = ~ sum(.)),
    missing_text = "",
    formatter = fmt_number,
    decimals = 0) %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = 'bold',
    grand_summary_row.text_transform = 'uppercase',
    data_row.padding = px(1),
    row_group.padding = px(1),
    summary_row.padding = px(1),
    grand_summary_row.padding = px(1))

gtsave(gtbl_sites_ind_MSF_covid_status,
       file.path(path.local.msf.tables.oc, paste0('gtbl_sites_ind_MSF_covid_status', '_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_sites_ind_MSF_covid_status %>%
         tab_header(
           title = html(paste('SUMMARY TABLE<br>
           Counts of patients consulted/admitted to an MSF supported COvid-19 health service by Covid status as of', format(date_max_report, "%d %b %Y"))),
           subtitle = 'Sites with (*) reported aggregated data'),
       file.path(path.local.msf.tables.oc, paste0(week_report, '_SUMMARY-TABLE_MSF-sites_by_patients_Covid-status.html'))) %>%
  invisible()
```



## Patients' characteristics

### Age distribution by patinfo_sex of patients consulted/admitted

```{r gtbl-age-distribution-by-patinfo_sex-all}

dta_age_sex <- dta_linelist %>%
  select(ind_MSF_covid_status, age_in_years, age_5gp, patinfo_sex) %>%
  drop_na() %>%
  transform(patinfo_sex = factor(patinfo_sex, levels = c('F', 'M'), 
                         labels = c('Female', 'Male')))

tbl_age_distribution_by_sex_all <- dta_age_sex %>%
  group_by(patinfo_sex) %>%
  summarise(
    age_min_max = paste0(min(age_in_years, na.rm = TRUE), ' - ', max(age_in_years, na.rm = TRUE)),
    age_mean    = format(round(mean(age_in_years, na.rm = TRUE), digits = 1), nsmall = 1),
    age_median  = format(round(median(age_in_years, na.rm = TRUE), digits = 1), nsmall = 1)) %>%
  add_row(
    dta_age_sex %>%
      summarise(
        patinfo_sex = 'All',
        age_min_max = paste0(min(age_in_years, na.rm = TRUE), ' - ',
                             max(age_in_years, na.rm = TRUE)),
        age_mean    = format(round(mean(age_in_years, na.rm = TRUE), 
                                   digits = 1), nsmall = 1),
        age_median  = format(round(median(age_in_years, na.rm = TRUE), 
                                   digits = 1), nsmall = 1)))


gtbl_age_distribution_by_sex_all <- tbl_age_distribution_by_sex_all %>%
  gt() %>%
  cols_label(
    patinfo_sex = 'Sex',
    age_min_max = 'Min-Max',
    age_mean = 'Mean',
    age_median = 'Median') %>%
  tab_header(
    title = 'Age distribution of all patients consulted/admitted') %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(1),
    row_group.padding = px(1))


gtbl_age_distribution_by_sex_all
```

```{r gtbl-age-patinfo_sex-continent}
dta_age_sex <- dta_linelist %>%
  mutate(patinfo_sex = factor(patinfo_sex, levels = c('F', 'M'), labels = c('Female', 'Male')) %>% fct_explicit_na())

tbl_age_sex <- dta_age_sex %>%
  group_by(patinfo_sex) %>%
  summarise(
    age_median = median(age_in_years, na.rm = TRUE)) %>%
  ungroup() %>%
  pivot_wider(names_from = patinfo_sex, values_from = age_median) %>%
  mutate(
    continent = 'Total')

tbl_age_sex_continent <- dta_age_sex %>%
  group_by(continent, patinfo_sex) %>%
  summarise(
    age_median = median(age_in_years, na.rm = TRUE)) %>%
  ungroup() %>%
  pivot_wider(names_from = patinfo_sex, values_from = age_median)


gtbl_age_sex_continent <- tbl_age_sex_continent %>%
  add_row(tbl_age_sex) %>%
  gt(rowname_col = 'continent') %>%
  fmt_number(
    columns = 2:4,
    decimals = 0) %>%
  tab_header(
    title = 'Median age (years) of patients by continent, all patients consulted/admitted') %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2))

gtbl_age_sex_continent
```

### Sex ratio

```{r}
tbl_sex_ratio <- dta_linelist %>%
  select(continent, patinfo_sex) %>%
  transform(patinfo_sex = factor(patinfo_sex, levels = c('M', 'F'), 
                                 labels = c('Male', 'Female')) %>%
              fct_explicit_na()) %>%
  group_by(continent) %>%
  count(patinfo_sex) %>%
  pivot_wider(names_from = patinfo_sex, values_from = n) %>%
  adorn_totals() %>%
  mutate(ratio_mf = round(Male/Female, 1))

gtbl_sex_ratio <- tbl_sex_ratio %>%
  gt(rowname_col = 'continent') %>%
  cols_label(
    ratio_mf = 'Male/Female ratio') %>%
  tab_header(
    title = 'Sex ratio of all patients consulted/admitted by continent') %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2))

gtbl_sex_ratio
```

### Summary table of patients consulted/admitted by type of case


<!-- Table age (median, IQR, min-max), sex and presence of at least one morbidity, by Covid status -->
```{r tbl-all-general-characteristics, layout = "l-page"}
# DONE by Francesco
## This is a complicated table, there might be a simpler way that this one

tbl_general_by_status <- dta_linelist %>%
  select(ind_MSF_covid_status, age_in_years, age_5gp, patinfo_sex, ind_Comcond_01) %>%
  group_by(ind_MSF_covid_status) %>%
  summarise(
    age_total   = paste0(sum(!is.na(age_in_years)) , '/',  n()),
    age_min_max = paste0(min(age_in_years, na.rm = TRUE), ' - ', max(age_in_years, na.rm = TRUE)),
    age_median  = format(round(median(age_in_years, na.rm = TRUE), digits = 1), nsmall = 1),
    age_gp1     = freq_prct(age_5gp, '0-4'),
    age_gp2     = freq_prct(age_5gp, '5-14'),
    age_gp3     = freq_prct(age_5gp, '15-44'),
    age_gp4     = freq_prct(age_5gp, '45-64'),
    age_gp5     = freq_prct(age_5gp, '65+'),
    sex_total   = paste0(sum(!is.na(patinfo_sex)) , '/',  n()),
    sex_males   = sum(patinfo_sex == 'M', na.rm = TRUE),
    sex_females = sum(patinfo_sex == 'F', na.rm = TRUE),
    sex_ratio   = format(round(sum(patinfo_sex == 'M', na.rm = TRUE)/sum(patinfo_sex == 'F', na.rm = TRUE), digits = 1), nsmall = 1),
    comorbidity_total = paste0(sum(!is.na(ind_Comcond_01)) , '/',  n()),
    comorbidity_presence = freq_prct(ind_Comcond_01, 1))


tbl_general_total <- dta_linelist %>%
  select(ind_MSF_covid_status, age_in_years, age_5gp, patinfo_sex, ind_Comcond_01) %>%
  summarise(
    age_total   = paste0(sum(!is.na(age_in_years)) , '/',  n()),
    age_min_max = paste0(min(age_in_years, na.rm = TRUE), ' - ', max(age_in_years, na.rm = TRUE)),
    age_median  = format(round(median(age_in_years, na.rm = TRUE), digits = 1), nsmall = 1),
    age_gp1     = freq_prct(age_5gp, '0-4'),
    age_gp2     = freq_prct(age_5gp, '5-14'),
    age_gp3     = freq_prct(age_5gp, '15-44'),
    age_gp4     = freq_prct(age_5gp, '45-64'),
    age_gp5     = freq_prct(age_5gp, '65+'),
    sex_total   = paste0(sum(!is.na(patinfo_sex)) , '/',  n()),
    sex_males   = sum(patinfo_sex == 'M', na.rm = TRUE),
    sex_females = sum(patinfo_sex == 'F', na.rm = TRUE),
    sex_ratio   = format(round(sum(patinfo_sex == 'M', na.rm = TRUE)/sum(patinfo_sex == 'F', na.rm = TRUE), digits = 1), nsmall = 1),
    comorbidity_total = paste0(sum(!is.na(ind_Comcond_01)) , '/',  n()),
    comorbidity_presence = freq_prct(ind_Comcond_01, 1)) %>%
  mutate(
    ind_MSF_covid_status = 'Total')

tbl_general <- tbl_general_by_status %>%
  add_row(tbl_general_total) %>%
  mutate(
    ind_MSF_covid_status = factor(ind_MSF_covid_status, levels = c('Confirmed', 'Probable', 'Suspected', 'Not a case', '(Unknown)', 'Total')))

vars_age <- tibble(levels = tbl_general %>% select(starts_with('age_')) %>% names(),
                   labels = c('Totals with age information, n/N', 'Minimum - maximum', 'Median', '0 - 4, n (%)', '5 - 14, n (%)', '15 - 44, n (%)', '45 - 64, n (%)', '65+, n (%)'))

vars_sex <- tibble(levels = tbl_general %>% 
                     select(starts_with('sex_')) %>% 
                     names(),
                   labels = c('Totals with sex information, n/N', 'Men, n', 'Women, n', 'Sex ratio (M/F)'))

vars_comorbidities <- tibble(levels = tbl_general %>% select(starts_with('comorbidity_')) %>% names(),
                             labels = c('Totals with comorbidity information, n/N', 'Presence of at least one comorbidity, n (%)'))

gtbl_general <- tbl_general %>%
  gather(characteristics, values, -ind_MSF_covid_status, factor_key = TRUE) %>%
  spread(ind_MSF_covid_status, values) %>%
  mutate(type_characteristic = case_when(
    characteristics ==  'total' ~ NA_character_,
    characteristics %in% vars_age$levels ~ 'Age (in years)',
    characteristics %in% vars_sex$levels ~ 'Sex',
    characteristics %in% vars_comorbidities$levels ~ 'Comorbidities')) %>%
  mutate(
    characteristics = factor(characteristics,
                             levels = c(vars_age$levels, vars_sex$levels, vars_comorbidities$levels),
                             labels = c(vars_age$labels, vars_sex$labels, vars_comorbidities$labels))) %>%
  gt(rowname_col = "characteristics",
     groupname_col = "type_characteristic") %>%
  cols_label(
    characteristics = 'Characteristics') %>%
  cols_align(
    align = 'left',
    columns = vars(characteristics)) %>%
  cols_align(
    align = 'center',
    columns = vars(Confirmed, Probable, Suspected, `Not a case`, `(Unknown)`)) %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = 'bold',
    data_row.padding = px(2),
    row_group.padding = px(2))

gtsave(gtbl_general,
       file.path(path.local.msf.tables.oc, paste0('gtbl_general','_', week_report, '.png'))) %>%
  invisible()


gtsave(gtbl_general %>%
         tab_header(
           title = paste('Characteristics of patients consulted in in an MSF supported Covid-19 health service by continent as of', format(date_max_report, "%d %B %Y"))),
       file.path(path.local.msf.tables.oc, paste0('gtbl_general', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_general %>%
  tab_header(
    title = paste('Characteristics of patients consulted/admitted by type of case as of', format(date_max_report, "%d %B %Y")))
```

### Age/sex pyramid by continent

<!-- Age/sex pyramid by continent -->
```{r pyramid-age-all-continent, fig.cap = 'Age pyramid of patients consulted/admitted to an MSF supported Covid-19 health service by continent', fig.width = 8, fig.height = 8}
# Done by Paul

missing_age_sex <- dta_linelist %>%
  summarise(age = sum(is.na(age_in_years)), patinfo_sex = sum(is.na(patinfo_sex)))

tbl_pyramid_continent <- dta_linelist %>%
  drop_na(patinfo_sex, age_in_years) %>%
  count(continent, patinfo_sex, age_9gp) %>%
  mutate(n = if_else(patinfo_sex == "M", -n, n))

pyramid_age_sex_all_continent <- ggplot(tbl_pyramid_continent, aes(x = age_9gp, y = n, fill = patinfo_sex)) +
  facet_wrap(~continent, ncol = 2, scales = "free_x") +
  geom_col() +
  geom_hline(yintercept = 0, colour = "black") +
  geom_text(aes(label = abs(n), hjust = if_else(n >= 0, 1.1, -0.1)), colour = "white", size = 3.5) +
  coord_flip() +
  scale_fill_manual(name = NULL, values = c("#4E79A7FF", "#A0CBE8FF"), breaks = c("M", "F"), labels = c("Males", "Females")) +
  scale_y_continuous(label = abs, limits = pyramid_limits) +
  theme(legend.position = "top") +
  labs(x = "Age Group", y = "Patients consulted/admitted", caption = glue::glue("Missing Data: Age {missing_age_sex$age}, Sex {missing_age_sex$patinfo_sex}"))

ggsave(file.path(path.local.msf.graphs.oc, paste0('pyramid_age_sex_all_continent', '_', week_report, '.png')),
       plot = pyramid_age_sex_all_continent,
       width = 8,
       height = 8)


pyramid_age_sex_all_continent
```

### Comorbidities

```{r tbl-symtopms-tbl_comcond_status, layout = "l-page"}

dta_comcond <- prepare_msf_dta_comcond(dta_linelist)

dta_comcond <- dta_comcond %>%
  select(-c(Comcond_other, ind_Comcond_count, ind_Comcond_01, 
            Comcond_count, Comcond_01)) %>%
  mutate(across(c(starts_with('Comcond_'),'ind_MSF_hiv_status','ind_MSF_hypertension', 'ind_MSF_malaria', 'ind_MSF_malnutrition', 'ind_MSF_smoking', 'ind_MSF_tb_active'), function(x){na_if(x,'Unknown')}))


tbl_comcond_status <- names(dta_comcond)[4:ncol(dta_comcond)] %>%
  map_df(~{
    dta_comcond %>%
      drop_na(any_of(.x)) %>%
      tbl_prop(.x, 'ind_MSF_covid_status', drop_levels = TRUE) %>%
      rename('value'= .x) %>%
      mutate(type_comorbidity = .x)
  }) %>%
  filter(value == 'Yes') %>%
  mutate(
    type_comorbidity = factor(type_comorbidity, levels = df_labels_comcond$levels, labels = df_labels_comcond$labels))


gtbl_comcond_status <- tbl_comcond_status %>%
  gt() %>%
  cols_hide('value') %>%
  cols_move_to_start(columns = vars(type_comorbidity)) %>%
  cols_label(
    type_comorbidity = "Comorbidities/Underlying conditions",
    n_Confirmed    = 'n',
    n_Probable     = 'n',
    n_Suspected    = 'n',
    `n_Not a case` = 'n',
    `n_(Unknown)`  = 'n',
    n_Total        = 'n',
    N_Confirmed    = 'N',
    N_Probable     = 'N',
    N_Suspected    = 'N',
    `N_Not a case` = 'N',
    `N_(Unknown)`  = 'N',
    N_Total        = 'N',
    p_Confirmed    = '(%)',
    p_Probable     = '(%)',
    p_Suspected    = '(%)',
    `p_Not a case` = '(%)',
    `p_(Unknown)`  = '(%)',
    p_Total        = '(%)') %>%
  cols_align(
    align = 'left',
    columns = vars(type_comorbidity)) %>%
  cols_align(
    align = 'right',
    columns = starts_with(c('n_', '.p_'))) %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = "({x})") %>%
  fmt_missing(
    columns = starts_with(c('n_', 'p_')),
    missing_text = "---") %>%
  tab_spanner(
    label = 'Confirmed',
    columns = ends_with('Confirmed')) %>%
  tab_spanner(
    label = 'Probable',
    columns = ends_with('Probable')) %>%
  tab_spanner(
    label = 'Suspected',
    columns = ends_with('Suspected')) %>%
  tab_spanner(
    label = 'Not a case',
    columns = ends_with('Not a case')) %>%
  tab_spanner(
    label = 'Unknown',
    columns = ends_with('(Unknown)')) %>%
  tab_spanner(
    label = 'All',
    columns = ends_with('Total')) %>%
  tab_style(
    style = list(
      cell_text(align = "right")),
    locations = cells_column_labels(columns = starts_with(c('n_', 'p_')))) %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2))

gtsave(gtbl_comcond_status,
       file.path(path.local.msf.tables.oc, paste0('gtbl_comcond_status','_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_comcond_status,
       file.path(path.local.msf.tables.oc, paste0('gtbl_comcond_status', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_comcond_status %>%
  tab_header(
    title = paste('Frequency and proportion of comorbidities among patients an MSF supported Covid-19 health service by Covid status as of', format(date_max_report, '%d %B %Y')))
```

### Symptoms at consultation/admission

<!-- Table - Frequency and proportion of symptoms by Covid status -->
```{r tbl-symtopms-all, layout = "l-page"}
# DONE by Francesco

dta_sympt <- dta_linelist %>%
  select(ind_MSF_covid_status, tidyselect::vars_select(names(dta_linelist), starts_with('MSF_symptom_') & !ends_with('_date_onset')))

names_sympt <- grep('^MSF_symptom_', names(dta_sympt),
                    perl = TRUE, value = TRUE) %>% sort()

names(names_sympt) <- names_sympt %>%
  gsub('^MSF_symptom_','', .) %>%
  gsub('_', ' ', .) %>%
  gsub('anosmia', 'loss of smell', .) %>%
  gsub('loss taste', 'loss of taste', .) %>%
  gsub('sorethoat', 'sorethroat', .) %>%
  stringr::str_to_sentence()

dta_sympt <- rename(dta_sympt, all_of(names_sympt))


tbl_sympt_status <- names(dta_sympt)[2:ncol(dta_sympt)] %>%
  map_df(~{
    dta_sympt %>%
      drop_na(any_of(.x)) %>%
      tbl_prop(.x, 'ind_MSF_covid_status', drop_levels = FALSE) %>%
      rename('value'= .x) %>%
      mutate(
        type_symptom = .x)
  }) %>%
  filter(value == 'Yes')


gtbl_sympt_all <- tbl_sympt_status %>%
  gt() %>%
  cols_hide('value') %>%
  cols_move_to_start(columns = vars(type_symptom)) %>%
  cols_label(
    type_symptom = 'Signes/Symptoms',
    n_Confirmed    = 'n',
    n_Probable     = 'n',
    n_Suspected    = 'n',
    `n_Not a case` = 'n',
    `n_(Unknown)`  = 'n',
    n_Total        = 'n',
    N_Confirmed    = 'N',
    N_Probable     = 'N',
    N_Suspected    = 'N',
    `N_Not a case` = 'N',
    `N_(Unknown)`  = 'N',
    N_Total        = 'N',
    p_Confirmed    = '(%)',
    p_Probable     = '(%)',
    p_Suspected    = '(%)',
    `p_Not a case` = '(%)',
    `p_(Unknown)`  = '(%)',
    p_Total        = '(%)') %>%
  tab_spanner(
    label = "Confirmed",
    columns = ends_with('_Confirmed')) %>%
  tab_spanner(
    label = "Probable",
    columns = ends_with('_Probable')) %>%
  tab_spanner(
    label = "Suspected",
    columns = ends_with('_Suspected')) %>%
  tab_spanner(
    label = "Not a case",
    columns = ends_with('_Not a case')) %>%
  tab_spanner(
    label = "(Unknown)",
    columns = ends_with('_(Unknown)')) %>%
  tab_spanner(
    label = "Total",
    columns = ends_with('_Total')) %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = "({x})") %>%
  fmt_missing(
    columns = starts_with('p_'),
    missing_text = "---") %>%
  cols_align(
    align = 'left',
    columns = vars(type_symptom)) %>%
  cols_align(
    align = 'right',
    columns = starts_with(c('n_', 'p_'))) %>%
  tab_style(
    style = list(
      cell_text(align = "right")),
    locations = cells_column_labels(columns = starts_with(c('n_', 'p_')))) %>%
  cols_width(
    vars(type_symptom) ~ px(140),
    starts_with('n_') ~ px(55),
    starts_with("p_") ~ px(50)) %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = 'bold',
    data_row.padding = px(2),
    row_group.padding = px(2))

gtsave(gtbl_sympt_all,
       file.path(path.local.msf.tables.oc, paste0('gtbl_sympt_all','_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_sympt_all,
       file.path(path.local.msf.tables.oc, paste0('gtbl_sympt_all', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_sympt_all %>%
  tab_header(
    title = paste('Frequency and proportion of symptoms among patients an MSF supported Covid-19 health service by Covid status as of', format(date_max_report, '%d %B %Y')))
```

### MSF_severity by Covid status

<!-- Table - Frequency and proportion of patients wit MSF_severity -->
```{r}
# by Francesco

dta_MSF_severity <- dta_linelist %>%
  select(ind_MSF_covid_status, MSF_severity) %>%
  mutate(
    MSF_severity = factor(MSF_severity, levels = c('Mild', 'Moderate', 'Severe', 'Critical'))
  )

tbl_MSF_severity_status <- dta_MSF_severity %>%
  drop_na(MSF_severity) %>%
  tbl_prop('MSF_severity', 'ind_MSF_covid_status', drop_levels = FALSE)


gtbl_MSF_severity_status <- tbl_MSF_severity_status %>%
  gt(rowname_col = "MSF_severity") %>%
  cols_hide(
    columns = starts_with('N_', ignore.case = FALSE)) %>%
  cols_label(
    MSF_severity = "Level of MSF_severity",
    n_Confirmed    = 'n',
    n_Probable     = 'n',
    n_Suspected    = 'n',
    `n_Not a case` = 'n',
    `n_(Unknown)`  = 'n',
    n_Total        = 'n',
    p_Confirmed    = '(%)',
    p_Probable     = '(%)',
    p_Suspected    = '(%)',
    `p_Not a case` = '(%)',
    `p_(Unknown)`  = '(%)',
    p_Total        = '(%)') %>%
  cols_align(
    align = 'left',
    columns = vars(MSF_severity)) %>%
  cols_align(
    align = 'right',
    columns = starts_with(c('n_', 'p_'))) %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = "({x})") %>%
  fmt_missing(
    columns = starts_with(c('n_', 'p_')),
    missing_text = "---") %>%
  tab_spanner(
    label = 'Confirmed',
    columns = ends_with('Confirmed')) %>%
  tab_spanner(
    label = 'Probable',
    columns = ends_with('Probable')) %>%
  tab_spanner(
    label = 'Suspected',
    columns = ends_with('Suspected')) %>%
  tab_spanner(
    label = 'Not a case',
    columns = ends_with('Not a case')) %>%
  tab_spanner(
    label = 'Unknown',
    columns = ends_with('(Unknown)')) %>%
  tab_spanner(
    label = 'All',
    columns = ends_with('Total')) %>%
  summary_rows(
    groups = NULL,
    columns = starts_with('n_', ignore.case = FALSE),
    fns = list('Total' = ~ sum(.)),
    missing_text = "",
    formatter = fmt_number,
    decimals = 0) %>%
  #  tab_footnote(
  #    footnote = paste0("Missing values = ", sum(is.na(dta_MSF_severity$MSF_severity))),
  #    locations = cells_summary(
  #      groups = NULL,
  #      columns = vars(MSF_severity),
  #      rows = 1)) %>%
  tab_style(
    style = list(
      cell_text(align = "right")),
    locations = cells_column_labels(columns = starts_with(c('n_', 'p_')))) %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2))

gtsave(gtbl_MSF_severity_status,
       file.path(path.local.msf.tables.oc, paste0('gtbl_MSF_severity_status','_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_MSF_severity_status,
       file.path(path.local.msf.tables.oc, paste0('gtbl_MSF_severity_status', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_MSF_severity_status %>%
  tab_header(
    title = paste('Frequency and proportion of level of MSF_severity among patients consulted at an MSF supported Covid-19 health service as of', format(date_max_report, '%d %B %Y')))
```

### MSF_severity of confirmed cases by continent

```{r}
dta_MSF_severity_confirmed_continent <- dta_linelist %>%
  select(continent, ind_MSF_covid_status, MSF_severity) %>%
  filter(ind_MSF_covid_status == 'Confirmed') %>%
  mutate(
    MSF_severity = factor(MSF_severity, levels = c('Mild', 'Moderate', 'Severe', 'Critical'))
  )


tbl_MSF_severity_confirmed_continent <- dta_MSF_severity_confirmed_continent %>%
  drop_na(MSF_severity) %>%
  tbl_prop('MSF_severity', 'continent', drop_levels = FALSE)

gtbl_MSF_severity_confirmed_continent <- tbl_MSF_severity_confirmed_continent %>%
  gt(rowname_col = "MSF_severity") %>%
  cols_hide(
    columns = starts_with('N_', ignore.case = FALSE)) %>%
  cols_label(
    n_Africa   = 'n',
    n_Americas = 'n',
    n_Asia     = 'n',
    n_Europe   = 'n',
    n_Total    = 'n',
    p_Africa   = '(%)',
    p_Americas = '(%)',
    p_Asia     = '(%)',
    p_Europe   = '(%)',
    p_Total    = '(%)') %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = "({x})") %>%
  fmt_missing(
    columns = starts_with(c('n_', 'p_')),
    missing_text = "---") %>%
  cols_align(
    align = 'right',
    columns = starts_with(c('n_', 'p_'))) %>%
  tab_spanner(
    label = 'Africa',
    columns = ends_with('_Africa')) %>%
  tab_spanner(
    label = 'Americas',
    columns = ends_with('_Americas')) %>%
  tab_spanner(
    label = 'Asia',
    columns = ends_with('_Asia')) %>%
  tab_spanner(
    label = 'Europe',
    columns = ends_with('_Europe')) %>%
  tab_spanner(
    label = 'Total',
    columns = ends_with('_Total')) %>%
  summary_rows(
    groups = NULL,
    columns = starts_with('n_', ignore.case = FALSE),
    fns = list('Total' = ~ sum(.)),
    missing_text = "",
    formatter = fmt_number,
    decimals = 0) %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = 'bold',
    data_row.padding = px(2),
    row_group.padding = px(2))

gtsave(gtbl_MSF_severity_confirmed_continent,
       file.path(path.local.msf.tables.oc, paste0('gtbl_MSF_severity_confirmed_continent','_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_MSF_severity_confirmed_continent,
       file.path(path.local.msf.tables.oc, paste0('gtbl_MSF_severity_confirmed_continent', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_MSF_severity_confirmed_continent %>%
  tab_header(
    title = paste('Frequency and proportion of level of MSF_severity among Covid confirmed patients consulted at an MSF supported Covid-19 health service as of', format(date_max_report, '%d %B %Y')))
```

### Case-fatality by Covid status

<!-- Table Case-fatality risk by Covid status and by continent -->

```{r tbl-cfr-status}
# DONE by Francesco

dta_cfr_status <- dta_linelist %>%
  select(continent, ind_MSF_covid_status, merge_admit, ind_outcome_patcourse_status) %>%
  filter(ind_outcome_patcourse_status %in% c('Cured', 'Died','Sent back home'),
       merge_admit == "Yes")

tbl_cfr_status_continent <- levels(dta_cfr_status$ind_MSF_covid_status) %>%
  map_df(~{
    dta_cfr_status %>%
      filter(ind_MSF_covid_status == .x) %>%
      tbl_prop('ind_outcome_patcourse_status', 'continent', drop_levels = FALSE) %>%
      mutate(
        ind_MSF_covid_status = .x)
  }) %>%
  filter(ind_outcome_patcourse_status == 'Died')


gtbl_cfr_status_continent <- tbl_cfr_status_continent %>%
  gt() %>%
  cols_hide('ind_outcome_patcourse_status') %>%
  cols_move_to_start(columns = vars(ind_MSF_covid_status)) %>%
  cols_label(
    ind_MSF_covid_status = 'Covid status',
    n_Africa   = 'n',
    n_Americas = 'n',
    n_Asia     = 'n',
    n_Europe   = 'n',
    n_Total    = 'n',
    N_Africa   = 'N',
    N_Americas = 'N',
    N_Asia     = 'N',
    N_Europe   = 'N',
    N_Total    = 'N',
    p_Africa   = '(%)',
    p_Americas = '(%)',
    p_Asia     = '(%)',
    p_Europe   = '(%)',
    p_Total    = '(%)') %>%
  tab_spanner(
    label = 'Africa',
    columns = ends_with('_Africa')) %>%
  tab_spanner(
    label = 'Americas',
    columns = ends_with('_Americas')) %>%
  tab_spanner(
    label = 'Asia',
    columns = ends_with('_Asia')) %>%
  tab_spanner(
    label = 'Europe',
    columns = ends_with('_Europe')) %>%
  tab_spanner(
    label = 'Total',
    columns = ends_with('_Total')) %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = '({x})') %>%
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  fmt_missing(
    columns = starts_with('p_'),
    missing_text = "---") %>%
  cols_align(
    align = 'left',
    columns = vars(ind_MSF_covid_status)) %>%
  cols_align(
    align = 'right',
    columns = starts_with('n_')) %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2))

gtsave(gtbl_cfr_status_continent,
       file.path(path.local.msf.tables.oc, paste0('gtbl_cfr_status_continent','_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_cfr_status_continent,
       file.path(path.local.msf.tables.oc, paste0('gtbl_cfr_status_continent','_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_cfr_status_continent %>%
  tab_header(
    title = 'Case-fatality risk by Covid status and continent among admitted patients an MSF supported Covid-19 health service',
    subtitle = 'In the denominator only patients with outcome as recovered, sent back home or died'
  )


# CFR text
cfr_confirmed_probable_suspected <- dta_linelist %>%
  select(ind_MSF_covid_status, ind_outcome_patcourse_status) %>%
  filter(ind_MSF_covid_status %in% c('Confirmed', 'Probable', 'Suspected')) %>%
  filter(ind_outcome_patcourse_status %in% c('Cured', 'Died', 'Sent back home'), ) %>%
  summarise(cfr = mean(ind_outcome_patcourse_status == 'Died')) %>%
  pull()
```

---

# Covid-19 confirmed patients

## Weekly number of confirmed patients consulted/admitted

```{r, eval = FALSE}

dta_linelist_confirmed <- dta_linelist %>%
  filter(ind_MSF_covid_status == 'Confirmed')

tbl_epicurve_consultation_continent <- dta_linelist_confirmed %>%
  filter(merge_admit != "Unknown") %>%
  count(continent, epi_week_consultation, merge_admit) %>%
  drop_na() %>%
  mutate(merge_admit = recode(merge_admit, "Yes" = "Admitted", "No" = "Not Admitted"))

tbl_admit_prop_continent <- dta_linelist_confirmed %>%
  filter(merge_admit != "Unknown") %>%
  drop_na(epi_week_consultation) %>%
  group_by(continent, epi_week_consultation) %>%
  summarise(
    n = n(),
    total = sum(merge_admit %in% c("Yes", "No")),
    admitted = sum(merge_admit == "Yes"),
    admit_prop = admitted / total
  ) %>%
  ungroup()

n_max <- max(tbl_admit_prop_continent$n, na.rm = TRUE)
cfr_max <- rounder(max(tbl_admit_prop_continent$admit_prop, na.rm = TRUE), .1)
cfr_max <- ifelse(cfr_max > 1, 1, cfr_max)
scaling_factor <- cfr_max / n_max # sets the y limit of CFR rounded up to nearest 10% above cfr max

missing_consultation <- sum(is.na(dta_linelist_confirmed$epi_week_consultation))
missing_admission    <- sum(dta_linelist_confirmed$merge_admit == "Unknown")

hist_epicurve_consult_admit_confirmed_continent <- tbl_epicurve_consultation_continent %>%
  ggplot(aes(epi_week_consultation, n)) +
  facet_wrap(~continent) +
  geom_col(aes(fill = merge_admit)) +
  geom_line(data = tbl_admit_prop_continent,
            aes(y = admit_prop / scaling_factor, colour = "Admitted %"),
            key_glyph = "timeseries", size = 0.5) +
  ggthemes::scale_fill_tableau(name = NULL, palette = "Tableau 20") +
  scale_x_date(name = "Week of Consultation", date_breaks = "3 week", date_labels = "%V",
               sec.axis = ggplot2::sec_axis(trans = ~ .), expand = expansion(mult = c(0.01, 0.01))) +
  scale_y_continuous(sec.axis = ggplot2::sec_axis(~ . * scaling_factor, name = "Admitted", labels = scales::percent_format(accuracy = 1)),
                     expand = expansion(mult = c(0, 0.02))) +
  labs(y = "Patients", colour = NULL, caption = glue::glue("Missing data: Consultation Date {missing_consultation}, Admission: {missing_admission}")) +
  theme_light() +
  theme(legend.position = "top", panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank())

ggsave(file.path(path.local.msf.graphs.oc, paste0('hist_epicurve_consult_admit_confirmed_continent', '_', week_report, '.png')),
       plot = hist_epicurve_consult_admit_confirmed_continent,
       width = 6,
       height = 4,
       scale = 1.1,
       dpi = 320)


hist_epicurve_consult_admit_confirmed_continent
```

## Age and sex distribution of Covid confirmed patients

```{r gtbl-age-distribution-by-sex-confirmed-only}

dta_age_sex_confirmed <- dta_linelist %>%
  select(ind_MSF_covid_status, age_in_years, age_5gp, patinfo_sex) %>%
  filter(ind_MSF_covid_status == 'Confirmed') %>%
  drop_na() %>%
  transform(patinfo_sex = factor(patinfo_sex, levels = c('F', 'M'), labels = c('Female', 'Male')))

tbl_age_distribution_by_sex_confirmed_only <- dta_age_sex_confirmed %>%
  group_by(patinfo_sex) %>%
  summarise(
    age_min_max = paste0(min(age_in_years, na.rm = TRUE), ' - ', max(age_in_years, na.rm = TRUE)),
    age_mean    = round(mean(age_in_years, na.rm = TRUE), digits = 1),
    age_median  = round(median(age_in_years, na.rm = TRUE), digits = 1)) %>%
  add_row(
    dta_age_sex_confirmed %>%
      summarise(
        patinfo_sex = 'All',
        age_min_max = paste0(min(age_in_years, na.rm = TRUE), ' - ', max(age_in_years, na.rm = TRUE)),
        age_mean    = round(mean(age_in_years, na.rm = TRUE), digits = 1),
        age_median  = round(median(age_in_years, na.rm = TRUE), digits = 1)))

gtbl_age_distribution_by_sex_confirmed_only <- tbl_age_distribution_by_sex_confirmed_only %>%
  gt() %>%
  cols_label(
    patinfo_sex = 'Sex',
    age_min_max = 'Min-Max',
    age_mean = 'Mean',
    age_median = 'Median') %>%
  tab_header(
    title = 'Covid-confirmed patients consulted/admitted')

gtbl_age_distribution_by_sex_confirmed_only

```


<!-- Age/sex pyramid -->
```{r pyramid-age-sex-confirmed, fig.cap = 'Age pyramid of Covid-confirmed patients consulted/admitted to an MSF supported health service', fig.width = 8, fig.height = 8}
# Done by Paul

missing_age_sex <- dta_linelist %>% filter(ind_MSF_covid_status == "Confirmed") %>% summarise(age = sum(is.na(age_in_years)), patinfo_sex = sum(is.na(patinfo_sex)))

tbl_pyramid_confirmed <- dta_linelist %>%
  drop_na(patinfo_sex, age_in_years) %>%
  filter(ind_MSF_covid_status %in% c("Confirmed")) %>%
  count(patinfo_sex, age_9gp) %>%
  mutate(n = if_else(patinfo_sex == "M", -n, n))

pyramid_age_sex_confirmed <- ggplot(tbl_pyramid_confirmed, aes(x = age_9gp, y = n, fill = patinfo_sex)) +
  geom_col() +
  geom_hline(yintercept = 0, colour = "black") +
  #geom_text(aes(label = abs(n), hjust = if_else(n >= 0, 1.1, -0.1)), colour = "white", size = 4) +
  coord_flip() +
  scale_fill_manual(name = NULL, values = c("#4E79A7FF", "#A0CBE8FF"), breaks = c("M", "F"), labels = c("Males", "Females")) +
  scale_y_continuous(limits = pyramid_limits, breaks = pyramid_brks, label = abs) +
  theme(legend.position = "top") +
  labs(x = "Age Group", y = "Confirmed Patients", caption = glue::glue("Missing Data: Age {missing_age_sex$age}, Sex {missing_age_sex$patinfo_sex}"))  +
  guides(y.sec = guide_axis_label_trans(~paste(.x))) # show age group labels on both axis

ggsave(file.path(path.local.msf.graphs.oc, paste0('pyramid_age_sex_confirmed', '_', week_report, '.png')),
       plot = pyramid_age_sex_confirmed,
       width = 8,
       height = 6)


pyramid_age_sex_confirmed

```

## Age and sex distribution of Covid confirmed patients by continent

<!-- Age/sex pyramid by continent -->

```{r pyramid-age-sex-confirmed-continent, fig.cap = 'Age pyramid of patients consulted/admitted to an MSF supported Covid-19 health service by continent', fig.width = 8, fig.height = 8}
# Done by Paul

missing_age_sex <- dta_linelist %>% filter(ind_MSF_covid_status == "Confirmed") %>% summarise(age = sum(is.na(age_in_years)), patinfo_sex = sum(is.na(patinfo_sex)))

tbl_pyramid_confirmed_continent <- dta_linelist %>%
  drop_na(patinfo_sex, age_in_years) %>%
  filter(ind_MSF_covid_status %in% c("Confirmed")) %>%
  count(continent, patinfo_sex, age_9gp) %>%
  mutate(n = if_else(patinfo_sex == "M", -n, n))

pyramid_age_sex_confirmed_continent <- ggplot(tbl_pyramid_confirmed_continent, aes(x = age_9gp, y = n, fill = patinfo_sex)) +
  facet_wrap(~continent, ncol = 2, scales = "free_x") +
  geom_col() +
  geom_hline(yintercept = 0, colour = "black") +
  geom_text(aes(label = abs(n), hjust = if_else(n >= 0, 1.1, -0.1)), colour = "white", size = 3.5) +
  coord_flip() +
  scale_fill_manual(name = NULL, values = c("#4E79A7FF", "#A0CBE8FF"), breaks = c("M", "F"), labels = c("Males", "Females")) +
  scale_y_continuous(label = abs, limits = pyramid_limits) +
  theme(legend.position = "top") +
  labs(x = "Age Group", y = "Confirmed Patients", caption = glue::glue("Missing Data: Age {missing_age_sex$age}, Sex {missing_age_sex$patinfo_sex}"))

ggsave(file.path(path.local.msf.graphs.oc, paste0('pyramid_age_sex_confirmed_continent', '_', week_report, '.png')),
       plot = pyramid_age_sex_confirmed_continent,
       width = 8,
       height = 8)


pyramid_age_sex_confirmed_continent
```

## Patients care

<!--
Table Patient’s care 1
- Admitted to a hospital/inpatient department,
- received oxygen,
- admitted to ICU,
- supported with ventilator,
- supported with ECMO # removed on 25 November 2020
-->

```{r tbl-care-procedures}
# DONE by Francesco

dta_care <- dta_linelist %>%
  filter(ind_MSF_covid_status %in% c('Confirmed')) %>%
  select(continent, country, starts_with('merge_'))


tbl_care_admitted <- tbl_prop(dta_care, 'merge_admit', 'continent') %>%
  mutate(name_var = 'Admitted')

tbl_care_oxygen <- tbl_prop(dta_care, 'merge_oxygen', 'continent') %>%
  mutate(name_var = 'Received oxygen')

tbl_care_icu <- tbl_prop(dta_care, 'merge_icu', 'continent') %>%
  mutate(name_var = 'Admitted to the Intensive Care Unit')

tbl_care_vent <- tbl_prop(dta_care, 'merge_vent', 'continent') %>%
  mutate(name_var = 'Supported by a ventilator')


tbl_care <- tbl_care_admitted %>% rename(values_var = merge_admit) %>%
  add_row(tbl_care_oxygen %>% rename(values_var = merge_oxygen)) %>%
  add_row(tbl_care_icu %>% rename(values_var = merge_icu)) %>%
  add_row(tbl_care_vent %>% rename(values_var = merge_vent))


gtbl_care <- tbl_care %>%
  gt(rowname_col = "values_var",
     groupname_col = "name_var") %>%
  cols_hide(
    columns = starts_with('N_', ignore.case = FALSE)) %>%
  cols_label(
    n_Africa   = 'n',
    n_Americas = 'n',
    n_Asia     = 'n',
    n_Europe   = 'n',
    n_Total    = 'n',
    p_Africa   = '(%)',
    p_Americas = '(%)',
    p_Asia     = '(%)',
    p_Europe   = '(%)',
    p_Total    = '(%)') %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = "({x})") %>%
  tab_spanner(
    label = 'Africa',
    columns = ends_with('_Africa')) %>%
  tab_spanner(
    label = 'Americas',
    columns = ends_with('_Americas')) %>%
  tab_spanner(
    label = 'Asia',
    columns = ends_with('_Asia')) %>%
  tab_spanner(
    label = 'Europe',
    columns = ends_with('_Europe')) %>%
  tab_spanner(
    label = 'Total',
    columns = ends_with('_Total')) %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = 'bold',
    data_row.padding = px(2),
    row_group.padding = px(2))

gtsave(gtbl_care,
       file.path(path.local.msf.tables.oc, paste0('gtbl_care', '_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_care,
       file.path(path.local.msf.tables.oc, paste0('gtbl_care', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_care %>%
  tab_header(
    title = "Major care procedures to Covid confirmed patient's in MSF heatlth facilities, by location")
```

## Age and sex of patients by outcome

```{r tbl-age-sex-outcome}
# DONE by Francesco

dta_age_sex_outcome <- dta_linelist %>%
  select(site_name, ind_MSF_covid_status, ind_outcome_patcourse_status,
         patinfo_sex = patinfo_sex, age_in_years) %>%
  filter(ind_MSF_covid_status %in% c('Confirmed', 'Probable', 'Suspected'))

tbl_age_sex_outcome <- dta_age_sex_outcome %>%
  group_by(ind_outcome_patcourse_status) %>%
  summarise(
    age_median = median(age_in_years, na.rm = TRUE),
    age__prcl  = glue('[{quantile(age_in_years, na.rm = TRUE)[[2]]} - {quantile(age_in_years, na.rm = TRUE)[[4]]}]'),
    age_range  = glue('[{min(age_in_years, na.rm = TRUE)} - {max(age_in_years, na.rm = TRUE)}]'),
    sex_males   = sum(patinfo_sex == 'M', na.rm = TRUE),
    sex_females = sum(patinfo_sex == 'F', na.rm = TRUE),
    sex_ratio   = sex_males/sex_females)

tbl_age_sex_outcome$sex_ratio[is.infinite(tbl_age_sex_outcome$sex_ratio)] <- NA

gtbl_age_sex_outcome <- tbl_age_sex_outcome %>%
  gt() %>%
  cols_label(
    ind_outcome_patcourse_status = 'Status',
    age_median = 'Median',
    age__prcl = '[25% - 75%]',
    age_range = '[min - max]',
    sex_males = 'Males',
    sex_females = 'Females',
    sex_ratio = 'M/F ratio') %>%
  fmt_number(
    columns = vars(sex_ratio),
    decimals = 1) %>%
  fmt_missing(
    columns = vars(sex_ratio),
    missing_text = '---') %>%
  tab_spanner(
    label = 'Age (years)',
    columns = starts_with('age_')) %>%
  tab_spanner(
    label = 'Sex',
    columns = starts_with('sex_')) %>%
  cols_align(
    align = 'left',
    columns = vars(ind_outcome_patcourse_status)) %>%
  cols_align(
    align = 'right',
    columns = starts_with('sex_')) %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(2))

gtsave(gtbl_age_sex_outcome,
       file.path(path.local.msf.tables.oc, paste0('gtbl_age_sex_outcome', '_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_age_sex_outcome,
       file.path(path.local.msf.tables.oc, paste0('gtbl_age_sex_outcome', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_age_sex_outcome %>%
  tab_header(
    title = 'Age and sex ditribution by type of patient outcome among confirmed, probable and suspected Covid cases')
```

## Case fatality by age, sex and number of comorbidities

<!-- CFR by age, sex and number of comorbidities -->

```{r tbl-cfr-age-sex-comorbidities}
# DONE by Francesco
# function tbl_prop in utils_vis.R

dta_cfr_age <- dta_linelist %>%
  mutate(age_5gp = fct_explicit_na(age_5gp)) %>% 
  filter(ind_outcome_patcourse_status %in% c('Cured', 'Died','Sent back home'),
       merge_admit == "Yes")

tbl_cfr_age <- levels(dta_cfr_age$age_5gp) %>%
  map_df(~{
    dta_cfr_age %>%
      filter(age_5gp == .x) %>%
      tbl_prop('ind_outcome_patcourse_status', 'continent', drop_levels = FALSE) %>%
      mutate(
        age_5gp = .x)
  }) %>%
  filter(ind_outcome_patcourse_status == 'Died')


dta_cfr_sex <- dta_linelist %>%
  filter(ind_outcome_patcourse_status %in% c('Cured', 'Died','Sent back home'),
       merge_admit == "Yes") %>% 
  mutate(sex = factor(patinfo_sex, levels = c('F', 'M'), labels = c('Females', 'Males'))) %>%
  mutate(sex = fct_explicit_na(sex))

tbl_cfr_sex <- levels(dta_cfr_sex$sex) %>%
  map_df(~{
    dta_cfr_sex %>%
      filter(sex == .x) %>%
      tbl_prop('ind_outcome_patcourse_status', 'continent', drop_levels = FALSE) %>%
      mutate(
        sex = .x)
  }) %>%
  filter(ind_outcome_patcourse_status == 'Died')

dta_cfr_comorbidities <- dta_linelist %>%
  filter(ind_outcome_patcourse_status %in% c('Cured', 'Died','Sent back home'),
       merge_admit == "Yes") %>% 
  mutate(ind_Comcond_count = as.factor(ind_Comcond_count)) %>%
  mutate(ind_Comcond_count = fct_explicit_na(ind_Comcond_count))


tbl_cfr_comorbidities <- levels(dta_cfr_comorbidities$ind_Comcond_count) %>%
  map_df(~{
    dta_cfr_comorbidities %>%
      filter(ind_Comcond_count == .x) %>%
      tbl_prop('ind_outcome_patcourse_status', 'continent', drop_levels = FALSE) %>%
      mutate(
        ind_Comcond_count = .x)
  }) %>%
  filter(ind_outcome_patcourse_status == 'Died') %>%
  arrange(ind_Comcond_count)


tbl_cfr_age_sex_comorbidities <- tbl_cfr_age %>%
  rename(var = age_5gp) %>%
  mutate(label_var = 'Age (years)') %>%
  add_row(tbl_cfr_sex %>%
            rename(var = sex) %>%
            mutate(label_var = 'Sex')) %>%
  add_row(tbl_cfr_comorbidities %>%
            rename(var = ind_Comcond_count) %>%
            mutate(label_var = 'Number of comorbidities'))


gtbl_cfr_age_sex_comorbidities <- tbl_cfr_age_sex_comorbidities %>%
  gt(rowname_col = "var",
     groupname_col = "label_var") %>%
  cols_hide(
    columns = vars('ind_outcome_patcourse_status')) %>%
  cols_label(
    n_Africa   = 'n',
    n_Americas = 'n',
    n_Asia     = 'n',
    n_Europe   = 'n',
    n_Total    = 'n',
    N_Africa   = 'N',
    N_Americas = 'N',
    N_Asia     = 'N',
    N_Europe   = 'N',
    N_Total    = 'N',
    p_Africa   = '(%)',
    p_Americas = '(%)',
    p_Asia     = '(%)',
    p_Europe   = '(%)',
    p_Total    = '(%)') %>%
  tab_spanner(
    label = 'Africa',
    columns = ends_with('_Africa')) %>%
  tab_spanner(
    label = 'Americas',
    columns = ends_with('_Americas')) %>%
  tab_spanner(
    label = 'Asia',
    columns = ends_with('_Asia')) %>%
  tab_spanner(
    label = 'Europe',
    columns = ends_with('_Europe')) %>%
  tab_spanner(
    label = 'Total',
    columns = ends_with('_Total')) %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = '({x})') %>%
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  fmt_missing(
    columns = starts_with('p_'),
    missing_text = "---") %>%
  cols_align(
    align = 'left',
    columns = vars(var)) %>%
  cols_align(
    align = 'right',
    columns = starts_with('n_')) %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2))

gtsave(gtbl_cfr_age_sex_comorbidities,
       file.path(path.local.msf.tables.oc, paste0('gtbl_cfr_age_sex_comorbidities', '_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_cfr_age_sex_comorbidities,
       file.path(path.local.msf.tables.oc, paste0('gtbl_cfr_age_sex_comorbidities', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_cfr_age_sex_comorbidities %>%
  tab_header(
    title = 'Case fatality risk by age, sex and number of comorbidities of admitted patients'
  )


# Numbers for editing report
cfr_confirmed_no_comorbidities <- tbl_cfr_comorbidities %>%
  filter(ind_Comcond_count == 0) %>%
  pull(p_Total)

cfr_confirmed_3_comorbidities <- tbl_cfr_comorbidities %>%
  filter(ind_Comcond_count == 3) %>%
  pull(p_Total)

```


<!--
Table Outcome 2 - Count of cured, deaths, transferred… by Admitted/Not Admitted
This table is NOT displayed any more
-->
```{r, eval = FALSE}
# DONE by Francesco

dta_outcome_admission <- dta_linelist %>%
  select(ind_MSF_covid_status, merge_admit, tidyselect::vars_select(names(dta_linelist), starts_with('outcome_'),                                                             starts_with('ind_outcome_')))

tbl_outcome_admission <- dta_outcome_admission %>%
  filter(ind_MSF_covid_status %in% c('Confirmed', 'Probable', 'Suspected')) %>%
  count(merge_admit, ind_outcome_patcourse_status, .drop = FALSE) %>%
  pivot_wider(names_from = merge_admit, values_from = n, values_fill = list(n = 0)) %>%
  adorn_totals('col') %>%
  mutate(
    prct_yes = round(Yes / sum(Yes) * 100, 1),
    prct_no  = round(No / sum(No) * 100, 1),
    prct_ukn = round(`(Unknown)` / sum(`(Unknown)`) * 100, 1),
    prct_tot = round(Total / sum(Total) * 100, 1)) %>%
  adorn_totals('row')

gtbl_outcome_admission <- tbl_outcome_admission %>%
  gt() %>%
  cols_label(
    ind_outcome_patcourse_status = 'Status',
    Yes     = 'N',
    No      = 'N',
    `(Unknown)` = 'N',
    Total   = 'N',
    prct_yes = '(%)',
    prct_no  = '(%)',
    prct_ukn = '(%)',
    prct_tot = '(%)') %>%
  tab_spanner(
    label = "Admitted",
    columns = vars(Yes, prct_yes)) %>%
  tab_spanner(
    label = "Non admitted",
    columns = vars(No, prct_no)) %>%
  tab_spanner(
    label = "Unknown",
    columns = vars(`(Unknown)`, prct_ukn)) %>%
  tab_spanner(
    label = "Total",
    columns = vars(Total, prct_tot)) %>%
  fmt_missing(
    columns = starts_with('prct_'),
    missing_text = "--") %>%
  text_transform(
    locations = cells_body(
      columns = starts_with('prct_')),
    fn = function(x) {
      x = paste0('(',x,')')}) %>%
  text_transform(
    locations = cells_body(
      columns = starts_with('prct_'),
      rows = ind_outcome_patcourse_status == 'Total'),
    fn = function(x) {
      x = '---'}) %>%
  cols_align(
    align = 'right',
    columns = vars(Yes, No, `(Unknown)`, Total)) %>%
  cols_align(
    align = 'right',
    columns = starts_with('prct_')) %>%
  tab_style(
    style = list(
      cell_text(align = "right")),
    locations = cells_column_labels(columns = vars(Yes, No, `(Unknown)`, Total))) %>%
  tab_style(
    style = list(
      cell_text(align = "right")),
    locations = cells_column_labels(columns = starts_with('prct_'))) %>%
  cols_width(
    vars(ind_outcome_patcourse_status) ~ px(210),
    vars(Yes, No, `(Unknown)` , Total) ~ px(70),
    starts_with("prct_") ~ px(50)) %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(2))

gtsave(gtbl_outcome_admission,
       file.path(path.local.msf.tables.oc, paste0('gtbl_outcome_admission', '_', week_report, '.png'))) %>%
  invisible()

gtsave(gtbl_outcome_admission,
       file.path(path.local.msf.tables.oc, paste0('gtbl_outcome_admission', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_outcome_admission %>%
  tab_header(
    title = 'Count of patients outcome by Admitted/Not Admitted'
  )

```


## Case fatality by level of care received

```{r tbl-cfr-status-hospi-icu-oxy-venti}
dta_cfr_status_hospi_icu <- dta_linelist %>%
  filter(ind_MSF_covid_status %in% c('Confirmed', 'Probable', 'Suspected')) %>%
  filter(ind_outcome_patcourse_status %in% c('Cured', 'Died', 'Sent back home')) %>%
  # filter(merge_admit != "(Unknown)") %>%
  select(ind_MSF_covid_status, merge_admit, ind_outcome_patcourse_status, merge_icu, merge_oxygen,
         merge_vent) %>%
  mutate(merge_admit = recode(merge_admit, "Yes" = "Admitted", "No" = "Not Admitted"))


# Hospi
dta_cfr_status_hospi <- unique(dta_cfr_status_hospi_icu$merge_admit) %>%
  map_df(~{
    dta_cfr_status_hospi_icu %>%
      filter(merge_admit == .x) %>%
      tbl_prop('ind_outcome_patcourse_status', 'ind_MSF_covid_status', drop_levels = FALSE) %>%
      mutate(
        merge_admit = .x)
  })  %>%
  filter(ind_outcome_patcourse_status == 'Died')

# ICU
dta_cfr_status_icu <- unique(dta_cfr_status_hospi_icu$merge_icu) %>%
  map_df(~{
    dta_cfr_status_hospi_icu %>%
      filter(merge_icu == .x) %>%
      tbl_prop('ind_outcome_patcourse_status', 'ind_MSF_covid_status', drop_levels = FALSE) %>%
      mutate(
        merge_icu = .x)
  })  %>%
  filter(ind_outcome_patcourse_status == 'Died')


# O2
dta_cfr_status_o2 <- unique(dta_cfr_status_hospi_icu$merge_oxygen) %>%
  map_df(~{
    dta_cfr_status_hospi_icu %>%
      filter(merge_oxygen == .x) %>%
      tbl_prop('ind_outcome_patcourse_status', 'ind_MSF_covid_status', drop_levels = FALSE) %>%
      mutate(
        merge_oxygen = .x)
  })  %>%
  filter(ind_outcome_patcourse_status == 'Died')


# Ventilation
dta_cfr_status_ventilation <- unique(dta_cfr_status_hospi_icu$merge_vent) %>%
  map_df(~{
    dta_cfr_status_hospi_icu %>%
      filter(merge_vent == .x) %>%
      tbl_prop('ind_outcome_patcourse_status', 'ind_MSF_covid_status', drop_levels = FALSE) %>%
      mutate(
        merge_vent = .x)
  })  %>%
  filter(ind_outcome_patcourse_status == 'Died')



tbl_cfr_status_hospi_icu_oxy <- dta_cfr_status_hospi %>%
  rename(var = merge_admit) %>%
  mutate(label_var = 'Admission status') %>%
  
  add_row(dta_cfr_status_icu %>%
            rename(var = merge_icu) %>%
            mutate(label_var = 'ICU')) %>%
  
  add_row(dta_cfr_status_o2 %>%
            rename(var = merge_oxygen) %>%
            mutate(label_var = 'Oxygen')) %>%
  
  add_row(dta_cfr_status_ventilation %>%
            rename(var = merge_vent) %>%
            mutate(label_var = 'Non invasive ventilation'))



gtbl_cfr_status_hospi_icu_oxy_venti <- tbl_cfr_status_hospi_icu_oxy %>%
  gt(rowname_col = "var",
     groupname_col = "label_var") %>%
  cols_hide(
    columns = vars('ind_outcome_patcourse_status', 'n_Not a case', 'n_(Unknown)',
                   'N_Not a case', 'N_(Unknown)', 'p_Not a case', 'p_(Unknown)')) %>%
  cols_label(
    n_Confirmed = 'n',
    n_Probable  = 'n',
    n_Suspected = 'n',
    n_Total     = 'n',
    
    N_Confirmed = 'N',
    N_Probable  = 'N',
    N_Suspected = 'N',
    N_Total     = 'N',
    
    p_Confirmed = '(%)',
    p_Probable  = '(%)',
    p_Suspected = '(%)',
    p_Total     = '(%)') %>%
  
  tab_spanner(
    label = 'Confirmed',
    columns = ends_with('_Confirmed')) %>%
  tab_spanner(
    label = 'Probable',
    columns = ends_with('_Probable')) %>%
  tab_spanner(
    label = 'Suspected',
    columns = ends_with('_Suspected')) %>%
  tab_spanner(
    label = 'Total',
    columns = ends_with('_Total')) %>%
  
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = '({x})') %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  fmt_missing(
    columns = starts_with('p_'),
    missing_text = "---") %>%
  cols_align(
    align = 'left',
    columns = vars(var)) %>%
  cols_align(
    align = 'right',
    columns = starts_with('n_')) %>%
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2))


gtsave(gtbl_cfr_status_hospi_icu_oxy_venti,
       file.path(path.local.msf.tables.oc, paste0('gtbl_cfr_status_hospi_icu_oxy_venti', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_cfr_status_hospi_icu_oxy_venti %>%
  tab_header(
    title = 'Case fatality risk by status, admission and ICU ')
```

## Case fatality by type of comorbidity

<!-- CFR by type of comorbidity among confirmed patients -->

```{r tbl-cfr-type-comorbidity}
# DONE by Francesco

dta_cfr_comcond <- prepare_msf_dta_comcond(dta_linelist) %>%
  mutate(
    Comcond_other = case_when(
      !is.na(Comcond_other) ~ 'Yes',
      TRUE ~ NA_character_)) %>%
  filter(ind_MSF_covid_status == 'Confirmed', 
         ind_outcome_patcourse_status %in% c('Cured', 'Died','Sent back home'),
       merge_admit == "Yes") %>%
  pivot_longer(-c(ind_MSF_covid_status, ind_outcome_patcourse_status, 
                  continent, ind_Comcond_count, ind_Comcond_01, Comcond_count,
                  Comcond_01), 
               names_to = "type_comorbidity") %>%
  mutate(
    type_comorbidity = factor(type_comorbidity, levels = df_labels_comcond$levels, labels = df_labels_comcond$labels)) %>%
  filter(!is.na(value), value == 'Yes')

tbl_cfr_type_comcond <- levels(dta_cfr_comcond$type_comorbidity) %>%
  map_df(~{
    dta_cfr_comcond %>%
      filter(type_comorbidity == .x) %>%
      tbl_prop('ind_outcome_patcourse_status', 'continent', drop_levels = FALSE) %>%
      mutate(
        type_comorbidity = .x)
  }) %>%
  filter(ind_outcome_patcourse_status == 'Died') %>%
  arrange(desc(p_Total))

gtbl_cfr_type_comcond <- tbl_cfr_type_comcond %>%
  gt() %>%
  cols_hide('ind_outcome_patcourse_status') %>%
  cols_move_to_start(columns = vars(type_comorbidity)) %>%
  cols_label(
    type_comorbidity = "Comorbidities/Underlying conditions",
    n_Africa   = 'n',
    n_Americas = 'n',
    n_Asia     = 'n',
    n_Europe   = 'n',
    n_Total    = 'n',
    N_Africa   = 'N',
    N_Americas = 'N',
    N_Asia     = 'N',
    N_Europe   = 'N',
    N_Total    = 'N',
    p_Africa   = '(%)',
    p_Americas = '(%)',
    p_Asia     = '(%)',
    p_Europe   = '(%)',
    p_Total    = '(%)')  %>%
  tab_spanner(
    label = 'Africa',
    columns = ends_with('_Africa')) %>%
  tab_spanner(
    label = 'Americas',
    columns = ends_with('_Americas')) %>%
  tab_spanner(
    label = 'Asia',
    columns = ends_with('_Asia')) %>%
  tab_spanner(
    label = 'Europe',
    columns = ends_with('_Europe')) %>%
  tab_spanner(
    label = 'Total',
    columns = ends_with('_Total')) %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = '({x})') %>%
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0/0") %>%
  fmt_missing(
    columns = starts_with('p_'),
    missing_text = "---") %>%
  cols_align(
    align = 'left',
    columns = vars(type_comorbidity)) %>%
  cols_align(
    align = 'right',
    columns = starts_with('n_')) %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2))


gtsave(gtbl_cfr_type_comcond,
       file.path(path.local.msf.tables.oc, paste0('gtbl_cfr_type_comcond', '_', week_report, '.png'))) %>%
  invisible()

#gtsave(gtbl_cfr_type_comcond,
#       file.path(path.local.msf.tables, paste0('gtbl_cfr_type_comcond', '_', week_report, '.html')),
#       inline_css = TRUE) %>%
#  invisible()


gtbl_cfr_type_comcond %>%
  tab_header(
    title = 'Case fatality risk by type of comorbidities among admitted, confirmed patients'
  )
```


<!-- CFR by age-group -->
```{r tbl-cfr-agegroup, fig.cap = 'Case fatality risk by age-group among confirmed patients'}
age_cut <- c(seq(0, 80, 10), Inf)
age_labs <- c("0-9", "10-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80+")

df_MSF_severity <- dta_linelist %>%
  mutate(age_group = cut(age_in_years, breaks = age_cut, labels = age_labs, right = FALSE, include.lowest = TRUE)) %>%
  filter(!is.na(age_group), 
         ind_MSF_covid_status == "Confirmed",
         ind_outcome_patcourse_status %in% c('Cured', 'Died','Sent back home'),
       merge_admit == "Yes") %>%
  group_by(age_group) %>%
  summarize(
    n = n(),
    n_hosp = sum(merge_admit == "Yes", na.rm = TRUE),
    n_hosp_known = sum(merge_admit %in% c("Yes", "No"), na.rm = TRUE),
    p_hosp = n_hosp / n_hosp_known,
    p_hosp_low95 = as.numeric(binom.test(n_hosp, n_hosp_known)$conf.int)[1],
    p_hosp_upp95 = as.numeric(binom.test(n_hosp, n_hosp_known)$conf.int)[2],
    n_died = sum(ind_outcome_patcourse_status == "Died" & merge_admit == "Yes", na.rm = TRUE),
    n_died_known = sum(ind_outcome_patcourse_status %in% c("Died", "Cured", "Sent back home") & merge_admit == "Yes", na.rm = TRUE),
    p_died_low95 = as.numeric(binom.test(n_died, n_died_known)$conf.int)[1],
    p_died_upp95 = as.numeric(binom.test(n_died, n_died_known)$conf.int)[2],
    p_died = n_died / n_died_known,
    # hosp_lab = paste0("(", n_hosp_known, ")"),
    # died_lab = paste0("(", n_died_known, ")"),
    hosp_lab = paste0("(", n_hosp, "/", n_hosp_known, ")"),
    died_lab = paste0("(", n_died, "/", n_died_known, ")")
  ) %>%
  ungroup()

# p1 <- ggplot(MSF_severity, aes(x = age_group)) +
#   geom_point(aes(y = p_hosp)) +
#   geom_linerange(aes(ymin = p_hosp_low95, ymax = p_hosp_upp95)) +
#   geom_text(aes(label = hosp_lab, y = p_hosp_upp95 + 0.06), size = 4.2) +
#   scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2), labels = scales::percent_format(accuracy = 1)) +
#   labs(x = NULL, y = "Hospitalisés")

dots_cfr_agegroup <- ggplot(df_MSF_severity, aes(x = age_group)) +
  geom_point(aes(y = p_died), size = 2) +
  geom_linerange(aes(ymin = p_died_low95, ymax = p_died_upp95)) +
  geom_text(aes(label = died_lab, y = p_died_upp95 + 0.02), size = 4.2) +
  scale_y_continuous(breaks = seq(0, 1, 0.2), limits = c(0, 1), labels = scales::percent_format(accuracy = 1), sec.axis = dup_axis(name = NULL)) +
  labs(x = "Age Group", y = "CFR", caption = "CFR = deaths / known outcomes (cured + died + sent back home) among the admitted\nLine range shows binomial 95% confidence intervals")

ggsave(file.path(path.local.msf.graphs.oc, paste0('dots_cfr_agegroup', '_', week_report, '.png')),
       plot = dots_cfr_agegroup,
       scale = 1.1,
       dpi = 320)


dots_cfr_agegroup
```

---

# Indicators over time among Covid19 confirmed cases

## Weekly case fatality risk

<!--
**With all confirmed cases as denominator: time-series and trends plot** with week as time unit of the following indicator:

Time-series plot:
- Proportion of admitted over the confirmed - line
- Proportion of patients with severe symptoms - line
- Delay from symptoms onset to consultation/admission - boxplot
- Length of staying for patients hospitalised - boxplot
- Case fatality risk for patients hospitalised - line
-->


<!-- CFR by week overall -->
```{r line-cfr-week, fig.cap = 'Evolution overtime of the case fatality risk among Covid confirmed patients consulted/admitted to an MSF supported health facility'}
# DONE by Paul
# Only including outcomes of cured or died as CFR denominator - is this ok? FG: Yes

# 2 most recent epi weeks to remove from chart
latest_2_epiweeks <- max_2(dta_linelist$epi_week_consultation)

tbl_cfr_epiweek_prop <- dta_linelist %>%
  filter(
    ind_MSF_covid_status %in% c("Confirmed", "Suspected", "Probable"),
    ind_outcome_patcourse_status %in% c("Cured", "Died", "Sent back home"),
    merge_admit == "Yes",
    !epi_week_consultation %in% latest_2_epiweeks
  ) %>%
  drop_na(epi_week_consultation) %>%
  group_by(epi_week_consultation) %>%
  summarise(
    total = sum(!is.na(ind_outcome_patcourse_status)),
    dead = sum(ind_outcome_patcourse_status == "Died", na.rm = T),
    prop_dead = (dead / total) #* 100
  ) %>%
  ungroup()

tbl_cfr_epiweek <- dta_linelist %>%
  filter(
    ind_MSF_covid_status %in% c("Confirmed", "Suspected", "Probable"),
    ind_outcome_patcourse_status %in% c("Cured", "Died", "Sent back home"),
    merge_admit == "Yes",
    !epi_week_consultation %in% latest_2_epiweeks
  ) %>%
  count(epi_week_consultation, ind_outcome_patcourse_status) %>%
  drop_na() %>%
  ungroup()

n_max <- max(tbl_cfr_epiweek_prop$total, na.rm = TRUE)
cfr_max <- rounder(max(tbl_cfr_epiweek_prop$prop_dead, na.rm = TRUE), .1)
scaling_factor <- cfr_max / n_max # sets the y limit of CFR rounded up to nearest 10% above cfr max

line_cfr_week <- ggplot(tbl_cfr_epiweek, aes(epi_week_consultation, n)) +
  geom_col(aes(fill = ind_outcome_patcourse_status)) +
  geom_line(data = tbl_cfr_epiweek_prop,
            aes(y = prop_dead / scaling_factor, colour = "CFR"),
            key_glyph = "timeseries", size = 1) +
  scale_x_date(name = "Week of Admission", date_breaks = "2 weeks", date_labels = "%V", sec.axis = ggplot2::sec_axis(trans = ~ .),
               expand = expansion(mult = c(0.01, 0.01))) + #, labels = function(x) format(x, "%b-%Y")
  scale_y_continuous(sec.axis = ggplot2::sec_axis(~ . * scaling_factor, name = "Case Fatality Risk (admitted)", labels = scales::percent_format(accuracy = 1)),
                     expand = expansion(mult = c(0, 0.02))) +
  ggthemes::scale_fill_tableau(palette = "Tableau 20") +
  labs(y = "Patients", colour = NULL, fill = "Outcome", caption = "Latest 2 weeks omitted for data completeness purposes") +
  theme(legend.position = "top", panel.grid.minor = element_blank())

ggsave(file.path(path.local.msf.graphs.oc, paste0('line_cfr_week', '_', week_report, '.png')),
       plot = line_cfr_week,
       width = 6,
       height = 4,
       scale = 1.1,
       dpi = 320)


line_cfr_week
```


<!-- CFR by week and by continent -->
```{r line-cfr-week-continent, fig.cap = 'Evolution overtime of the case fatality risk among Covid confirmed patients consulted/admitted to an MSF supported health facility, separated by continent', fig.width = 8, fig.height = 8}
# DONE by Paul
# Only including outcomes of cured or died as CFR denominator - is this ok? FG: Yes

# 2 most recent epi-weeks to remove from chart
latest_2_epiweeks <- max_2(dta_linelist$epi_week_consultation)

tbl_cfr_epiweek_continent_prop <- dta_linelist %>%
  filter(
    ind_MSF_covid_status %in% c("Confirmed", "Suspected", "Probable"),
    ind_outcome_patcourse_status %in% c("Cured", "Died", "Sent back home"),
    merge_admit == "Yes",
    !epi_week_consultation %in% latest_2_epiweeks
  ) %>%
  group_by(continent, epi_week_consultation) %>%
  summarise(
    total = sum(!is.na(ind_outcome_patcourse_status)),
    dead = sum(ind_outcome_patcourse_status == "Died", na.rm = T),
    prop_dead = (dead / total) #* 100
  ) %>%
  ungroup()

tbl_cfr_epiweek_continent <- dta_linelist %>%
  filter(
    ind_MSF_covid_status %in% c("Confirmed", "Suspected", "Probable"),
    ind_outcome_patcourse_status %in% c("Cured", "Died", "Sent back home"),
    merge_admit == "Yes",
    !epi_week_consultation %in% latest_2_epiweeks
  ) %>%
  count(continent, epi_week_consultation, ind_outcome_patcourse_status) %>%
  drop_na()

n_max <- max(tbl_cfr_epiweek_continent_prop$total, na.rm = TRUE)
cfr_max <- rounder(max(tbl_cfr_epiweek_continent_prop$prop_dead, na.rm = TRUE), .1)
cfr_max <- ifelse(cfr_max > 1, 1, cfr_max)
scaling_factor <- cfr_max / n_max # sets the y limit of CFR rounded up to nearest 10% above cfr max

line_cfr_week_continent <- ggplot(tbl_cfr_epiweek_continent, aes(epi_week_consultation, n)) +
  facet_wrap(~continent, ncol = 2) +
  geom_col(aes(fill = ind_outcome_patcourse_status)) +
  geom_line(data = tbl_cfr_epiweek_continent_prop,
            aes(y = prop_dead / scaling_factor, colour = "CFR"),
            key_glyph = "timeseries", size = 1) +
  scale_x_date(name = "Week of Admission", date_breaks = "3 weeks", date_labels = "%V", sec.axis = ggplot2::sec_axis(trans = ~ .),
               expand = expansion(mult = c(0.01, 0.01))) + #, labels = function(x) format(x, "%b-%Y")
  scale_y_continuous(sec.axis = ggplot2::sec_axis(~ . * scaling_factor, name = "Case Fatality Risk (admitted)", labels = scales::percent_format(accuracy = 1)),
                     expand = expansion(mult = c(0, 0.02))) +
  ggthemes::scale_fill_tableau(palette = "Tableau 20") +
  labs(y = "Patients", colour = NULL, fill = "Outcome", caption = "Latest 2 weeks omitted for data completeness purposes") +
  theme(legend.position = "top", panel.grid.minor = element_blank())

ggsave(file.path(path.local.msf.graphs.oc, paste0('line_cfr_week_continent', '_', week_report, '.png')),
       plot = line_cfr_week_continent,
       width = 8,
       height = 8,
       scale = 1.1,
       dpi = 320)


line_cfr_week_continent
```

## Delay from onset of symptpoms to consultation/admission

```{r boxplot-delay-admission-week, fig.cap = 'Evolution overtime of the delay to admission among Covid confirmed patients consulted/admitted to an MSF supported health facility'}
# DONE by Paul

dta_delay <- dta_linelist %>%
  filter(ind_MSF_covid_status %in% c("Confirmed")) %>%
  select(epi_week_consultation, patcourse_dateonset, MSF_date_consultation, ind_MSF_covid_status) %>%
  drop_na(epi_week_consultation) %>%
  mutate(
    delay_before_consultation = as.numeric(MSF_date_consultation - patcourse_dateonset)) %>%
  filter(between(delay_before_consultation, left = 0, right = 50)) # filter any delays negative or over 50 days


boxplot_delay_before_consultation <- ggplot(dta_delay, aes(x = epi_week_consultation, y = delay_before_consultation)) +
  geom_boxplot(aes(group = epi_week_consultation)) +
  #geom_smooth(fill = "lightblue") +
  scale_x_date(name = "Week of Consultation", date_breaks = "2 weeks", date_labels = "%V", sec.axis = ggplot2::sec_axis(trans = ~ .)) +
  labs(y = "Delay to consultation (days)")

ggsave(file.path(path.local.msf.graphs.oc, paste0('boxplot_delay_before_consultation', '_', week_report, '.png')),
       plot = boxplot_delay_before_consultation,
       width = 6,
       height = 4,
       scale = 1.1,
       dpi = 320)

boxplot_delay_before_consultation

```

## Length of stay in hospital

```{r boxplot-length-stay-week, fig.cap = 'Evolution overtime of of length of stay among Covid confirmed, probable and suspected patients in an MSF supported health facility'}
# DONE by Paul

# 2 most recent epi weeks to remove from chart
latest_2_epiweeks <- max_2(dta_linelist$epi_week_consultation)

# length stay
dta_MSF_length_stay <- dta_linelist %>%
  select(epi_week_consultation, ind_MSF_covid_status, merge_admit, MSF_length_stay) %>%
  filter(ind_MSF_covid_status %in% c('Confirmed'), merge_admit == 'Yes') %>%
  drop_na(epi_week_consultation, MSF_length_stay) %>%
  filter(between(MSF_length_stay, left = 0, right = 50), !epi_week_consultation %in% latest_2_epiweeks)


boxplot_MSF_length_stay_week <- ggplot(dta_MSF_length_stay, aes(epi_week_consultation, MSF_length_stay, group = epi_week_consultation)) +
  geom_boxplot() +
  #geom_jitter(colour = "red", alpha = 0.3) +
  scale_x_date(name = "Week of Admission", date_breaks = "2 weeks", date_labels = "%V", sec.axis = ggplot2::sec_axis(trans = ~ .)) +
  labs(y = "Length of stay in hospital (days)", caption = "Latest 2 weeks omitted for data completeness purposes")

ggsave(file.path(path.local.msf.graphs.oc, paste0('boxplot_MSF_length_stay', '_', week_report, '.png')),
       plot = boxplot_MSF_length_stay_week,
       width = 6,
       height = 4,
       scale = 1.1,
       dpi = 320)

boxplot_MSF_length_stay_week
```

# Analysis of non-cases

```{r tbl-non-cases-diagnosis, layout="l-body-outset"}

dta_non_cases_diagt <- dta_linelist %>%
  filter(ind_MSF_covid_status == 'Not a case') %>%
  select(MSF_main_diagnosis, ind_outcome_patcourse_status) %>%
  mutate(
    MSF_main_diagnosis = fct_explicit_na(MSF_main_diagnosis, na_level = "Not reported"))

tbl_non_cases_diagt <- tbl_prop(dta_non_cases_diagt, 'MSF_main_diagnosis', 'ind_outcome_patcourse_status', prop_by = 'row')

gtbl_non_cases_diagt <- tbl_non_cases_diagt %>%
  gt() %>%
  cols_hide(
    columns = starts_with('N_', ignore.case = FALSE)) %>%
  cols_label(
    MSF_main_diagnosis = '',
    n_Cured = 'n',
    n_Died = 'n',
    `n_Left against medical advice` = 'n',
    n_Transferred = 'n',
    `n_Sent back home` = 'n',
    n_Other = 'n',
    `n_(Pending/Unknown)` = 'n',
    n_Total = 'n',
    p_Cured = '(%)',
    p_Died = '(%)',
    `p_Left against medical advice` = '(%)',
    p_Transferred = '(%)',
    `p_Sent back home` = '(%)',
    p_Other = '(%)',
    `p_(Pending/Unknown)` = '(%)',
    p_Total = '(%)') %>%
  tab_spanner(
    label = 'Recovered',
    columns = ends_with('_Cured')) %>%
  tab_spanner(
    label = 'Died',
    columns = ends_with('_Died')) %>%
  tab_spanner(
    label = html('Left against<br>medical advice'),
    columns = ends_with('_Left against medical advice')) %>%
  tab_spanner(
    label = 'Transferred',
    columns = ends_with('_Transferred')) %>%
  tab_spanner(
    label = html('Sent<br>back home'),
    columns = ends_with('_Sent back home')) %>%
  tab_spanner(
    label = 'Other',
    columns = ends_with('_Other')) %>%
  tab_spanner(
    label = html('Pending/<br>Unknown'),
    columns = ends_with('_Pending/Unknown')) %>%
  tab_spanner(
    label = 'Total',
    columns = ends_with('_Total')) %>%
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100,
    pattern = "({x})") %>%
  cols_align(
    align = 'right',
    columns = starts_with('n_')) %>%
  cols_align(
    align = 'left',
    columns = starts_with('p_')) %>%
  cols_align(
    align = 'left',
    columns = vars(MSF_main_diagnosis)) %>%
  tab_style(
    style = list(
      cell_text(align = "right")),
    locations = cells_column_labels(starts_with('n_'))) %>%
  tab_style(
    style = list(
      cell_text(align = "left")),
    locations = cells_column_labels(starts_with('p_'))) %>%
  tab_options(
    column_labels.font.weight = "bold",
    data_row.padding = px(5))


gtsave(gtbl_non_cases_diagt,
       file.path(path.local.msf.tables.oc, paste0('gtbl_non_cases_diagt', '_', week_report, '.png'))) %>%
  invisible()


gtsave(gtbl_non_cases_diagt,
       file.path(path.local.msf.tables.oc, paste0('gtbl_non_cases_diagt', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtbl_non_cases_diagt %>%
  tab_header(
    title = 'Frequency of main diagnoses among non confirmed patients who attended an MSF Covid facility'
  )
```



# Analysis by project

Inclusion criteria: patients were included in the study if COVID suspected or probable or confirmed case independently of level of care provided in the project.


## Testing

```{r tbl-testing}
tble19 <- dta_testing %>%
  group_by(country, project) %>%
  summarise(n_patient_PCR = sum(merge_test_PCR == "Yes", na.rm = TRUE),
            n_patient_PCR_positive = sum(merge_test_PCR_positive == "Yes", na.rm = TRUE),
            perct_patient_pos = ifelse(n_patient_PCR == 0,
                                       NA_integer_,
                                       n_patient_PCR_positive / n_patient_PCR),
            a_patient_positive = format_nbp(n_patient_PCR_positive, perct_patient_pos),
            
            n_test_PCR = sum(n_test_PCR, na.rm = TRUE),
            n_test_PCR_positive =sum(n_test_PCR_positive, na.rm = TRUE),
            perct_test_pos = ifelse(n_test_PCR == 0,
                                    NA_integer_,
                                    n_test_PCR_positive / n_test_PCR),
            a_test_positive = format_nbp(n_test_PCR_positive, perct_test_pos)
  ) %>%
  select(-c(n_patient_PCR_positive, perct_patient_pos,  n_test_PCR_positive, perct_test_pos))




gtble19 <- tble19 %>%
  gt() %>%
  cols_label(
    country          = 'Country',
    project          = 'Project',
    
    n_test_PCR     = 'Total',
    a_test_positive = "Positive",
    
    n_patient_PCR     = "Total",
    a_patient_positive = "Positive") %>%
  
  tab_spanner(
    label = 'Number of tests',
    columns = contains("_test_")) %>%
  tab_spanner(
    label = 'Number of patients',
    columns = contains('_patient_')) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)


gtsave(gtble19,
       file.path(path.local.msf.tables.oc, paste0('gtble19', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble19 %>%
  tab_header(
    title = 'Number of tests and patient tested')

```


## Time symptom onset to presentations

```{r fig4, fig.cap = "Time from symptom onset to presentation"}
fig4 <- dta_linelist_project %>%
  mutate(delay_before_consultation = as.numeric(MSF_date_consultation - patcourse_dateonset)) %>%
  filter(between(delay_before_consultation, left = 0, right = 50)) %>%
  # select(country, project, merge_admit, MSF_severity, MSF_length_stay) %>%
  group_by(continent, country, project) %>%
  summarise(x = quantile(delay_before_consultation, c(0.25, 0.5, 0.75), na.rm = TRUE),
            q = c(0.25, 0.5, 0.75),
            n = n()) %>%
  pivot_wider(names_from = q,
              values_from =x,
              names_prefix = "q_") %>%
  mutate(project = paste(country, project, sep = " - ")) %>%
  
  ggplot(aes(x = fct_reorder(project, desc(q_0.5)),
             y = q_0.5,
             colour = continent)) +
  geom_point(aes(size = n),
             shape = 19) +
  geom_errorbar(aes(ymin = q_0.25,
                    ymax = q_0.75),
                width = 0) +
  coord_flip() +
  labs(x = "",
       y = "Median time",
       title = "Median time between symptom onset\nand presentation",
       subtitle = "Median and interquartile range") +
  ggthemes::scale_colour_tableau(name = "Continent", palette = "Red-Blue-Brown") +
  scale_size_continuous(name = "Number of\npresentations",
                        breaks = c(10, 100, 1000, 5000)) +
  theme_light() +
  theme(legend.position = 'right',
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

ggsave(file.path(path.local.msf.graphs.oc, paste0('fig4', '_', week_report, '.png')),
       plot = fig4,
       width = 6,
       height = 4,
       scale = 1.1,
       dpi = 320)

fig4
```



## Patient visits and hospitalisation

### Summary
```{r tble1}
tble1 <- dta_linelist_project %>%
  select(country, project, patcourse_admit, outcome_patcourse_admit, merge_admit) %>%
  group_by(country, project) %>%
  summarise(n_patients           = n(),
            
            n_admitted_admission = sum(patcourse_admit == "Yes", na.rm = TRUE),
            p_admitted_admission = n_admitted_admission / n_patients,
            
            n_admited_post       = sum(outcome_patcourse_admit == "Yes" &
                                         patcourse_admit != "Yes", na.rm = TRUE),
            p_admited_post       = n_admited_post / n_patients,
            
            n_admit_all          = sum(merge_admit == "Yes", na.rm = TRUE),
            p_admit_all          = n_admit_all / n_patients)


gtble1 <- tble1 %>%
  gt() %>%
  cols_label(
    country              = 'Country',
    project              = 'Project',
    n_patients           = 'Visits',
    n_admitted_admission = 'N',
    p_admitted_admission = "%",
    n_admited_post       = 'N',
    p_admited_post       = "%",
    n_admit_all          = 'N',
    p_admit_all          = "%") %>%
  
  tab_spanner(
    label = 'Hospitalised at presentation',
    columns = ends_with('_admitted_admission')) %>%
  
  tab_spanner(
    label = 'Hospitalised during course of disease',
    columns = ends_with('_admited_post')) %>%
  
  tab_spanner(
    label = 'Hospitalised (any time)',
    columns = ends_with('_admit_all')) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100) %>%
  
  # cols_align(
  #   align = 'right',
  #   columns = starts_with('n_')) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = vars(n_patients, n_admitted_admission, n_admited_post, n_admit_all),
    fns = list(Total = "sum"),
    decimals = 0)


gtsave(gtble1,
       file.path(path.local.msf.tables.oc, paste0('gtble1', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble1 %>%
  tab_header(
    title = 'Patient visits and hospitalisation status at admission and over course of disease, by project')
```

### By MSF_severity

```{r tble2}
tble2 <- dta_linelist_project %>%
  select(country, project, merge_admit, 
         MSF_severity, MSF_length_stay) %>%
  group_by(country, project) %>%
  summarise(n_patients = n(),
            n_admis    = sum(merge_admit == "Yes", na.rm = TRUE),
            
            n_mild       = sum(MSF_severity == "Mild", na.rm = TRUE),
            n_mild_admis = sum(MSF_severity == "Mild" & merge_admit == "Yes"),
            p_mild_admis = n_mild_admis / n_mild,
            
            n_moderate       = sum(MSF_severity == "Moderate", na.rm = TRUE),
            n_moderate_admis = sum(MSF_severity == "Moderate" & merge_admit == "Yes", na.rm = TRUE),
            p_moderate_admis = n_moderate_admis / n_moderate,
            
            n_severe       = sum(MSF_severity == "Severe", na.rm = TRUE),
            n_severe_admis  = sum(MSF_severity == "Severe" & merge_admit == "Yes", na.rm = TRUE),
            p_severe_admis = n_severe_admis / n_severe,
            
            n_critical        = sum(MSF_severity == "Critical", na.rm = TRUE),
            n_critical_admis  = sum(MSF_severity == "Critical" & merge_admit == "Yes", na.rm = TRUE),
            p_critical_admis  = n_critical_admis / n_critical,
            
            n_na       = sum(is.na(MSF_severity)),
            p_na       = n_na / n_patients)


gtble2 <- tble2 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    n_patients = 'Visits',
    n_admis    = 'Admissions',
    
    n_mild       = 'N visits',
    n_mild_admis = 'N admissions',
    p_mild_admis = "% admitted",
    
    n_moderate       = 'N visits',
    n_moderate_admis = 'N admissions',
    p_moderate_admis = "% admitted",
    
    n_severe       = 'N visits',
    n_severe_admis = 'N admissions',
    p_severe_admis = "% admitted",
    
    n_critical       = 'N visits',
    n_critical_admis = 'N admissions',
    p_critical_admis = "% admitted",
    
    n_na = 'MSF_severity',
    p_na = "% non-available") %>%
  
  tab_spanner(
    label = 'All MSF_severity',
    columns = vars(n_patients, n_admis)) %>%
  tab_spanner(
    label = 'Mild',
    columns = contains('_mild')) %>%
  tab_spanner(
    label = 'Moderate',
    columns = contains('_moderate')) %>%
  tab_spanner(
    label = 'Severe',
    columns = contains('_severe')) %>%
  tab_spanner(
    label = 'Critical',
    columns = contains('_critical')) %>%
  tab_spanner(
    label = 'Missing data',
    columns = ends_with('_na')) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  fmt_number(
    columns = starts_with('p_'),
    decimals = 1,
    scale_by = 100) %>%
  
  cols_align(
    align = 'center') %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding  = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)

gtsave(gtble2,
       file.path(path.local.msf.tables.oc, paste0('gtble2', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble2 %>%
  tab_header(
    title = 'Patient visits and hospitalisation by project and MSF_severity')
```


### Length of stay and MSF_severity

```{r tbl-length-stay-MSF_severity}
tble3 <- dta_linelist_project %>%
  select(country, project, merge_admit, MSF_severity, MSF_length_stay) %>%
  group_by(country, project) %>%
  summarise(
    n_patients = n(),
    n_admis    = sum(merge_admit == "Yes", na.rm = TRUE),
    
    q_mild_25 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Mild"], c(0.25), na.rm = TRUE),
    q_mild_50 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Mild"], c(0.5), na.rm = TRUE),
    q_mild_75 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Mild"], c(0.75), na.rm = TRUE),
    
    q_moderate_25 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Moderate"], c(0.25), na.rm = TRUE),
    q_moderate_50 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Moderate"], c(0.5), na.rm = TRUE),
    q_moderate_75 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Moderate"], c(0.75), na.rm = TRUE),
    
    q_critical_25 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Critical"], c(0.25), na.rm = TRUE),
    q_critical_50 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Critical"], c(0.5), na.rm = TRUE),
    q_critical_75 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Critical"], c(0.75), na.rm = TRUE),
    
    q_severe_25 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Severe"], c(0.25), na.rm = TRUE),
    q_severe_50 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Severe"], c(0.5), na.rm = TRUE),
    q_severe_75 = quantile(MSF_length_stay[merge_admit == "Yes" & MSF_severity == "Severe"], c(0.75), na.rm = TRUE),
    
    na_MSF_severity = sum(is.na(MSF_severity)),
    na_length_of_stay = sum(is.na(MSF_length_stay)),
    na_both = sum(is.na(MSF_severity) & is.na(MSF_length_stay))
  ) %>%
  
  mutate(iqr_mild     = glue("{q_mild_50} [{q_mild_25} - {q_mild_75}]"),
         iqr_moderate = glue("{q_moderate_50} [{q_moderate_25} - {q_moderate_75}]"),
         iqr_severe   = glue("{q_critical_50} [{q_critical_25} - {q_critical_75}]"),
         iqr_critical = glue("{q_severe_50} [{q_severe_25} - {q_severe_75}]")) %>%
  select(-starts_with("q_")) %>%
  relocate(country, project,
           starts_with("n_"),
           starts_with("iqr_"),
           starts_with("na_"))  %>%
  group_by(country)


gtble3 <- tble3 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    
    n_patients = 'Visits',
    n_admis    = 'Admissions',
    
    iqr_mild     = 'Mild',
    iqr_moderate = 'Moderate',
    iqr_severe   = 'Severe',
    iqr_critical = 'Critical',
    
    na_MSF_severity       = "MSF_severity",
    na_length_of_stay = "Length of stay",
    na_both           = "Both") %>%
  
  tab_spanner(
    label = 'Length  of Stay (Median [IQR])',
    columns = starts_with('iqr_')) %>%
  tab_spanner(
    label = 'Missing data',
    columns = starts_with('na')) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center') %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)


gtsave(gtble3,
       file.path(path.local.msf.tables.oc, paste0('gtble3', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble3 %>%
  tab_header(
    title = 'Median length of stay according to MSF_severity')
```




## Age and gender

```{r fig5, fig.cap = "Age and gender of patients"}
fig5 <- dta_linelist_project %>%
  drop_na(patinfo_sex) %>%
  mutate(Sex = factor(patinfo_sex,
                      levels = c('M', 'F'),
                      labels = c('Male', 'Female'))) %>%
  group_by(continent, country, project, Sex) %>%
  summarise(x = quantile(age_in_years, c(0.25, 0.5, 0.75), na.rm = TRUE),
            q = c(0.25, 0.5, 0.75),
            n = n()) %>%
  pivot_wider(names_from = q,
              values_from =x,
              names_prefix = "q_") %>%
  mutate(project = paste(country, project, sep = " - ")) %>%
  
  ggplot(aes(x = fct_reorder(project, desc(q_0.5)),
             y = q_0.5,
             colour = Sex,
             shape = Sex
  )) +
  geom_point(aes(size = n),
             position = position_dodge(width=0.9)) +
  geom_errorbar(aes(ymin = q_0.25,
                    ymax = q_0.75),
                width = 0,
                position = position_dodge(width=0.9)) +
  coord_flip() +
  labs(x = "",
       y = "Median age",
       title = "Median age (in years) of all presentations",
       subtitle = "Median and interquartile range") +
  
  scale_colour_manual(name = "Sex",
                      values = c("#4E79A7FF", "#A0CBE8FF"),
                      breaks = c("Male", "Female"),
                      labels = c("Male", "Female")) +
  
  scale_size_continuous(name = "Visits",
                        breaks = c(10, 100, 1000, 5000)) +
  
  theme_light() +
  theme(legend.position = 'right',
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())


ggsave(file.path(path.local.msf.graphs.oc, paste0('fig5', '_', week_report, '.png')),
       plot = fig5,
       width = 8,
       height = 6,
       scale = 1.1,
       dpi = 320)

fig5
```

```{r tble6}
tble6 <- dta_linelist_project %>%
  select(country, project, ind_MSF_covid_status, merge_admit, Sex = patinfo_sex, age_in_years) %>%
  group_by(country, project, Sex) %>%
  summarise(
    n_patients = n(),
    n_admis    = sum(merge_admit == "Yes", na.rm = TRUE),
    
    n_probable_admis    = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes", na.rm = TRUE),
    q_probable_25_admis = quantile(age_in_years[ind_MSF_covid_status == "Probable" & merge_admit == "Yes"], c(0.25), na.rm = TRUE),
    q_probable_05_admis = quantile(age_in_years[ind_MSF_covid_status == "Probable" & merge_admit == "Yes"], c(0.5), na.rm = TRUE),
    q_probable_75_admis = quantile(age_in_years[ind_MSF_covid_status == "Probable" & merge_admit == "Yes"], c(0.75), na.rm = TRUE),
    f_probable_admis    = format_med(q_probable_05_admis, q_probable_25_admis, q_probable_75_admis),
    
    n_confirmed_admis    = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes", na.rm = TRUE),
    q_confirmed_25_admis = quantile(age_in_years[ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes"], c(0.25), na.rm = TRUE),
    q_confirmed_05_admis = quantile(age_in_years[ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes"], c(0.5), na.rm = TRUE),
    q_confirmed_75_admis = quantile(age_in_years[ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes"], c(0.75), na.rm = TRUE),
    f_confirmed_admis = format_med(q_confirmed_05_admis, q_confirmed_25_admis, q_confirmed_75_admis),
    
    n_probable_out    = sum(ind_MSF_covid_status == "Probable" & merge_admit == "No", na.rm = TRUE),
    q_probable_25_out = quantile(age_in_years[ind_MSF_covid_status == "Probable" & merge_admit == "No"], c(0.25), na.rm = TRUE),
    q_probable_05_out = quantile(age_in_years[ind_MSF_covid_status == "Probable" & merge_admit == "No"], c(0.5), na.rm = TRUE),
    q_probable_75_out = quantile(age_in_years[ind_MSF_covid_status == "Probable" & merge_admit == "No"], c(0.75), na.rm = TRUE),
    f_probable_out    = format_med(q_probable_05_out, q_probable_25_out, q_probable_75_out),
    
    n_confirmed_out    = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "No", na.rm = TRUE),
    q_confirmed_25_out = quantile(age_in_years[ind_MSF_covid_status == "Confirmed" & merge_admit == "No"], c(0.25), na.rm = TRUE),
    q_confirmed_05_out = quantile(age_in_years[ind_MSF_covid_status == "Confirmed" & merge_admit == "No"], c(0.5), na.rm = TRUE),
    q_confirmed_75_out = quantile(age_in_years[ind_MSF_covid_status == "Confirmed" & merge_admit == "No"], c(0.75), na.rm = TRUE),
    f_confirmed_out    = format_med(q_confirmed_05_out, q_confirmed_25_out, q_confirmed_75_out)
  ) %>%
  select(-starts_with("q_"))


gtble6 <- tble6 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    n_patients = 'Visits',
    n_admis    = 'Admissions',
    
    n_probable_admis = 'N probable',
    f_probable_admis = 'Median age probable [IQR]',
    
    n_confirmed_admis = 'N confirmed',
    f_confirmed_admis = 'Median age confirmed [IQR]',
    
    n_probable_out = 'N probable',
    f_probable_out = 'Median age probable [IQR]',
    
    n_confirmed_out = 'N confirmed',
    f_confirmed_out = 'Median age confirmed [IQR]',
  ) %>%
  
  tab_spanner(
    label = 'Hospitalised',
    columns = ends_with("_admis")) %>%
  tab_spanner(
    label = 'Never hospitalised',
    columns = ends_with('_out')) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)

gtsave(gtble6,
       file.path(path.local.msf.tables.oc, paste0('gtble6', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble6 %>%
  tab_header(
    title = 'Median age stratified by gender, project and hospitalisation status')

```




## MSF_severity

```{r tbl-MSF_severity}
tble14 <- dta_linelist_project %>%
  select(country, project, merge_admit, ind_MSF_covid_status, MSF_severity) %>%
  group_by(country, project) %>%
  summarise(
    n_suspected = sum(ind_MSF_covid_status == "Suspected", na.rm = TRUE),
    n_probable  = sum(ind_MSF_covid_status == "Probable", na.rm = TRUE),
    n_confirmed = sum(ind_MSF_covid_status == "Confirmed", na.rm = TRUE),
    
    n_admis = sum(merge_admit == "Yes", na.rm = TRUE),
    
    n_mild     = sum(MSF_severity == "Mild", na.rm = TRUE),
    n_moderate = sum(MSF_severity == "Moderate", na.rm = TRUE),
    n_severe   = sum(MSF_severity == "Severe", na.rm = TRUE),
    n_critical = sum(MSF_severity == "Critical", na.rm = TRUE),
    
    n_missing_MSF_severity = sum(is.na(MSF_severity)),
    
    n_patients = n())


gtble14 <- tble14 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    
    n_suspected = "Suspected",
    n_probable  = "Probable",
    n_confirmed = "Confirmed",
    
    n_admis = "Hospitalised",
    
    n_mild     = "Mild",
    n_moderate = "Moderate",
    n_severe   = "Severe",
    n_critical = "Critical",
    
    n_missing_MSF_severity = "Missing data MSF_severity",
    n_patients = "Total visits") %>%
  
  tab_spanner(
    label = 'Total presentation',
    columns = vars(n_suspected, n_probable, n_confirmed)) %>%
  tab_spanner(
    label = 'MSF_severity',
    columns = vars(n_mild, n_moderate, n_severe, n_critical)) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)

gtsave(gtble14,
       file.path(path.local.msf.tables.oc, paste0('gtble14', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble14 %>%
  tab_header(
    title = 'Case MSF_severity per project')

```


```{r fig-MSF_severity-mortality}
dta_linelist_project %>%
  filter(ind_outcome_patcourse_status == "Died") %>%
  filter(merge_admit == "Yes") %>%
  
  group_by(project, epi_week_admission, MSF_severity) %>%
  summarise(n = n()) %>%
  drop_na(epi_week_admission) %>%
  
  ggplot(aes(x = epi_week_admission,
             y = n,
             colour = MSF_severity)) +
  geom_line() +
  geom_point(alpha = 0.8,
             size = 2) +
  
  scale_x_date(name = "Week of Admission",
               date_breaks = "2 weeks",
               date_labels = "%V",
               sec.axis = ggplot2::sec_axis(trans = ~ .),
               expand = expansion(mult = c(0.01, 0.01))) +
  
  labs(x = "Week of admission",
       y = "Nb of death",
       title = glue::glue("Mortality by MSF_severity for hospitalized patients")) +
  ggthemes::scale_colour_tableau(name = "MSF_severity",
                                 palette = "Tableau 20")

ggsave(filename = paste0('mortality_MSF_severity', '_', week_report, '.png'),
       path = path.local.msf.graphs.oc,
       width = 8,
       height = 4,
       scale = 1.1,
       dpi = 320
)

```


```{r tbl-MSF_severity-mortality}
tble15 <- dta_linelist_project %>%
  select(country, project, merge_admit, ind_MSF_covid_status, MSF_severity, ind_outcome_patcourse_status, merge_icu) %>%
  group_by(country, project) %>%
  summarise(
    n_suspected = sum(ind_MSF_covid_status == "Suspected" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    n_probable  = sum(ind_MSF_covid_status == "Probable"  & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    n_confirmed = sum(ind_MSF_covid_status == "Confirmed" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    
    n_admis = sum(merge_admit == "Yes", na.rm = TRUE),
    
    n_mild     = sum(MSF_severity == "Mild", na.rm = TRUE),
    n_moderate = sum(MSF_severity == "Moderate", na.rm = TRUE),
    n_severe   = sum(MSF_severity == "Severe", na.rm = TRUE),
    n_critical = sum(MSF_severity == "Critical", na.rm = TRUE),
    
    n_deaths_icu = sum(ind_outcome_patcourse_status == "Died" & merge_icu == "Yes", na.rm = TRUE),
    
    n_patients = n())


gtble15 <- tble15 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    
    n_suspected = "Suspected",
    n_probable  = "Probable",
    n_confirmed = "Confirmed",
    
    n_admis = "Hospitalised",
    
    n_mild     = "Mild",
    n_moderate = "Moderate",
    n_severe   = "Severe",
    n_critical = "Critical",
    
    n_deaths_icu = "Death amond ICU patients",
    n_patients = "Total visits") %>%
  
  tab_spanner(
    label = 'Total presentation',
    columns = vars(n_suspected, n_probable, n_confirmed)) %>%
  tab_spanner(
    label = 'Death by MSF_severity',
    columns = vars(n_mild, n_moderate, n_severe, n_critical)) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)

gtsave(gtble15,
       file.path(path.local.msf.tables.oc, paste0('gtble15', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble15 %>%
  tab_header(
    title = 'Mortality by MSF_severity')
```




## Comorbidities

```{r fig-comorb, fig.width=8, fig.height=8, fig.cap = "comorbidities of hospitalized patients per project"}
fig7 <- dta_linelist_project %>%
  filter(merge_admit == "Yes") %>%
  mutate(project = paste(country, project, sep = " - ")) %>%
  group_by(project) %>%
  summarise(
    n_hospi = n(),
    
    n_diabetes = sum(ind_Comcond_diabetes == "Yes", na.rm = TRUE),
    p_diabetes = n_diabetes / n_hospi,
    
    n_ind_MSF_hypertension  = sum(ind_MSF_hypertension == "Yes", na.rm = TRUE),
    p_ind_MSF_hypertension  = n_ind_MSF_hypertension / n_hospi,
    
    n_chronic_renal = sum(Comcond_renal == "Yes", na.rm = TRUE),
    p_chronic_renal = n_chronic_renal / n_hospi) %>%
  pivot_longer(cols = c(p_diabetes, p_ind_MSF_hypertension, p_chronic_renal),
               names_to = "Comorbidities",
               values_to = "percent_of_hospi",
               names_prefix = "p_") %>%
  filter(percent_of_hospi > 0) %>%
  
  ggplot(aes(x = project,
             y = percent_of_hospi,
             colour = Comorbidities)) +
  geom_point(size = 3,
             alpha = 0.8) +
  coord_flip() +
  facet_wrap(~Comorbidities, ncol = 1, scales = "free_y") +
  
  ggthemes::scale_colour_tableau(name = "Comorbidities",
                                 palette = "Tableau 20",
                                 breaks = c("diabetes", "ind_MSF_hypertension", "chronic_renal"),
                                 labels = c("Diabete", "Hypertension", "Renal chronic illness")) +
  
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  
  theme_light() +
  theme(legend.position = 'right',
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  
  labs(x = "",
       y = "Percentage of hospitalised patients",
       title = "Mortality and comorbidities")

ggsave(file.path(path.local.msf.graphs.oc, paste0('fig7', '_', week_report, '.png')),
       plot = fig7,
       width = 8,
       height = 8,
       scale = 1.1,
       dpi = 320)

fig7
```


```{r tbl-comorb}
list_comorb <- c("ind_MSF_ind_MSF_obesity", "ind_ind_Comcond_diabetes", "ind_MSF_hypertension", "ind_Comcond_renal")


tble16 <- dta_linelist_project %>%
  select(country, project, merge_admit, ind_Comcond_01, ind_Comcond_diabetes, ind_MSF_hypertension,
         ind_Comcond_renal, ind_outcome_patcourse_status, ind_MSF_obesity) %>%
  group_by(country, project) %>%
  summarise(
    n_all = n(), # Visits (cases) at project level
    
    n_comorb = sum(ind_Comcond_01 == 1, na.rm = TRUE), # at least one comorb (cases)
    p_comorb = n_comorb / n_all,
    a_comorb = format_nbp(n_comorb, p_comorb),
    
    n_comorb_hospi = sum(ind_Comcond_01 == 1 & merge_admit == "Yes", na.rm = TRUE), #hospitalised, at least one comorb
    p_comorb_hospi = n_comorb_hospi / n_comorb,
    a_comorb_hospi = format_nbp(n_comorb_hospi, p_comorb_hospi),
    
    n_comorb_death = sum(ind_Comcond_01 == 1 & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    p_comorb_death = n_comorb_death / n_comorb_hospi,
    a_comorb_death = format_nbp(n_comorb_death, p_comorb_death),
    
    
    n_ind_MSF_obesity = sum(ind_MSF_obesity == "Yes", na.rm = TRUE),
    p_ind_MSF_obesity = n_ind_MSF_obesity / n_comorb,
    a_ind_MSF_obesity = format_nbp(n_ind_MSF_obesity, p_ind_MSF_obesity),
    
    n_ind_MSF_obesity_hospi = sum(ind_MSF_obesity == "Yes" & merge_admit == "Yes", na.rm = TRUE),
    p_ind_MSF_obesity_hospi = n_ind_MSF_obesity_hospi / n_ind_MSF_obesity,
    a_ind_MSF_obesity_hospi = format_nbp(n_ind_MSF_obesity_hospi, p_ind_MSF_obesity_hospi),
    
    n_ind_MSF_obesity_death = sum(ind_MSF_obesity == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    p_ind_MSF_obesity_death = n_ind_MSF_obesity_death / n_ind_MSF_obesity_hospi,
    a_ind_MSF_obesity_death = format_nbp(n_ind_MSF_obesity_death, p_ind_MSF_obesity_death),
    
    
    n_diabetes = sum(ind_Comcond_diabetes == "Yes", na.rm = TRUE),
    p_diabetes = n_diabetes / n_comorb,
    a_diabetes = format_nbp(n_diabetes, p_diabetes),
    
    n_diabetes_hospi = sum(ind_Comcond_diabetes == "Yes" & merge_admit == "Yes", na.rm = TRUE),
    p_diabetes_hospi = n_diabetes_hospi / n_diabetes,
    a_diabetes_hospi = format_nbp(n_diabetes_hospi, p_diabetes_hospi),
    
    n_diabetes_death = sum(ind_Comcond_diabetes == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    p_diabetes_death = n_diabetes_death / n_diabetes_hospi,
    a_diabetes_death = format_nbp(n_diabetes_death, p_diabetes_death),
    
    
    n_ind_MSF_hypertension  = sum(ind_MSF_hypertension == "Yes", na.rm = TRUE),
    p_ind_MSF_hypertension  = n_ind_MSF_hypertension / n_comorb,
    a_ind_MSF_hypertension  = format_nbp(n_ind_MSF_hypertension, p_ind_MSF_hypertension),
    
    n_ind_MSF_hypertension_hospi  = sum(ind_MSF_hypertension == "Yes" & merge_admit == "Yes", na.rm = TRUE),
    p_ind_MSF_hypertension_hospi  = n_ind_MSF_hypertension_hospi / n_ind_MSF_hypertension,
    a_ind_MSF_hypertension_hospi  = format_nbp(n_ind_MSF_hypertension_hospi, p_ind_MSF_hypertension_hospi),
    
    n_ind_MSF_hypertension_death  = sum(ind_MSF_hypertension == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    p_ind_MSF_hypertension_death  = n_ind_MSF_hypertension_death / n_ind_MSF_hypertension_hospi,
    a_ind_MSF_hypertension_death  = format_nbp(n_ind_MSF_hypertension_death, p_ind_MSF_hypertension_death),
    
    
    n_chronic_renal = sum(ind_Comcond_renal == "Yes", na.rm = TRUE),
    p_chronic_renal = n_chronic_renal / n_comorb,
    a_chronic_renal = format_nbp(n_chronic_renal, p_chronic_renal),
    
    n_chronic_renal_hospi = sum(ind_Comcond_renal == "Yes" & merge_admit == "Yes", na.rm = TRUE),
    p_chronic_renal_hospi = n_chronic_renal_hospi / n_chronic_renal,
    a_chronic_renal_hospi = format_nbp(n_chronic_renal_hospi, p_chronic_renal_hospi),
    
    n_chronic_renal_death = sum(ind_Comcond_renal == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    p_chronic_renal_death = n_chronic_renal_death / n_chronic_renal_hospi,
    a_chronic_renal_death = format_nbp(n_chronic_renal_death, p_chronic_renal_death)
  ) %>%
  select(country, project, starts_with("a_"))



gtble16 <- tble16 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    
    a_comorb       = 'N',
    a_comorb_hospi = 'Hospitalised',
    a_comorb_death = 'Deceased',
    
    a_ind_MSF_obesity       = 'N',
    a_ind_MSF_obesity_hospi = 'Hospitalised',
    a_ind_MSF_obesity_death = 'Deceased',
    
    a_diabetes       = 'N',
    a_diabetes_hospi = 'Hospitalised',
    a_diabetes_death = 'Deceased',
    
    a_ind_MSF_hypertension       = 'N',
    a_ind_MSF_hypertension_hospi = 'Hospitalised',
    a_ind_MSF_hypertension_death = 'Deceased',
    
    a_chronic_renal       = 'N',
    a_chronic_renal_hospi = 'Hospitalised',
    a_chronic_renal_death = 'Deceased') %>%
  
  tab_spanner(
    label = 'At least one documented comorbidity',
    columns = vars(a_comorb, a_comorb_hospi, a_comorb_death)) %>%
  tab_spanner(
    label = 'Obesity',
    columns = contains("_ind_MSF_obesity")) %>%
  tab_spanner(
    label = 'Diabete',
    columns = contains("_diabetes")) %>%
  tab_spanner(
    label = 'Hypertension',
    columns = contains("_ind_MSF_hypertension")) %>%
  tab_spanner(
    label = 'Chronic renal',
    columns = contains("_chronic_renal")) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)

gtsave(gtble16,
       file.path(path.local.msf.tables.oc, paste0('gtble16', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble16 %>%
  tab_header(
    title = 'Number of patients with comorbidities and proportion of hospitalizations and deaths')
```




## Coinfections

```{r fig-coinfections, fig.cap = "coinfections in hospitalized patients"}
fig8 <- dta_linelist_project %>%
  filter(merge_admit == "Yes") %>%
  select(country, project, merge_admit, ind_outcome_patcourse_status, ind_MSF_malaria, ind_MSF_hiv_status, ind_MSF_tb_active,
         ind_MSF_tb_treatment_past, ind_MSF_malnutrition) %>%
  mutate(project = paste(country, project, sep = " - ")) %>%
  group_by(project) %>%
  summarise(
    n_hospi =  n(),
    
    n_ind_MSF_malaria   = sum(ind_MSF_malaria == "Yes", na.rm = TRUE),
    n_hiv       = sum(ind_MSF_hiv_status == "Yes", na.rm = TRUE),
    n_ind_MSF_tb_active = sum(ind_MSF_tb_active == "Yes", na.rm = TRUE),
    n_tb_past   = sum(ind_MSF_tb_treatment_past == "Yes", na.rm = TRUE),
    
    p_ind_MSF_malaria   = n_ind_MSF_malaria / n_hospi,
    p_hiv       = n_hiv / n_hospi,
    p_ind_MSF_tb_active = n_ind_MSF_tb_active / n_hospi,
    p_tb_past   = n_tb_past / n_hospi
  ) %>%
  
  pivot_longer(cols = c(p_ind_MSF_malaria, p_hiv, p_ind_MSF_tb_active, p_tb_past),
               names_to = "Coinfection",
               values_to = "percent_of_hospi",
               names_prefix = "p_") %>%
  filter(percent_of_hospi > 0) %>%
  
  
  ggplot(aes(x = project,
             y = percent_of_hospi,
             colour = Coinfection)) +
  geom_point(size = 3,
             alpha = 0.8) +
  coord_flip() +
  facet_wrap(~Coinfection, ncol = 2, scales = "free_y") +
  
  ggthemes::scale_colour_tableau(name = "Coinfection",
                                 palette = "Tableau 20",
                                 breaks = c("hiv", "ind_MSF_malaria", "ind_MSF_tb_active", "tb_past"),
                                 labels = c("HIV", "Malaria", "TB active", "TB in the past")) +
  
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  
  
  theme_light() +
  theme(legend.position = 'right',
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(x = "",
       y = "Percentage of hospitalised patients",
       title = "Coinfections in hospitalised patients")




ggsave(file.path(path.local.msf.graphs.oc, paste0('fig8', '_', week_report, '.png')),
       plot = fig8,
       width = 10,
       height = 6,
       scale = 1.1,
       dpi = 320)

fig8
```


```{r tbl-coinfections}

list_coinf <- c("ind_MSF_malaria", "ind_MSF_hiv_status", "ind_MSF_tb_active", "ind_MSF_tb_treatment_past", "ind_MSF_malnutrition")


tble17 <-dta_linelist_project %>%
  select(country, project, merge_admit, ind_outcome_patcourse_status, ind_MSF_malaria, ind_MSF_hiv_status,
         ind_MSF_tb_active, ind_MSF_tb_treatment_past, ind_MSF_malnutrition) %>%
  group_by(country, project) %>%
  mutate(coinf = ifelse(ind_MSF_malaria == "Yes" | ind_MSF_hiv_status == "Yes" | ind_MSF_tb_active == "Yes" |
                          ind_MSF_tb_treatment_past == "Yes" | ind_MSF_malnutrition == "Yes",
                        1, 0)) %>%
  summarise(n_coinf = sum(coinf == 1, na.rm = TRUE),
            p_coinf = n_coinf / n(),
            a_coinf = format_nbp(n_coinf, p_coinf),
            
            n_coinf_hospi = sum(coinf == 1 & merge_admit == "Yes", na.rm = TRUE),
            p_coinf_hospi = n_coinf_hospi / n_coinf,
            a_coinf_hospi = format_nbp(n_coinf_hospi, p_coinf_hospi),
            
            n_coinf_death = sum(coinf == 1 & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
            p_coinf_death = n_coinf_death / n_coinf,
            a_coinf_death = format_nbp(n_coinf_death, p_coinf_death),
            
            
            n_ind_MSF_malaria = sum(ind_MSF_malaria == "Yes", na.rm = TRUE),
            p_ind_MSF_malaria = n_ind_MSF_malaria / n_coinf,
            a_ind_MSF_malaria = format_nbp(n_ind_MSF_malaria, p_ind_MSF_malaria),
            
            n_ind_MSF_malaria_hospi = sum(ind_MSF_malaria == "Yes" & merge_admit == "Yes", na.rm = TRUE),
            p_ind_MSF_malaria_hospi = n_ind_MSF_malaria_hospi / n_ind_MSF_malaria,
            a_ind_MSF_malaria_hospi = format_nbp(n_ind_MSF_malaria_hospi, p_ind_MSF_malaria_hospi),
            
            n_ind_MSF_malaria_death = sum(ind_MSF_malaria == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
            p_ind_MSF_malaria_death = n_ind_MSF_malaria_death / n_ind_MSF_malaria,
            a_ind_MSF_malaria_death = format_nbp(n_ind_MSF_malaria_death, p_ind_MSF_malaria_death),
            
            
            n_hiv = sum(ind_MSF_hiv_status == "Yes", na.rm = TRUE),
            p_hiv = n_hiv / n_coinf,
            a_hiv = format_nbp(n_hiv, p_hiv),
            
            n_hiv_hospi = sum(ind_MSF_hiv_status == "Yes" & merge_admit == "Yes", na.rm = TRUE),
            p_hiv_hospi = n_hiv_hospi / n_hiv,
            a_hiv_hospi = format_nbp(n_hiv_hospi, p_hiv_hospi),
            
            n_hiv_death = sum(ind_MSF_hiv_status == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
            p_hiv_death = n_hiv_death / n_hiv,
            a_hiv_death = format_nbp(n_hiv_death, p_hiv_death),
            
            
            n_ind_MSF_tb_active = sum(ind_MSF_tb_active == "Yes", na.rm = TRUE),
            p_ind_MSF_tb_active = n_ind_MSF_tb_active / n_coinf,
            a_ind_MSF_tb_active = format_nbp(n_ind_MSF_tb_active, p_ind_MSF_tb_active),
            
            n_ind_MSF_tb_active_hospi = sum(ind_MSF_tb_active == "Yes" & merge_admit == "Yes", na.rm = TRUE),
            p_ind_MSF_tb_active_hospi = n_ind_MSF_tb_active_hospi / n_ind_MSF_tb_active,
            a_ind_MSF_tb_active_hospi = format_nbp(n_ind_MSF_tb_active_hospi, p_ind_MSF_tb_active_hospi),
            
            n_ind_MSF_tb_active_death = sum(ind_MSF_tb_active == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
            p_ind_MSF_tb_active_death = n_ind_MSF_tb_active_death / n_ind_MSF_tb_active,
            a_ind_MSF_tb_active_death = format_nbp(n_ind_MSF_tb_active_death, p_ind_MSF_tb_active_death),
            
            
            n_tb_past = sum(ind_MSF_tb_treatment_past == "Yes", na.rm = TRUE),
            p_tb_past = n_tb_past / n_coinf,
            a_tb_past = format_nbp(n_tb_past, p_tb_past),
            
            n_tb_past_hospi = sum(ind_MSF_tb_treatment_past == "Yes" & merge_admit == "Yes", na.rm = TRUE),
            p_tb_past_hospi = n_tb_past_hospi / n_tb_past,
            a_tb_past_hospi = format_nbp(n_tb_past_hospi, p_tb_past_hospi),
            
            n_tb_past_death = sum(ind_MSF_tb_treatment_past == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
            p_tb_past_death = n_tb_past_death / n_tb_past,
            a_tb_past_death = format_nbp(n_tb_past_death, p_tb_past_death),
            
            n_ind_MSF_malnutrition = sum(ind_MSF_malnutrition == "Yes", na.rm = TRUE),
            p_ind_MSF_malnutrition = n_ind_MSF_malnutrition / n_coinf,
            a_ind_MSF_malnutrition = format_nbp(n_ind_MSF_malnutrition, p_ind_MSF_malnutrition),
            
            n_ind_MSF_malnutrition_hospi = sum(ind_MSF_malnutrition == "Yes" & merge_admit == "Yes", na.rm = TRUE),
            p_ind_MSF_malnutrition_hospi = n_ind_MSF_malnutrition_hospi / n_ind_MSF_malnutrition,
            a_ind_MSF_malnutrition_hospi = format_nbp(n_ind_MSF_malnutrition_hospi, p_ind_MSF_malnutrition_hospi),
            
            n_ind_MSF_malnutrition_death = sum(ind_MSF_malnutrition == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
            p_ind_MSF_malnutrition_death = n_ind_MSF_malnutrition_death / n_ind_MSF_malnutrition,
            a_ind_MSF_malnutrition_death = format_nbp(n_ind_MSF_malnutrition_death, p_ind_MSF_malnutrition_death),
            
  )

gtble17 <- tble17 %>%
  select(country, project, starts_with("a_"))  %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    
    a_coinf       = 'N' ,
    a_coinf_hospi = 'Hospitalised',
    a_coinf_death = 'Deceased',
    
    a_ind_MSF_malaria       = 'N' ,
    a_ind_MSF_malaria_hospi = 'Hospitalised',
    a_ind_MSF_malaria_death = 'Deceased',
    
    a_hiv       = 'N',
    a_hiv_hospi = 'Hospitalised',
    a_hiv_death = 'Deceased',
    
    a_ind_MSF_tb_active       = 'N',
    a_ind_MSF_tb_active_hospi = 'Hospitalised',
    a_ind_MSF_tb_active_death = 'Deceased',
    
    a_tb_past        = 'N',
    a_tb_past_hospi  = 'Hospitalised',
    a_tb_past_death = 'Deceased',
    
    a_ind_MSF_malnutrition       = 'N',
    a_ind_MSF_malaria_hospi      = 'Hospitalised',
    a_ind_MSF_malnutrition_death = 'Deceased') %>%
  
  tab_spanner(
    label = 'At least one documented co-infection',
    columns = vars(a_coinf, a_coinf_hospi, a_coinf_death)) %>%
  tab_spanner(
    label = 'Malaria',
    columns = contains("_ind_MSF_malaria")) %>%
  tab_spanner(
    label = 'HIV',
    columns = contains("_hiv")) %>%
  tab_spanner(
    label = 'Tuberculosis active',
    columns = contains("_ind_MSF_tb_active")) %>%
  tab_spanner(
    label = 'Tuberculosis treatment in the past',
    columns = contains("_tb_past")) %>%
  tab_spanner(
    label = 'Malnutrition',
    columns = contains("_ind_MSF_malnutrition")) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)

gtsave(gtble17,
       file.path(path.local.msf.tables.oc, paste0('gtble17', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble17 %>%
  tab_header(
    title = 'Number of patients with coinfections and proportion of hospitalizations and death')
```






## Intervention received
### ICU

```{r tbl-icu}
tble7 <- dta_linelist_project %>%
  select(country, project, merge_admit, patcourse_icu, merge_icu) %>%
  group_by(country, project) %>%
  summarise(
    n_patients = n(),
    n_admis    = sum(merge_admit == "Yes", na.rm = TRUE),
    
    n_icu_admission = sum(merge_admit == "Yes" & patcourse_icu == "Yes", na.rm = TRUE),
    p_icu_admission = n_icu_admission / n_admis,
    m_icu_admission = format_nbp(n_icu_admission, p_icu_admission),
    
    n_icu_later  = sum(merge_admit == "Yes" & patcourse_icu != "Yes" & merge_icu == "Yes", na.rm = TRUE),
    p_icu_later = n_icu_later / n_admis,
    m_icu_later = format_nbp(n_icu_later, p_icu_later)
  ) %>%
  select(country, project, n_patients, n_admis,
         m_icu_admission,  m_icu_later)



gtble7 <-  tble7 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    n_patients = 'Visits',
    n_admis    = 'Admissions',
    
    m_icu_admission = 'ICU at adm.',
    m_icu_later     = 'ICU later',
  ) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)

gtsave(gtble7,
       file.path(path.local.msf.tables.oc, paste0('gtble7', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble7 %>%
  tab_header(
    title = 'Total number of patient visits, admitted, and admitted in ICU')
```


```{r tbl-icu-status}
tble8 <- dta_linelist_project %>%
  select(country, project, ind_MSF_covid_status, merge_admit, patcourse_icu, merge_icu) %>%
  group_by(country, project) %>%
  summarise(
    n_patients = n(),
    n_admis    = sum(merge_admit == "Yes", na.rm = TRUE),
    
    n_probable            = sum(ind_MSF_covid_status == "Probable", na.rm = TRUE),
    n_probable_admis      = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes", na.rm = TRUE),
    p_probable_admis      = n_probable_admis / n_probable,
    m_probable_admis      = format_nbp(n_probable_admis, p_probable_admis),
    
    n_probable_icu_admis  = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes" & patcourse_icu == "Yes", na.rm = TRUE),
    p_probable_icu_admis  = n_probable_icu_admis / n_probable_admis,
    m_probable_icu_admis  = format_nbp(n_probable_icu_admis, p_probable_icu_admis),
    
    n_probable_icu_later  = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes" & patcourse_icu != "Yes" & merge_icu == "Yes", na.rm = TRUE),
    p_probable_icu_later  = n_probable_icu_later / n_probable_admis,
    m_probable_icu_later  = format_nbp(n_probable_icu_later, p_probable_icu_later),
    
    n_suspected            = sum(ind_MSF_covid_status == "Suspected", na.rm = TRUE),
    n_suspected_admis      = sum(ind_MSF_covid_status == "Suspected" & merge_admit == "Yes", na.rm = TRUE),
    p_suspected_admis      = n_suspected_admis / n_suspected,
    m_suspected_admis      = format_nbp(n_suspected_admis, p_suspected_admis),
    
    n_suspected_icu_admis  = sum(ind_MSF_covid_status == "Suspected" & merge_admit == "Yes" & patcourse_icu == "Yes", na.rm = TRUE),
    p_suspected_icu_admis  = n_suspected_icu_admis / n_suspected_admis,
    m_suspected_icu_admis  = format_nbp(n_suspected_icu_admis, p_suspected_icu_admis),
    
    n_suspected_icu_later  = sum(ind_MSF_covid_status == "Suspected" & merge_admit == "Yes" & patcourse_icu != "Yes" & merge_icu == "Yes", na.rm = TRUE),
    p_suspected_icu_later  = n_suspected_icu_later / n_suspected_admis,
    m_suspected_icu_later  = format_nbp(n_suspected_icu_later, p_suspected_icu_later),
    
    n_confirmed            = sum(ind_MSF_covid_status == "Confirmed", na.rm = TRUE),
    n_confirmed_admis      = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes", na.rm = TRUE),
    p_confirmed_admis      = n_confirmed_admis / n_confirmed,
    m_confirmed_admis      = format_nbp(n_confirmed_admis, p_confirmed_admis),
    
    n_confirmed_icu_admis  = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes" & patcourse_icu == "Yes", na.rm = TRUE),
    p_confirmed_icu_admis  = n_confirmed_icu_admis / n_confirmed_admis,
    m_confirmed_icu_admis  = format_nbp(n_confirmed_icu_admis, p_confirmed_icu_admis),
    
    n_confirmed_icu_later  = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes" & patcourse_icu != "Yes" & merge_icu == "Yes", na.rm = TRUE),
    p_confirmed_icu_later  = n_confirmed_icu_later / n_confirmed_admis,
    m_confirmed_icu_later  = format_nbp(n_confirmed_icu_later, p_confirmed_icu_later),
  ) %>%
  select(country, project, n_patients,
         # n_admis,
         n_probable,  m_probable_admis,  m_probable_icu_admis,  m_probable_icu_later,
         n_suspected, m_suspected_admis, m_suspected_icu_admis, m_suspected_icu_later,
         n_confirmed, m_confirmed_admis, m_confirmed_icu_admis, m_confirmed_icu_later)




gtble8 <-  tble8 %>%
  gt() %>%
  cols_label(
    country          = 'Country',
    project          = 'Project',
    n_patients       = 'Visits',
    # n_admited        = 'Admissions',
    
    n_probable            = 'Visits',
    m_probable_admis      = 'Hospit.',
    m_probable_icu_admis  = 'ICU at adm.',
    m_probable_icu_later  = 'ICU after adm.',
    
    n_suspected            = 'Visits',
    m_suspected_admis      = 'Hospit.',
    m_suspected_icu_admis  = 'ICU at adm.',
    m_suspected_icu_later  = 'ICU after adm.',
    
    n_confirmed            = 'Visits',
    m_confirmed_admis      = 'Hospit.',
    m_confirmed_icu_admis  = 'ICU at adm.',
    m_confirmed_icu_later  = 'ICU after adm.') %>%
  
  tab_spanner(
    label = 'Probable',
    columns = contains("_probable")) %>%
  tab_spanner(
    label = 'Suspected',
    columns = contains('_suspected')) %>%
  tab_spanner(
    label = 'Confirmed',
    columns = contains('_confirmed')) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)

gtsave(gtble8,
       file.path(path.local.msf.tables.oc, paste0('gtble8', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble8 %>%
  tab_header(
    title = 'Total number of patient visits by case status and ICU admission status')

```


### Oxygen


```{r tbl-oxy}
tble9 <- dta_linelist_project %>%
  select(country, project, merge_admit, MSF_received_oxygen, merge_oxygen) %>%
  group_by(country, project) %>%
  summarise(
    n_patients = n(),
    n_admis    = sum(merge_admit == "Yes", na.rm = TRUE),
    
    n_oxygen_admission = sum(merge_admit == "Yes" & MSF_received_oxygen == "Yes", na.rm = TRUE),
    p_oxygen_admission = n_oxygen_admission / n_admis,
    m_oxygen_admission = format_nbp(n_oxygen_admission, p_oxygen_admission),
    
    n_oxygen_later  = sum(merge_admit == "Yes" & MSF_received_oxygen != "Yes" & merge_oxygen == "Yes", na.rm = TRUE),
    p_oxygen_later  = n_oxygen_later / n_admis,
    m_oxygen_later  = format_nbp(n_oxygen_later, p_oxygen_later)
  ) %>%
  select(country, project, n_patients, n_admis,
         m_oxygen_admission,  m_oxygen_later)




gtble9 <-  tble9 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    n_patients = 'Visits',
    n_admis    = 'Admissions',
    
    m_oxygen_admission = 'Oxygen at adm.',
    m_oxygen_later     = 'Oxygen later',
  ) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)

gtsave(gtble9,
       file.path(path.local.msf.tables.oc, paste0('gtble9', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble9 %>%
  tab_header(
    title = 'Total number of patient visits and oxygen status')

```


```{r tbl-oxy-status}
tble10 <- dta_linelist_project %>%
  select(country, project, ind_MSF_covid_status, merge_admit, MSF_received_oxygen, merge_oxygen) %>%
  group_by(country, project) %>%
  summarise(
    n_patients = n(),
    n_admis    = sum(merge_admit == "Yes", na.rm = TRUE),
    
    n_probable            = sum(ind_MSF_covid_status == "Probable", na.rm = TRUE),
    n_probable_admis      = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes", na.rm = TRUE),
    p_probable_admis      =  n_probable_admis / n_probable,
    m_probable_admis         = format_nbp(n_probable_admis, p_probable_admis),
    
    n_probable_oxygen_admis  = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes" & MSF_received_oxygen == "Yes", na.rm = TRUE),
    p_probable_oxygen_admis  = n_probable_oxygen_admis / n_probable_admis,
    m_probable_oxygen_admis  = format_nbp(n_probable_oxygen_admis, p_probable_oxygen_admis),
    
    n_probable_oxygen_later  = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes" & MSF_received_oxygen != "Yes" & merge_oxygen == "Yes", na.rm = TRUE),
    p_probable_oxygen_later  = n_probable_oxygen_later / n_probable_admis,
    m_probable_oxygen_later  = format_nbp(n_probable_oxygen_later, p_probable_oxygen_later),
    
    n_suspected            = sum(ind_MSF_covid_status == "Suspected", na.rm = TRUE),
    n_suspected_admis      = sum(ind_MSF_covid_status == "Suspected" & merge_admit == "Yes", na.rm = TRUE),
    p_suspected_admis      = n_suspected_admis / n_suspected,
    m_suspected_admis      = format_nbp(n_suspected_admis, p_suspected_admis),
    
    n_suspected_oxygen_admis  = sum(ind_MSF_covid_status == "Suspected" & merge_admit == "Yes" & MSF_received_oxygen == "Yes", na.rm = TRUE),
    p_suspected_oxygen_admis  = n_suspected_oxygen_admis / n_suspected_admis,
    m_suspected_oxygen_admis  = format_nbp(n_suspected_oxygen_admis, p_suspected_oxygen_admis),
    
    n_suspected_oxygen_later  = sum(ind_MSF_covid_status == "Suspected" & merge_admit == "Yes" & MSF_received_oxygen != "Yes" & merge_oxygen == "Yes", na.rm = TRUE),
    p_suspected_oxygen_later  = n_suspected_oxygen_later / n_suspected_admis,
    m_suspected_oxygen_later  = format_nbp(n_suspected_oxygen_later, p_suspected_oxygen_later),
    
    
    n_confirmed            = sum(ind_MSF_covid_status == "Confirmed", na.rm = TRUE),
    n_confirmed_admis      = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes", na.rm = TRUE),
    p_confirmed_admis      = n_confirmed_admis / n_confirmed,
    m_confirmed_admis      = format_nbp(n_confirmed_admis, p_confirmed_admis),
    
    n_confirmed_oxygen_admis  = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes" & MSF_received_oxygen == "Yes", na.rm = TRUE),
    p_confirmed_oxygen_admis  = n_confirmed_oxygen_admis / n_confirmed_admis,
    m_confirmed_oxygen_admis  = format_nbp(n_confirmed_oxygen_admis, p_confirmed_oxygen_admis),
    
    n_confirmed_oxygen_later  = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes" & MSF_received_oxygen != "Yes" & merge_oxygen == "Yes", na.rm = TRUE),
    p_confirmed_oxygen_later  = n_confirmed_oxygen_later / n_confirmed_admis,
    m_confirmed_oxygen_later  = format_nbp(n_confirmed_oxygen_later, p_confirmed_oxygen_later),
  ) %>%
  select(country, project, n_patients,
         n_probable,  m_probable_admis,  m_probable_oxygen_admis,  m_probable_oxygen_later,
         n_suspected, m_suspected_admis, m_suspected_oxygen_admis, m_suspected_oxygen_later,
         n_confirmed, m_confirmed_admis, m_confirmed_oxygen_admis, m_confirmed_oxygen_later)



gtble10 <- tble10 %>%
  gt() %>%
  cols_label(
    country          = 'Country',
    project          = 'Project',
    n_patients       = 'Visits',
    
    n_probable               = 'Visits',
    m_probable_admis         = 'Hospit.',
    m_probable_oxygen_admis  = 'Oxygen at adm.',
    m_probable_oxygen_later  = 'Oxygen after adm.',
    
    n_suspected               = 'Visits',
    m_suspected_admis         = 'Hospit.',
    m_suspected_oxygen_admis  = 'Oxygen at adm.',
    m_suspected_oxygen_later  = 'Oxygen after adm.',
    
    n_confirmed               = 'Visits',
    m_confirmed_admis         = 'Hospit.',
    m_confirmed_oxygen_admis  = 'Oxygen at adm.',
    m_confirmed_oxygen_later  = 'Oxygen after adm.') %>%
  
  tab_spanner(
    label = 'Probable',
    columns = contains("_probable")) %>%
  tab_spanner(
    label = 'Suspected',
    columns = contains('_suspected')) %>%
  tab_spanner(
    label = 'Confirmed',
    columns = contains('_confirmed')) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)


gtsave(gtble10,
       file.path(path.local.msf.tables.oc, paste0('gtble10', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble10 %>%
  tab_header(
    title = 'Total number of patient visits by case status and received_oxygen admission status')
```




### Ventilation

```{r tbl-vent}
tble12 <- dta_linelist_project %>%
  select(country, project, merge_admit, patcourse_vent, merge_vent) %>%
  group_by(country, project) %>%
  summarise(
    n_patients = n(),
    n_admis    = sum(merge_admit == "Yes", na.rm = TRUE),
    
    n_vent_admission = sum(merge_admit == "Yes" & patcourse_vent == "Yes", na.rm = TRUE),
    p_vent_admission = n_vent_admission / n_admis,
    m_vent_admission = format_nbp(n_vent_admission, p_vent_admission),
    
    n_vent_later  = sum(merge_admit == "Yes" & patcourse_vent != "Yes" & merge_vent == "Yes", na.rm = TRUE),
    p_vent_later  = n_vent_later / n_admis,
    m_vent_later  = format_nbp(n_vent_later, p_vent_later)
  ) %>%
  select(country, project, n_patients, n_admis,
         m_vent_admission,  m_vent_later)


gtble12 <-  tble12 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    n_patients = 'Visits',
    n_admis    = 'Admissions',
    
    m_vent_admission = 'Ventilation at adm.',
    m_vent_later     = 'Ventilation later',
  ) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)

gtsave(gtble12,
       file.path(path.local.msf.tables.oc, paste0('gtble12', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble12 %>%
  tab_header(
    title = 'Total number of patient visits and ventilation status')

```


```{r tbl-vent-statu}
tble13 <- dta_linelist_project %>%
  select(country, project, ind_MSF_covid_status, merge_admit, patcourse_vent, merge_vent) %>%
  group_by(country, project) %>%
  summarise(
    n_patients = n(),
    n_admis    = sum(merge_admit == "Yes", na.rm = TRUE),
    
    n_probable            = sum(ind_MSF_covid_status == "Probable", na.rm = TRUE),
    n_probable_admis      = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes", na.rm = TRUE),
    p_probable_admis      = n_probable_admis / n_probable,
    m_probable_admis      = format_nbp(n_probable_admis, p_probable_admis),
    
    n_probable_vent_admis  = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes" & patcourse_vent == "Yes", na.rm = TRUE),
    p_probable_vent_admis  = n_probable_vent_admis / n_probable_admis,
    m_probable_vent_admis  = format_nbp(n_probable_vent_admis, p_probable_vent_admis),
    
    n_probable_vent_later  = sum(ind_MSF_covid_status == "Probable" & merge_admit == "Yes" & patcourse_vent != "Yes" & merge_vent == "Yes", na.rm = TRUE),
    p_probable_vent_later  = n_probable_vent_later / n_probable_admis,
    m_probable_vent_later  = format_nbp(n_probable_vent_later, p_probable_vent_later),
    
    n_suspected            = sum(ind_MSF_covid_status == "Suspected", na.rm = TRUE),
    n_suspected_admis      = sum(ind_MSF_covid_status == "Suspected" & merge_admit == "Yes", na.rm = TRUE),
    p_suspected_admis      = n_suspected_admis / n_suspected,
    m_suspected_admis      = format_nbp(n_suspected_admis, p_suspected_admis),
    
    n_suspected_vent_admis  = sum(ind_MSF_covid_status == "Suspected" & merge_admit == "Yes" & patcourse_vent == "Yes", na.rm = TRUE),
    p_suspected_vent_admis  = n_suspected_vent_admis / n_suspected_admis,
    m_suspected_vent_admis  = format_nbp(n_suspected_vent_admis, p_suspected_vent_admis),
    
    n_suspected_vent_later  = sum(ind_MSF_covid_status == "Suspected" & merge_admit == "Yes" & patcourse_vent != "Yes" & merge_vent == "Yes", na.rm = TRUE),
    p_suspected_vent_later  = n_suspected_vent_later / n_suspected_admis,
    m_suspected_vent_later  = format_nbp(n_suspected_vent_later, p_suspected_vent_later),
    
    
    n_confirmed            = sum(ind_MSF_covid_status == "Confirmed", na.rm = TRUE),
    n_confirmed_admis      = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes", na.rm = TRUE),
    p_confirmed_admis      = n_confirmed_admis / n_confirmed,
    m_confirmed_admis      = format_nbp(n_confirmed_admis, p_confirmed_admis),
    
    n_confirmed_vent_admis  = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes" & patcourse_vent == "Yes", na.rm = TRUE),
    p_confirmed_vent_admis  = n_confirmed_vent_admis / n_confirmed_admis,
    m_confirmed_vent_admis  = format_nbp(n_confirmed_vent_admis, p_confirmed_vent_admis),
    
    n_confirmed_vent_later  = sum(ind_MSF_covid_status == "Confirmed" & merge_admit == "Yes" & patcourse_vent != "Yes" & merge_vent == "Yes", na.rm = TRUE),
    p_confirmed_vent_later  = n_confirmed_vent_later / n_confirmed_admis,
    m_confirmed_vent_later  = format_nbp(n_confirmed_vent_later, p_confirmed_vent_later),
  ) %>%
  select(country, project, n_patients,
         n_probable,  m_probable_admis,  m_probable_vent_admis,  m_probable_vent_later,
         n_suspected, m_suspected_admis, m_suspected_vent_admis, m_suspected_vent_later,
         n_confirmed, m_confirmed_admis, m_confirmed_vent_admis, m_confirmed_vent_later)



gtble13 <- tble13 %>%
  gt() %>%
  cols_label(
    country    = 'Country',
    project    = 'Project',
    n_patients = 'Visits',
    
    n_probable             = 'Visits',
    m_probable_admis       = 'Hospit.',
    m_probable_vent_admis  = 'Ventilation at adm.',
    m_probable_vent_later  = 'Ventilation after adm.',
    
    n_suspected             = 'Visits',
    m_suspected_admis       = 'Hospit.',
    m_suspected_vent_admis  = 'Ventilation at adm.',
    m_suspected_vent_later  = 'Ventilation after adm.',
    
    n_confirmed             = 'Visits',
    m_confirmed_admis       = 'Hospit.',
    m_confirmed_vent_admis  = 'Ventilation at adm.',
    m_confirmed_vent_later  = 'Ventilation after adm.') %>%
  
  tab_spanner(
    label = 'Probable',
    columns = contains("_probable")) %>%
  tab_spanner(
    label = 'Suspected',
    columns = contains('_suspected')) %>%
  tab_spanner(
    label = 'Confirmed',
    columns = contains('_confirmed')) %>%
  
  fmt_missing(
    columns = starts_with('n_'),
    missing_text = "0") %>%
  
  cols_align(
    align = 'center',
    columns = everything()) %>%
  
  tab_options(
    column_labels.font.weight = "bold",
    row_group.font.weight = "bold",
    data_row.padding = px(2),
    row_group.padding = px(2),
    row.striping.include_table_body = TRUE)  %>%
  
  summary_rows(
    columns = starts_with("n_"),
    fns = list(Total = "sum"),
    decimals = 0)


gtsave(gtble13,
       file.path(path.local.msf.tables.oc, paste0('gtble13', '_', week_report, '.html')),
       inline_css = TRUE) %>%
  invisible()


gtble13 %>%
  tab_header(
    title = 'Total number of patient visits by case status and ventilation status')
```



### Mortality by intervention received

```{r fig6, fig.width = 8, fig.height = 8, , fig.cap = "Mortality of hospitalised patients by therapy received and by project"}
fig6 <- dta_linelist_project %>%
  filter(merge_admit == "Yes") %>%
  
  group_by(continent, country, project) %>%
  
  summarise(
    n_admit = n(),
    
    n_oxy       = sum(merge_oxygen == "Yes", na.rm = TRUE),
    n_death_oxy = sum(merge_oxygen == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    p_Oxygen    = round(100 * n_death_oxy / n_oxy, 1),
    
    n_vent        = sum(merge_vent == "Yes", na.rm = TRUE),
    n_death_vent  = sum(merge_vent == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    p_Ventilated = round(100 * n_death_vent / n_vent, 1),
    
    n_icu       = sum(merge_icu == "Yes", na.rm = TRUE),
    n_death_icu = sum(merge_icu == "Yes" & ind_outcome_patcourse_status == "Died", na.rm = TRUE),
    p_ICU       = round(100 * n_death_icu / n_icu, 1)
  ) %>%
  mutate(project = paste(country, project, sep = " - ")) %>%
  pivot_longer(cols = c(p_Oxygen, p_Ventilated, p_ICU),
               names_to = "therapy",
               names_prefix = "p_",
               values_to = "mortality") %>%
  mutate(therapy = factor(therapy,
                          levels = c("Oxygen", "Ventilated", "ICU")))  %>%
  
  ggplot(aes(x = reorder(project, desc(country)),
             y = mortality,
             size = n_admit,
             colour = therapy)) +
  geom_point(alpha = 0.8,
             position = position_dodge2(0.2)) +
  coord_flip() +
  facet_wrap(~therapy, ncol = 1, scales = "free") +
  labs(x = "",
       y = "Mortality (% of deceased among cases)",
       title = "Mortality by therapy received, by project") +
  
  ggthemes::scale_colour_tableau(name = "Therapy received", palette = "Tableau 20") +
  
  scale_size_continuous(name = "Number of hosp. cases\nin project",
                        # breaks = c(10, 100, 1000, 4000, 6000)
  ) +
  
  theme_light() +
  theme(legend.position = 'right',
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())


ggsave(file.path(path.local.msf.graphs.oc, paste0('fig6', '_', week_report, '.png')),
       plot = fig6,
       width = 8,
       height = 10,
       scale = 1.1,
       dpi = 320)

fig6
```



## Aditional figures by project

```{r include=FALSE}
tbl_plot1 <- dta_linelist_with_aggregated %>%
  mutate(project2 = stringr::str_remove(project,
                                        pattern = " \\(\\*\\)") %>%
           stringr::str_replace_all(" ", "_") %>% 
           stringr::str_replace_all("_-_", "_") %>% 
           stringr::str_replace_all("\\(", "") %>% 
           stringr::str_replace_all("\\)", "") %>% 
           stringr::str_replace_all("-", "") %>% 
           stringr::str_replace_all("/", "_")) %>%   
  mutate(country_project = paste(iso_a3, project2, sep = "_") ) %>% 
  count(country_project, ind_MSF_covid_status, epi_week_consultation) %>%
  drop_na(epi_week_consultation) %>% 
  group_by(country_project) %>% 
  nest() %>%
  mutate(plot1 = map2(.x = data,
                      .y = country_project,
                      ~ fct_plot_1(data = .x,
                                   country_project = .y,
                                   path_save = path.local.msf.graphs.oc.per.project)))

# print(tbl_plot1$plot1)



tbl_plot <- dta_linelist_project %>%
  group_by(country_project) %>%
  nest() %>%
  mutate(plot2 = map2(.x = data,
                      .y = country_project,
                      ~ fct_plot_2(data = .x,
                                   country_project = .y,
                                   path_save = path.local.msf.graphs.oc.per.project))) %>% 
  mutate(plot3 = map2(.x = data,
                      .y = country_project,
                      ~ fct_plot_3(data = .x,
                                   country_project = .y,
                                   path_save = path.local.msf.graphs.oc.per.project)))


# print(tbl_plot$plot2)
# print(tbl_plot$plot3)

a <- left_join(tbl_plot1, 
               tbl_plot, by = "country_project" ) %>% 
  select(plot1, plot2, plot3)
```


```{r echo=FALSE, message=FALSE, results ='asis', fig.width=5, fig.height=4}

for(i in 1:nrow(a)){
  
  cat('\n')
  cat('\n###', a[[i, "country_project"]], '\n')
  
  print_plot(a[[i, "plot1"]][[1]])
  print_plot(a[[i, "plot2"]][[1]])
  print_plot(a[[i, "plot3"]][[1]])

  cat('\n')
  cat('\n')
  cat('\n')
}
```

